### 2014-03-18 ms.tex
Initial draft

%\documentclass[10pt,iop]{emulateapj}
\documentclass[preprint]{aastex}

\usepackage{url}
\usepackage{multirow}
\usepackage{amsmath}
\usepackage{xcolor}
%\citestyle{aa}

%\bibliographystyle{apj_w_etal}

\newcommand{\etal}{{et al.\/}}
\newcommand{\Prob}{\mathtt{P}}
\newcommand{\logL}{\log\mathcal{L}}
\newcommand{\unit}[1]{\footnotesize #1}
\newcommand{\PAPER}{\mathrm{PAPER}}
\bibliographystyle{apj_w_etal}

\newcommand{\Nconf}{31}
\newcommand{\Nsrc}{32}
\definecolor{orange}{RGB}{255,127,0}

	% End definitions

%\slugcomment{DRAFT: \today}

\shorttitle{EoR time}
\shortauthors{Jacobs et al.}

\begin{document}

\title{The PAPER-32 power spectrum measured at several redshifts}
\author{
Daniel C. Jacobs\altaffilmark{1},
Aaron R. Parsons\altaffilmark{2,8},
James E. Aguirre\altaffilmark{3},
Zaki Ali\altaffilmark{2},
Judd Bowman\altaffilmark{1},
Richard F. Bradley\altaffilmark{4,5,6},
Chris L.  Carilli\altaffilmark{7},
David R. DeBoer\altaffilmark{8},
Matthew R. Dexter\altaffilmark{8},
Nicole E. Gugliucci\altaffilmark{5},
Pat Klima\altaffilmark{5},
Dave H. E. MacMahon\altaffilmark{8}
Jason R. Manley\altaffilmark{9},
David F. Moore\altaffilmark{3},
Jonathan C. Pober\altaffilmark{2},
Irina I. Stefan\altaffilmark{10},
William P. Walbrugh\altaffilmark{9}}

\altaffiltext{1}{School of Earth and Space Exploration, Arizona State U., Tempe, AZ}
\altaffiltext{2}{Astronomy Dept., U. California, Berkeley, CA}
\altaffiltext{3}{Dept. of Physics and Astronomy, U. Pennsylvania, Philadelphia, PA}
\altaffiltext{4}{Dept. of Electrical and Computer Engineering, U. Virginia, Charlottesville, VA}
\altaffiltext{5}{National Radio Astronomy Obs., Charlottesville, VA}
\altaffiltext{6}{Dept. of Astronomy, U. Virginia, Charlottesville, VA}
\altaffiltext{7}{National Radio Astronomy Obs., Socorro, NM}
\altaffiltext{8}{Radio Astronomy Lab., U. California, Berkeley, CA}
\altaffiltext{9}{Square Kilometer Array, South Africa Project, Cape Town, South Africa}
\altaffiltext{10}{Cavendish Lab., Cambridge, UK}

\begin{abstract}
We present new observations from the Donald C. Backer Precision Array for Probing the Epoch of Reionization telescope which probe the redshift range of $6<z<9$, extending previously published single redshift results to cover the full range accessible to the instrument.  The removal of foregrounds depends heavily on the spectral smoothness of both instrumental response and foregrounds.  These spectra demonstrate a robust foreground removal of foreground signals to the thermal noise limit, leaving systematics on the scale of the noise. We find that these residual systematics are present throughout the observed frequency band are most likely some mixture residual foregrounds and instrumental systematics.
\end{abstract}

\keywords{reionization}

\section{Introduction}
The Epoch of Reionization (EoR), when the first stars ionized the pervasive cosmological Hydrogen in the last global phase change, is predicted to be observable in highly redshifted 21 cm radiation.  The Donald C. Backer Precision Array for Probing the Epoch of Reionization (PAPER, XXX)\footnote{\url{http:\\eor.berkeley.edu}} is a low frequency radio interferometer experiment dedicated to opening this window on the universe.  Challenges include foregrounds which are brighter by several orders of magnitude and limited collecting areas of first generation instruments which necessitate long integration times. Direct observation of hydrogen before and during re-ionization is predicted to deliver a wealth of cosmological and astrophysical data, including the nature of the first stellar objects and the timing and rate of galaxy formation, reviews on the physics of reionization as well as theory about the nature of foregrounds maybe be found in \citet{Furlanetto:2006p2267,Morales:2010p8093,Pritchard:2012p9555}.
Telescopes seeking to measure this signal include the Giant Metre-wave Radio Telescope (GMRT; \cite{Paciga:2013p9943}), the Low Frequency Array (LOFAR; \cite{Yatawatta:2013p9699}) and the Murchison Widefield Array (MWA; \cite{Bowman:2013p9950} and \cite{Tingay:2013p9022}). PAPER is located in the Karoo desert at the Square Kilometer Array-South Africa cite. PAPER has doubled in size on a yearly basis since 2009, making  EoR observations with each array.  Here we report on deep integrations made with a 32 element array in 2011, first described in \cite{Parsons:2013p9876}, hereafter P14.  In that work we described in detail a method used to estimate power spectra in the presence of bright foregrounds. Using this method we combined 120 XXX of data to place upper limits on the power spectrum of EoR. Here we extend this analysis of the same data include multiple redshift bins allowing us to place constraints on the time history of reionization.

This paper summarizes the observations and reduction methodology in \ref{sec:obs_meth}, present the new upper limits in \ref{sec:results}. In \ref{sec:models} we examine the impacts of the new data points on theory and conclude in \ref{sec:conclusion}.

%
%This signal is though to be the richest data set on the sky!  Many telescopes are searching for the power spectral signature of 
%HI.  Coarsely, models of 21cm emission can be distilled into two parameters. The redshift at which point the universe was 50\% ionized ($z_i$) and the transition time from mostly neutral to mostly ionized $dz$. Thus full coverage of the spectrum is essential.

%This paper extends the result ARP2013a to cover the redshift range XXX.

\section{Observations and reduction}
As the observations and methods in use are unchanged from those described in P14, we will offer a quick summary and refer the reader to \cite{Parsons:2013p9876} for a more in depth discussion.  A general overview of the PAPER system can be found in \cite{Parsons:2010p6757}, calibration of the primary beam in \cite{Pober:2012p8800}, and imaging results in \cite{Jacobs:2011p8438,Jacobs:2013p10014,Stefan:2013p9926}.  Sensitivity analysis described in \cite{Parsons:2012p9028} revealed that for the low gain elements employed by PAPER, a highly redundant ``grid'' type arrangement offers a significant sensitivity boost.  Each baseline samples a different mode of the EoR power spectrum. Each redundant baseline one samples a cosmological mode at high SNR before being combined with other different $k$ modes.  The PAPER South Africa 32 antenna deployment (PSA32) was arranged in a 4x8 grid.  Here we include only the three shortest types of spacings, those between adjacent columns within at least one row of each other, a selection containing 70 30m baselines.

Observations spanning the band between 100 to 200-MHz ($13.1>z>6.1$) were recorded at a resolution of 48kHz into 10.7s bins beginning Dec 7, 2011 and ending XXX for a total of XXX nights.  WIthin this set we included observations in the LST range 1h30m - 6h30m where the sky dominated system temperature is at a minimum.

\section{Reduction}
Here we summarize our data reduction steps, for detailed descriptions see P14.

In several stages throughout the analysis process we take a 2D Fourier transform of the visibility spectra $V(t,\nu)$ into ``delay'/fringe rate' space.  In this space, smooth spectrum sources are physically localized to delays shorter than the light travel time length of the baseline (e.g. sources at the horizon, in the direction of the baseline vector) and fringe rates shorter than the sidereal rate.  In the Fourier space sources are highly localized but are convolved with the dual of the spectrum sampling function, which is uneven due to flagging of interference.  To account for this missing data we use a CLEAN like, iterative, peak finder and subtraction algorithm which is limited to finding peaks within the physically allowable ranges, and occasionally beyond, where necessary to account for spectral and temporal dispersion.

The data analysis pipeline essentially consists of iterative application of the delay/fringe rate transform process, with an ever tightening CLEAN box, interleaved with stages of averaging (time, night, and baseline), before finally computing a power spectrum.  This final step takes advantage of the redundant baselines to make an unbiased power spectrum estimate by cross-multiplying identical baselines.

\subsection{Initial Averaging}
  First, the raw data are down-selected to just the 30m long baselines. %I know this isn't technically true, but I think its important that the downselect is obvious.
   The visibilities are then compressed in the frequency and time directions by filtering delay modes above the horizon and highest frequency fringe rates for a 300m baseline.  This filtering is done in tandem with a radio frequency interference (RFI) flagging step, using the CLEAN residuals to flag 4$\sigma$ RFI in the absence of bright sky signals and then feeding the flags back into another iteration of the compression step. This mitigate the effects of RFI leaking around the filter to generate a cleaner averaged data set and results in a time and frequency bin of length 45.7s and width 400kHz. % XXX check  
  This process reduces the data volume by a factor of \~40.

\subsection{Calibration}
We model per-antenna complex gains with a phase slope and spectral amplitude.  Because the array samples correlations redundantly, the relative calibration between antennae is numerically overdetirmined and tractable as a linear algebra problem.  These solutions vary little over the observing period and exhibit less than 1\% rms variation. %XXX is that so???

By itself, the redundant solution is completely independent of any sky model. The redundant solution contains two free delay parameters and an overall amplitude scale. We fit the two delay parameters to a model of Pictor A, Fornax A, and the Crab Nebula during a time when the sky is dominated by these three sources while marginalizing over the unknown ratio between the three sources. With the delays in place we are now able form a beam on Pictor A and (for each channel) set the overall amplitude to the calibration value found in \cite{jacobs:2013b}.

%The success of the remainder of the processing stages hinges on the spectral stability of the data, which can easily be assessed by comparing redundant baselines. In figure XXX we see the rms of the ratios between redundant calibrations, before and after the calibration has been applied.  
\begin{figure}
\centering
\includegraphics[width=\textwidth]{figures/trms_by_baseline.png}
\caption{\label{fig:noise} The residual noise after filtering foregrounds is consistent with the theoretical noise (dotted line).  Here we estimate residual noise by differencing spectral channels and redundant baselines. The noise spectrum is characterized by an overall slope towards lower frequencies owing to the power law slope of the galactic spectrum which drives the system temperature.  At both high and low edges the noise is also driven up by the use of a hamming window in the foreground filter. Less is known about foregrounds at the edges of the band which drives up the residuals.}
\end{figure}
%XXX more, and a plot of those redundancy residuals
\subsection{Foregrounds}
Foregrounds are filtered from the calibrated data by removing all bright delay components with light travel times less then the baseline length. Where during the previous compression step a  horizon of 300m (1800ns)  was used to calculate the window size we now choose a window corresponding to the 30m baselines under study.  At this point a four hour long running mean is subtracted. This removes excess correlation due to cross-talk in the analog signal chain. The residuals are then flagged once more for RFI before the XXX nights of data are averaged into 10 minute long local sidereal time (LST) bins.  During averaging we found that some lst bins were dominated by very bright data points due to some non-noise-like data points. As these were very rare we found that removing the ten brightest points from every bin, a small median filter, was quite effective.   The root mean square of the residual signal (seen in Figure \ref{fig:noise}) at the end of this process is close to the 1mK level expected given the total integration time.

%This step is the first point at which signal loss could occur. To measure the loss we add white noise into the data at a level sufficient to double the system temperature (to make it easily detectable on the output), subtract  and then measure the fractional difference post filtration. We recover XXX\% of the injected signal.

%delay transform, covariance projection (lossless, and lossy), bin and bootstrap average,
\subsection{Power Spectrum}

\section{Discusion}

\begin{figure}
\includegraphics[width=\textwidth]{figures/psa32_pspec_k3pk_vs_z.png}
\caption{Power spectrum amplitude at k=0.3 hMpc$^-1$.  Paciga 2013 GMRT marked with an 'x', Parsons 2013 PAPER limit marked with black, this work marked with blue diamonds.}
\end{figure}

\section{Conclusion}

Here we have, using the same dataset extended the expanded the redshift coverage of \cite{Parsons:2013p9876} cover the range $10>z>7$.  These measurements demonstrate a precision removal of foregrounds to one part in ten million across the entire redshift range of interest. Though, with only 32 antenna, this data lacks the sensitivity to exclude all but the most extreme models, it does demonstrate the ability of the wideband filtration method to affectively remove foregrounds to the precision needed to integrate XXX nights to the thermal noise limit.  As future observations with PAPER  will add sensitivity primarily by increasing to 128 antennae this projects well for our ability to reach design sensitivity.

%\begin{align}
%\label{eq:BrightnessFluctuations}
% \Delta T &\approx \frac{T_S - T_{CMB}}{1 +z} \tau_{21} \nonumber\\
% &\approx T_0(z) \left( \frac{T_S - T_{CMB}}{T_S} \right) x_{HI} (1 + \delta) 
%\end{align}
%where $T_S$ is the 21cm line spin temperature, $x_{HI}$ is the local ionization fraction and $\delta$ is the mass over-density. All, save $T_0$, are position dependent.  $T_0$ encodes the global temperature evolution due to cosmological expansion
%\begin{align}
% T_0(z) & = 23~{\rm mK}~\left( \frac{\Omega_b h^2}{0.02} \right) \left[ \left(
% \frac{0.15}{\Omega_m h^2} \right) \left( \frac{1+z}{10} \right) \right]^{1/2}\nonumber\\
% & = 25~{\rm mK}~\left( \frac{1+z}{10} \right)^{1/2}
%\label{eq:Prefactor}
%\end{align}
%\citep[see, e.g., ][]{zaldarriaga_et_al2004,furlanetto_et_al2006}, where we have used the
%Planck 2013 parameters \citep{planck_et_al2013}.

\bibliography{library}

\end{document}

### 2014-03-27 ms.tex
Adding updated pspec vs z figures.

@@
 \subsection{Power Spectrum}
-
+\begin{figure}
+\centering
+\includegraphics[width=0.45\textwidth]{figures/pspec_Mar24_psa32_zbins_1_80_119_I.png}
+\includegraphics[width=0.45\textwidth]{figures/pspec_Mar24_psa32_zbins_1_100_139_I.png}
+\includegraphics[width=0.45\textwidth]{figures/pspec_Mar24_psa32_zbins_1_110_149_I.png}
+\includegraphics[width=0.45\textwidth]{figures/pspec_Mar24_psa32_zbins_1_119_150_I.png}
+\caption{\label{fig:pspecs} Clockwise from top left: power spectra centered on redshifts 8.5, 7.9, 7.6 and 7.5 (frequencies: 149.5, 159.5 164.5, and 167 MHz) at a bandwidth of 20MHz}
+\end{figure}
 \section{Discusion}
@@
 \begin{figure}
-\includegraphics[width=\textwidth]{figures/psa32_pspec_k3pk_vs_z.png}
-\caption{Power spectrum amplitude at k=0.3 hMpc$^-1$.  Paciga 2013 GMRT marked with an 'x', Parsons 2013 PAPER limit marked with black, this work marked with blue diamonds.}
+\includegraphics[width=\textwidth]{figures/psa_p3k_vs_z_28.png}
+\caption{Power spectrum amplitude at k=0.28 hMpc$^-1$.  Paciga 2013 GMRT marked with an 'x', Parsons 2013 PAPER limit marked with thin black, this work marked with thick blue diamonds.}
 \end{figure}
 
-\section{Conclusion}
+\section{Conclusions}
 

### 2014-03-28 ms.tex
Added more k bins and completed power spectrum method section.

@@
 PAPER is located in the Karoo desert at the Square Kilometer Array-South Africa cite.
-PAPER has doubled in size on a yearly basis since 2009, making  EoR observations with each array.
+PAPER has doubled in size on a yearly basis since 2009, making observations with each array.
 Here we report on deep integrations made with a 32 element array in 2011, first described in \cite{Parsons:2013p9876}, hereafter P14.
 In that work we described in detail a method used to estimate power spectra in the presence of bright foregrounds.
-Using this method we combined 120 XXX of data to place upper limits on the power spectrum of EoR.
-Here we extend this analysis of the same data include multiple redshift bins allowing us to place constraints on the time history of reionization.
+Using this method we combined 120 XXX of data to place upper limits on the power spectrum of HI in the early universe.
+Here we extend this analysis of the same data to  multiple redshift bins allowing us to place constraints throughout the likely time period of reionization.
 
-This paper summarizes the observations and reduction methodology in \ref{sec:obs_meth}, present the new upper limits in \ref{sec:results}.
-In \ref{sec:models} we examine the impacts of the new data points on theory and conclude in \ref{sec:conclusion}.
+This paper summarizes the observations and reduction methodology in \ref{sec:obs_meth}, present the new upper limits in \ref{sec:results} in \ref{sec:conclusion} offer some preliminary conclusions.
 
@@
 \section{Observations and reduction}
+\label{sec:observations}
 As the observations and methods in use are unchanged from those described in P14, we will offer a quick summary and refer the reader to \cite{Parsons:2013p9876} for a more in depth discussion.
@@
 \section{Reduction}
+\label{sec:obs_meth}
 Here we summarize our data reduction steps, for detailed descriptions see P14.
 
-In several stages throughout the analysis process we take a 2D Fourier transform of the visibility spectra $V(t,\nu)$ into ``delay'/fringe rate' space.
+In several stages throughout the analysis process we take a 2D Fourier transform of the visibility spectra $V(t,\nu)$ into ``delay'/fringe rate'' space.
 In this space, smooth spectrum sources are physically localized to delays shorter than the light travel time length of the baseline (e.g.
 sources at the horizon, in the direction of the baseline vector) and fringe rates shorter than the sidereal rate.
-In the Fourier space sources are highly localized but are convolved with the dual of the spectrum sampling function, which is uneven due to flagging of interference.
-To account for this missing data we use a CLEAN like, iterative, peak finder and subtraction algorithm which is limited to finding peaks within the physically allowable ranges, and occasionally beyond, where necessary to account for spectral and temporal dispersion.
+In this Fourier space sources are highly localized, deviations from a flat spectrum manifesting as a slight dispersion.
+In this space the spectrum sampling function, which is uneven due to flagging of interference takes the form of a convolution by a point-spread-function, much in the same way an imperfect sampling of the uv plane gives rise to the psf of an interferometer.
+To account for this we use a CLEAN like, iterative, peak finder and subtraction algorithm which is limited to finding peaks within the physically allowable ranges, and occasionally beyond, where necessary to account for spectral and temporal dispersion.
 
@@
 \subsection{Initial Averaging}
-  First, the raw data are down-selected to just the 30m long baselines. %I know this isn't technically true, but I think its important that the downselect is obvious.
+  First, the raw data are down-selected to just the 70 30m long baselines described in Section \ref{sec:observations}.
+%I know this isn't technically true, but I think its important that the downselect is obvious.
    The visibilities are then compressed in the frequency and time directions by filtering delay modes above the horizon and highest frequency fringe rates for a 300m baseline.
-This filtering is done in tandem with a radio frequency interference (RFI) flagging step, using the CLEAN residuals to flag 4$\sigma$ RFI in the absence of bright sky signals and then feeding the flags back into another iteration of the compression step.
-This mitigate the effects of RFI leaking around the filter to generate a cleaner averaged data set and results in a time and frequency bin of length 45.7s and width 400kHz.
-% XXX check
-  This process reduces the data volume by a factor of \~40.
+This filtering is done in tandem with a radio frequency interference (RFI) flagging step, using the CLEAN residuals to flag 4$\sigma$ deviations in the absence of bright sky signals and then feeding the flags back into another iteration of the compression step.
+This mitigate the effects of RFI leaking around the delay filter to generate a cleaner averaged data set and results in a time and frequency bin of length 39.6s and width 492.61kHz.
+  This process reduces the data volume by a factor of \~20, or roughly an order of magnitude improvement on traditional time and frequency averaging.
 
 \subsection{Calibration}
-We model per-antenna complex gains with a phase slope and spectral amplitude.
-Because the array samples correlations redundantly, the relative calibration between antennae is numerically overdetirmined and tractable as a linear algebra problem.
+We model the gain as a per-antenna amplitude, a phase slope -physically a single time delay- and a single passband for all antennae.
+Because the array samples correlations redundantly, the relative calibration between antennae is numerically overdetermined and tractable as a linear algebra problem.
 These solutions vary little over the observing period and exhibit less than 1\% rms variation.
-%XXX is that so???
+This method, as applied to this PAPER data set is discussed in depth by Zheng et al (in prep).
 
@@
 Foregrounds are filtered from the calibrated data by removing all bright delay components with light travel times less then the baseline length.
-Where during the previous compression step a  horizon of 300m (1800ns)  was used to calculate the window size we now choose a window corresponding to the 30m baselines under study.
-At this point a four hour long running mean is subtracted.
+Where during the previous compression step a liberal horizon of 300m (1800ns, much longer than the 30 meter baselines under study)  was used to calculate the window size, we now choose a window corresponding to the 30m baselines under study.
+The CLEAN model that is iteratively built is then subtracted from the data leaving residuals which we carry into the next stage.
+Next, a four hour long running mean is subtracted.
 This removes excess correlation due to cross-talk in the analog signal chain.
-The residuals are then flagged once more for RFI before the XXX nights of data are averaged into 10 minute long local sidereal time (LST) bins.
-During averaging we found that some lst bins were dominated by very bright data points due to some non-noise-like data points.
+The residuals are then flagged once more for RFI before the XXX nights of data are averaged into 36.4 second long local sidereal time (LST) bins.
+During averaging we found that some lst bins were dominated by very bright data points due to some exceedingly bright samples.
 As these were very rare we found that removing the ten brightest points from every bin, a small median filter, was quite effective.
@@
 \subsection{Power Spectrum}
+The output of the forgoing is a single calibrated and foreground filtered sidereal night.
+The power spectrum is estimated by selecting a 10MHz bandwidth range centered on the redshift of interest, performing a delay transform and cross correlating all the bins.
+To preserve the isolation of any foregrounds which remain, we increase the spectral range by 5MHz on each side and multiply by a blackman-harris window thus providing a much higher dynamic range delay spectrum psf.
+This leaves us with 40 delay samples on each of 70 baselines which are divided into three redundant groups.
+Within these groups we compute the  correlation between each mode.
+The cross correlation of the same delay modes between different redundant baselines provides our estimate of the power spectrum.
+Cross correlation between different modes provides our measure of systematic bias.
+These correlations should be identical between all redundant baselines to within the level of the noise.
+In practice, however there are some correlations deviate significantly from the median.
+These are removed by subtracting an average sky and then projecting out large residual modes.
+For more see Appendix of C of \cite{Parsons:2013p9876}.
+
+The residual correlations are all samples of the power spectrum.
+To estimate the final power spectrum and its uncertainty we compute the mean and variance of many random bootstrap-selected subsamples, sampling the dimensions: sidereal time, redundant baseline pair, and delay sign\footnote{As visibilities are complex, both the positive and negative delays  carry different information.}
+
 \begin{figure}
@@
 \includegraphics[width=0.45\textwidth]{figures/pspec_Mar24_psa32_zbins_1_119_150_I.png}
-\caption{\label{fig:pspecs} Clockwise from top left: power spectra centered on redshifts 8.5, 7.9, 7.6 and 7.5 (frequencies: 149.5, 159.5 164.5, and 167 MHz) at a bandwidth of 20MHz}
+\caption{\label{fig:pspecs} Right to left from top: power spectra centered on redshifts 8.5, 7.9, 7.6 and 7.5 (frequencies: 149.5, 159.5 164.5, and 167 MHz) at a bandwidth of 20MHz}
 \end{figure}
-\section{Discusion}
+\section{Results}
+\label{sec:results}
+In  Figure \ref{fig:pspecs} we see power spectra at different redshift bins, and in Figure \ref{fig:slices} we see power spectra at different $k$ modes as a function of redshift.
 
 \begin{figure}
-\includegraphics[width=\textwidth]{figures/psa_p3k_vs_z_28.png}
-\caption{Power spectrum amplitude at k=0.28 hMpc$^-1$.  Paciga 2013 GMRT marked with an 'x', Parsons 2013 PAPER limit marked with thin black, this work marked with thick blue diamonds.}
+\centering
+\includegraphics[width=0.48\textwidth]{figures/psa_p3k_vs_z_28.png}
+\includegraphics[width=0.48\textwidth]{figures/psa_p3k_vs_z_30.png}
+\includegraphics[width=0.48\textwidth]{figures/psa_p3k_vs_z_33.png}
+\includegraphics[width=0.48\textwidth]{figures/psa_p3k_vs_z_38.png}
+\caption{Power spectrum amplitude vs redshift at a selection of k modes.
+Right to left from top, k=0.28,0.3,0.33,0.38 hMpc$^-1$.
+Paciga 2013 GMRT marked with an 'x', Parsons 2013 PAPER limit marked with thin black, this work marked with thick blue diamonds.
+\label{fig:slices}}
 \end{figure}
@@
 \section{Conclusions}
-
+\label{sec:conclusion}
 Here we have, using the same dataset extended the expanded the redshift coverage of \cite{Parsons:2013p9876} cover the range $10>z>7$.

### 2014-04-24 ms.tex
adding new figures to tex and editing through to Discussion

@@
 Pat Klima\altaffilmark{5},
+Adrian Liu\altaffilmark{2},
 Dave H. E. MacMahon\altaffilmark{8}
@@
 \section{Introduction}
-The Epoch of Reionization (EoR), when the first stars ionized the pervasive cosmological Hydrogen in the last global phase change, is predicted to be observable in highly redshifted 21 cm radiation.
+The Epoch of Reionization, when the first stars ionized the pervasive cosmological Hydrogen in the last global phase change, is predicted to be observable in highly redshifted 21 cm radiation.
 The Donald C.
@@
 Direct observation of hydrogen before and during re-ionization is predicted to deliver a wealth of cosmological and astrophysical data, including the nature of the first stellar objects and the timing and rate of galaxy formation, reviews on the physics of reionization as well as theory about the nature of foregrounds maybe be found in \citet{Furlanetto:2006p2267,Morales:2010p8093,Pritchard:2012p9555}.
+
 Telescopes seeking to measure this signal include the Giant Metre-wave Radio Telescope (GMRT; \cite{Paciga:2013p9943}), the Low Frequency Array (LOFAR; \cite{Yatawatta:2013p9699}) and the Murchison Widefield Array (MWA; \cite{Bowman:2013p9950} and \cite{Tingay:2013p9022}).
@@
 PAPER has doubled in size on a yearly basis since 2009, making observations with each array.
+
 Here we report on deep integrations made with a 32 element array in 2011, first described in \cite{Parsons:2013p9876}, hereafter P14.
-In that work we described in detail a method used to estimate power spectra in the presence of bright foregrounds.
-Using this method we combined 120 XXX of data to place upper limits on the power spectrum of HI in the early universe.
-Here we extend this analysis of the same data to  multiple redshift bins allowing us to place constraints throughout the likely time period of reionization.
-
-This paper summarizes the observations and reduction methodology in \ref{sec:obs_meth}, present the new upper limits in \ref{sec:results} in \ref{sec:conclusion} offer some preliminary conclusions.
+In that work we described in detail out reduction method which enabled a deep limit on the power spectrum in the presence of bright foregrounds.
+Using this method we combined 92 nights of data to place upper limits on the power spectrum of HI in the early universe.
+However, this analysis was limited to a single redshift.
+The epoch of reionization signal is expected to evolve strongly with redshift, and it is this variation with cosmic history that will allow us to begin to place constraints on the physics of reionization \cite{Pritchard:2008p8123,Pober:2014p10390}.
+Therefore, while a detection of the 21 cm signal at even a single frequency would be a tremendous breakthrough, analysis techniques must be developed to capitalize on the wide bandwidths of the current generations of high redshift 21cm telescopes.
+The primary obstacle to detection, the wide bandwidth Using the same data as P14, this paper presents new power spectrum upper limits covering the redshift range $10.5>z>7.2$.
+In section  \ref{sec:obs_meth} we summarize the observation and reduction methodology, present the new upper limits in \ref{sec:results} in \ref{sec:conclusion} offer some conclusions and discussion of future work.
 
@@
 Sensitivity analysis described in \cite{Parsons:2012p9028} revealed that for the low gain elements employed by PAPER, a highly redundant ``grid'' type arrangement offers a significant sensitivity boost.
-Each baseline samples a different mode of the EoR power spectrum.
-Each redundant baseline one samples a cosmological mode at high SNR before being combined with other different $k$ modes.
+In an interferometer each baseline samples a different mode of the power spectrum.
+In a grid, there are many samples of each of these cosmological modes which can be averaged to a high sensitivity before being combined with other different $k$ modes.
 The PAPER South Africa 32 antenna deployment (PSA32) was arranged in a 4x8 grid.
-Here we include only the three shortest types of spacings, those between adjacent columns within at least one row of each other, a selection containing 70 30m baselines.
+Here we include only the three shortest types of spacings where the reionization power is expected to be brightest.
+This selection includes those between adjacent columns within at least one row of each other, a selection containing 70 30m baselines.
 
-Observations spanning the band between 100 to 200-MHz ($13.1>z>6.1$) were recorded at a resolution of 48kHz into 10.7s bins beginning Dec 7, 2011 and ending XXX for a total of XXX nights.
-WIthin this set we included observations in the LST range 1h30m - 6h30m where the sky dominated system temperature is at a minimum.
+Observations spanning the band between 100 to 200-MHz ($13.1>z>6.1$) were recorded at a resolution of 48kHz and 10.7s  beginning Dec 7, 2011 and ending March 19, 2012 for a total of 92 nights.
+WIthin this set we included observations in the LST range 1h - 9h where the sky dominated system temperature is at a minimum.
 
@@
 
-The data analysis pipeline essentially consists of iterative application of the delay/fringe rate transform process, with an ever tightening CLEAN box, interleaved with stages of averaging (time, night, and baseline), before finally computing a power spectrum.
-This final step takes advantage of the redundant baselines to make an unbiased power spectrum estimate by cross-multiplying identical baselines.
+The data analysis pipeline essentially consists of iterative application of the delay/fringe rate transform process, with an ever tightening allowable number of modes, interleaved with stages of averaging (time, frequency, night), before finally computing a power spectrum.
+This final step takes advantage of the redundant baselines to make an unbiased power spectrum estimate by cross-multiplying identical baselines and then averaging the power spectrum modes.
 
@@
   First, the raw data are down-selected to just the 70 30m long baselines described in Section \ref{sec:observations}.
-%I know this isn't technically true, but I think its important that the downselect is obvious.
+%I know this isn't technically true, but I think its important that the downselect is obvious up front.
    The visibilities are then compressed in the frequency and time directions by filtering delay modes above the horizon and highest frequency fringe rates for a 300m baseline.
-This filtering is done in tandem with a radio frequency interference (RFI) flagging step, using the CLEAN residuals to flag 4$\sigma$ deviations in the absence of bright sky signals and then feeding the flags back into another iteration of the compression step.
-This mitigate the effects of RFI leaking around the delay filter to generate a cleaner averaged data set and results in a time and frequency bin of length 39.6s and width 492.61kHz.
-  This process reduces the data volume by a factor of \~20, or roughly an order of magnitude improvement on traditional time and frequency averaging.
+This filtering is done in tandem with a radio frequency interference (RFI) flagging step, using the residuals which have had bright sky-like signals removed to flag 4$\sigma$ deviations before feeding the flags back into another iteration of the compression step.
+This mitigates the effects of bright, narrow band, interference  being scattered by the delay filter to generate a cleaner averaged data set and results in a time and frequency bin of length 39.6s and width 492.61kHz.
+  This process reduces the data volume by a factor of \~20, or roughly an order of magnitude improvement on traditional time and frequency averaging which in this array would be limited to ~100kHz and 10s to avoid averaging away fringes.
 
@@
 We model the gain as a per-antenna amplitude, a phase slope -physically a single time delay- and a single passband for all antennae.
-Because the array samples correlations redundantly, the relative calibration between antennae is numerically overdetermined and tractable as a linear algebra problem.
-These solutions vary little over the observing period and exhibit less than 1\% rms variation.
-This method, as applied to this PAPER data set is discussed in depth by Zheng et al (in prep).
+Because the array samples correlations redundantly, the relative calibration between antennae is numerically overdetermined and tractable as a linear algebra problem \cite{Liu:2010p10391}.
+As described in P14, we compute the ratio between redundant baselines, fit for a gain and phase slope and then algebraically solve for a per-antenna solution.
+Here we have avoided calibrating each channel independently to preserve as much frequency variation as possible.
+These solutions vary little over the observing period, exhibiting less than 1\% rms variation.
+A single solution derived for for the Dec 7 data set is used for the entire observing run.
+Time and frequency variation of redundant solutions is explored in detail in Zheng et al (in prep).
 
+Relative calibration is all that is necessary to form a power spectrum, however to set the flux scale, we must form a beam on a bright known point source.
 By itself, the redundant solution is completely independent of any sky model.
 The redundant solution contains two free delay parameters and an overall amplitude scale.
-We fit the two delay parameters to a model of Pictor A, Fornax A, and the Crab Nebula during a time when the sky is dominated by these three sources while marginalizing over the unknown ratio between the three sources.
+We fit the two delay parameters to a model of Pictor A, Fornax A, and the Crab Nebula during a time when the sky is dominated by these three sources while marginalizing over the unknown apparent flux ratio between the three sources\footnote{Unknown in the sense of a joint uncertainty in source flux and primary beam pattern.}.
 With the delays in place we are now able form a beam on Pictor A and (for each channel) set the overall amplitude to the calibration value found in \cite{jacobs:2013b}.
 
+ \begin{figure}
+\centering
+\includegraphics[width=\textwidth]{figures/psa32_flagging_zbins.png}
+\caption{\label{fig:bins} The average amount of data remaining after interference flagging over the 3 month period between Dec 2011 and March 2012 (black line) is quite high.
+Redshift bins (in grey, redshift center indicated with text label) are chosen to include spectral channels with uniform weight, i.e.
+no missing channels while maximizing coverage over the band.
+Redshift 7.7 is included for comparison with P14.
+Channels with no missing data are indicated by the dotted line.
+Each redshift bin is 20 MHz wide, but weighted by a Blackman-Harris window function which heavily down-weights the outer 10MHz for a Noise Equivalent Bandwidth of 10MHz.
+The interference is almost exclusively dominated by two features: ORBCOMM satellites at 137MHz and an unidentified intermittent line emitter at 175MHz.
+The roll off at 115MHz is due the rising noise at low frequencies being incorrectly flagged as interference.
+}
+\end{figure} 
 %The success of the remainder of the processing stages hinges on the spectral stability of the data, which can easily be assessed by comparing redundant baselines.
@@
 \centering
-\includegraphics[width=\textwidth]{figures/trms_by_baseline.png}
-\caption{\label{fig:noise} The residual noise after filtering foregrounds is consistent with the theoretical noise (dotted line).
-Here we estimate residual noise by differencing spectral channels and redundant baselines.
-The noise spectrum is characterized by an overall slope towards lower frequencies owing to the power law slope of the galactic spectrum which drives the system temperature.
-At both high and low edges the noise is also driven up by the use of a hamming window in the foreground filter.
-Less is known about foregrounds at the edges of the band which drives up the residuals.}
+\includegraphics[width=\textwidth]{figures/psa32_trms_with_and_without_fr_filtering_z.png}
+\caption{\label{fig:noise} Noise estimated by differencing adjacent frequencies (magenta) and between redundant baselines (blue), compared with an estimate of the theoretical noise level (dotted) indicating the rising significance of foregrounds as more data are added and sensitivity increases.
+The top three lines show noise after filtering foregrounds and binning into 40s long sidereal bins.
+At this noise level the frequency and baseline differences are roughly similar, both demonstrate the same small bumps of increased noise due to interference flagging.
+The bottom three lines show the noise level after integrating up to the maximum fringe rate of 914s.
+Here we see the interference bumps suppressed as more samples with different flagging are averaged together.
+We also see the difference between baselines become consistently larger than the difference between channels.
+In all cases the noise spectrum is characterized by an overall slope towards lower frequencies owing to the power law slope of the galactic spectrum which drives the system temperature.
+At the band edges noise rises dramatically, more so in the difference between baselines, then the spectral difference.
+This is, in part, due to the use of the Blackman-Harris window in the foreground filter is optimized for filtering foregrounds at the center of the band with minimal contamination of higher delay modes.
+We also note that the theoretical noise is consistently lower than the observed.
+This is most likely due to inaccuracies in the counting of data points included in the 92-day average, a key input into the theoretical noise limit.
+For this reason we estimate all error bars with a bootstrap method rather than by direct propagation.
+Correct propagation of error bars is a subject for future improvement.}
 \end{figure}
@@
 This removes excess correlation due to cross-talk in the analog signal chain.
-The residuals are then flagged once more for RFI before the XXX nights of data are averaged into 36.4 second long local sidereal time (LST) bins.
-During averaging we found that some lst bins were dominated by very bright data points due to some exceedingly bright samples.
-As these were very rare we found that removing the ten brightest points from every bin, a small median filter, was quite effective.
+The residuals are then flagged once more for RFI before the 92 nights of data are averaged into 36.4 second long local sidereal time (LST) bins.
+During averaging we found that some LST bins were dominated by very bright data points due to some exceedingly bright outlier samples.
+As these were very rare we found that removing the 10\% brightest from every bin, a small median filter, was quite effective.
+The source of these outliers is not known,  a likely possibility is instability in the analog signal chain stimulated by weather or bright interference which have been observed in subsequent seasons.
 The root mean square of the residual signal (seen in Figure \ref{fig:noise}) at the end of this process is close to the 1mK level expected given the total integration time.
+
+Though the frequency and repeated nightly observations have been averaged to their maximum extent, at 40s integrations the time axis has yet to be fully exploited.
+Sky-like fringes on a 30m baseline rotate much slower than 40 seconds.
+Performing a final fringe-rate filter, this time limiting to fringe-rates expected on a 30m baseline (down from 300m in the last iteration) we arrive at a data-set averaged to 900s, the maximum possible while still maintaining coherency.
 
@@
 The output of the forgoing is a single calibrated and foreground filtered sidereal night.
-The power spectrum is estimated by selecting a 10MHz bandwidth range centered on the redshift of interest, performing a delay transform and cross correlating all the bins.
-To preserve the isolation of any foregrounds which remain, we increase the spectral range by 5MHz on each side and multiply by a blackman-harris window thus providing a much higher dynamic range delay spectrum psf.
+The power spectrum is estimated in the delay spectrum of  a 10MHz bandwidth range centered on the redshift of interest.
+To preserve the isolation of any foregrounds which remain, we increase the spectral range by 5MHz on each side and multiply by a Blackman-Harris window thus providing a much higher dynamic range delay spectrum point spread function.
+
 This leaves us with 40 delay samples on each of 70 baselines which are divided into three redundant groups.
@@
 The cross correlation of the same delay modes between different redundant baselines provides our estimate of the power spectrum.
-Cross correlation between different modes provides our measure of systematic bias.
-These correlations should be identical between all redundant baselines to within the level of the noise.
-In practice, however there are some correlations deviate significantly from the median.
-These are removed by subtracting an average sky and then projecting out large residual modes.
-For more see Appendix of C of \cite{Parsons:2013p9876}.
+Cross correlation between different delay modes provides our measure of systematic bias.
+Sky-like correlations should be identical between all redundant baselines to within the level of the noise.
+In practice some correlations deviate significantly from the median.
+These are removed by iteratively dividing the covariance into a model of systematics and a model of sky-like emission and then projecting out large residual modes.
+This is done by dividing the baselines into different groups such that all cross-multiplications are done without introducing noise bias.
+For more see Appendix of C of P14.
 
 The residual correlations are all samples of the power spectrum.
-To estimate the final power spectrum and its uncertainty we compute the mean and variance of many random bootstrap-selected subsamples, sampling the dimensions: sidereal time, redundant baseline pair, and delay sign\footnote{As visibilities are complex, both the positive and negative delays  carry different information.}
+To estimate the final power spectrum and its uncertainty we compute the mean and variance of many random randomly-selected subsamples, sampling the dimensions: sidereal time, redundant baseline pair, and delay sign\footnote{As visibilities are complex, both the positive and negative delays  carry  information.}
 
@@
 \centering
-\includegraphics[width=0.45\textwidth]{figures/pspec_Mar24_psa32_zbins_1_80_119_I.png}
-\includegraphics[width=0.45\textwidth]{figures/pspec_Mar24_psa32_zbins_1_100_139_I.png}
-\includegraphics[width=0.45\textwidth]{figures/pspec_Mar24_psa32_zbins_1_110_149_I.png}
-\includegraphics[width=0.45\textwidth]{figures/pspec_Mar24_psa32_zbins_1_119_150_I.png}
-\caption{\label{fig:pspecs} Right to left from top: power spectra centered on redshifts 8.5, 7.9, 7.6 and 7.5 (frequencies: 149.5, 159.5 164.5, and 167 MHz) at a bandwidth of 20MHz}
+\includegraphics[width=0.45\textwidth]{{figures/pspec_log_z_10.29}.png}
+\includegraphics[width=0.45\textwidth]{{figures/pspec_log_z_8.54}.png}
+\includegraphics[width=0.45\textwidth]{{figures/pspec_log_z_7.94}.png}
+\includegraphics[width=0.45\textwidth]{{figures/pspec_log_z_7.68}.png}
+\includegraphics[width=0.45\textwidth]{{figures/pspec_log_z_7.55}.png}
+\caption{\label{fig:pspecs} Left to right from top: power spectra centered on redshifts 10.3,8.5, 7.9, 7.7 and 7.6 (frequencies: 127, 149.5, 159.5 164.5, and 167 MHz) at an effective bandwidth of 10MHz.
+Though the redshift 7.7 bin overlaps adjacent points it has provided for comparison against P14 which targeted this redshift with this data and reduction pipeline.}
 \end{figure}
@@
 
+Discussion points:
+\begin{enumerate}
+\item as a check, the redshift 7.7 is very similar 
+\item power spectrum at 10.3 is completely foreground contaminated. this appears to be consistent with the noise curves where the residual is much higher.
+\item Most of the data points lie above the theoretical noise curves. this because the noise model is a simple root(2Nbt) estimate.
+\item we could get a redshift bin at 7 if we had a prescription for including non-uniform weights.  
+\end{enumerate}
+
 \begin{figure}
 \centering
-\includegraphics[width=0.48\textwidth]{figures/psa_p3k_vs_z_28.png}
-\includegraphics[width=0.48\textwidth]{figures/psa_p3k_vs_z_30.png}
-\includegraphics[width=0.48\textwidth]{figures/psa_p3k_vs_z_33.png}
-\includegraphics[width=0.48\textwidth]{figures/psa_p3k_vs_z_38.png}
+\includegraphics[width=0.48\textwidth]{{figures/psa32_pspec_k_0.10_lin}.png}
+\includegraphics[width=0.48\textwidth]{{figures/psa32_pspec_k_0.20_lin}.png}
+\includegraphics[width=0.48\textwidth]{{figures/psa32_pspec_k_0.30_lin}.png}
+\includegraphics[width=0.48\textwidth]{{figures/psa32_pspec_k_0.40_lin}.png}
 \caption{Power spectrum amplitude vs redshift at a selection of k modes.
-Right to left from top, k=0.28,0.3,0.33,0.38 hMpc$^-1$.
-Paciga 2013 GMRT marked with an 'x', Parsons 2013 PAPER limit marked with thin black, this work marked with thick blue diamonds.
+Right to left from top, k=0.1,0.2,0.3,0.4 hMpc$^-1$.
+Paciga 2013 GMRT marked with an 'x', Parsons 2014 PAPER limit marked with thin black, this work marked with thick blue diamonds.
 \label{fig:slices}}
@@
 \label{sec:conclusion}
+
 Here we have, using the same dataset extended the expanded the redshift coverage of \cite{Parsons:2013p9876} cover the range $10>z>7$.
 These measurements demonstrate a precision removal of foregrounds to one part in ten million across the entire redshift range of interest.
-Though, with only 32 antenna, this data lacks the sensitivity to exclude all but the most extreme models, it does demonstrate the ability of the wideband filtration method to affectively remove foregrounds to the precision needed to integrate XXX nights to the thermal noise limit.
+Though, with only 32 antenna, this data lacks the sensitivity to exclude all but the most extreme models, it does demonstrate the ability of the wideband filtration method to affectively remove foregrounds to the precision needed to integrate 92 nights (1809 hours) to the thermal noise limit.
 As future observations with PAPER  will add sensitivity primarily by increasing to 128 antennae this projects well for our ability to reach design sensitivity.

### 2014-04-28 ms.tex
adding some writing in the discussion

@@
 At this noise level the frequency and baseline differences are roughly similar, both demonstrate the same small bumps of increased noise due to interference flagging.
-The bottom three lines show the noise level after integrating up to the maximum fringe rate of 914s.
+The bottom three lines show the noise level after integrating up to the maximum fringe rate of 776s.
 Here we see the interference bumps suppressed as more samples with different flagging are averaged together.
@@
 %XXX more, and a plot of those redundancy residuals
-\subsection{Foregrounds}
+\subsection{Foreground Filtering and Night Averaging}
 Foregrounds are filtered from the calibrated data by removing all bright delay components with light travel times less then the baseline length.
@@
 As these were very rare we found that removing the 10\% brightest from every bin, a small median filter, was quite effective.
-The source of these outliers is not known,  a likely possibility is instability in the analog signal chain stimulated by weather or bright interference which have been observed in subsequent seasons.
-The root mean square of the residual signal (seen in Figure \ref{fig:noise}) at the end of this process is close to the 1mK level expected given the total integration time.
+The source of these outliers is not known,  a likely possibility is instability in the analog signal chain stimulated by weather or bright interference, a circumstance that has since been observed in later seasons.
 
@@
 Performing a final fringe-rate filter, this time limiting to fringe-rates expected on a 30m baseline (down from 300m in the last iteration) we arrive at a data-set averaged to 900s, the maximum possible while still maintaining coherency.
+The root mean square of the residual signal (seen in Figure \ref{fig:noise}) at the end of this process is close to the 3mK level expected given the total integration time.
 
@@
 \subsection{Power Spectrum}
+\label{sec:power_spectrum}
 The output of the forgoing is a single calibrated and foreground filtered sidereal night.
@@
 \label{sec:results}
-In  Figure \ref{fig:pspecs} we see power spectra at different redshift bins, and in Figure \ref{fig:slices} we see power spectra at different $k$ modes as a function of redshift.
+\subsection{Foreground Filtering and Noise Levels}
+\label{sec:noise}
+The root-mean-square of the residual measures how well we've removed foregrounds.
+In Figure \ref{fig:noise} we examine $T_{rms}$ as calculated by differencing between adjacent channels and between redundant baselines.
+We see that in the last step of coherent integration, the step between 40 second to 789 second integrations, the noise level decreases by the expected factor of $\sqrt{789/40}$.
+At both stages the noise level found by both methods is roughly consistent with the theoretical level, only deviating significantly at the edges where the effective signal to noise due to the Blackman Harris window drops precipitously.
 
-Discussion points:
+The theoretical noise level is a straightforward calculation, but given the method employed here, should be given due description.
+Not every LST bin is observed an equal number of times nor is the noise level constant, as the galaxy moves overhead.
+the number of points in each bin is mostly determined by the observing schedule, which for PAPER is simply sunrise to sunset, but also by any flagging of data or observing outage (only 92 of the 103 days in the observing period have recorded data).
+Unfortunately, technical problems limited the retention of the exact cumulative count of points averaged into each LST bin during the averaging process.
+Here we have chosen to use the first order estimate based on the observing schedule to estimate the number of observations in each LST.
+Some bins are observed only a few times, while a slim range in the LST range 6 - 9 hours is observed on every night.
+On average each bin is measured 64 times.
+Given this accounting of the number of samples $N(lst)$, we estimate the final rms temperature to be
+
+\[
+T_{rms}  = \frac{T_{g}(lst)}{\sqrt{2BtN(lst)}} + T_{rcvr}
+\]
+where B and t are the 496kHz channel width and lst bin length, respectively.
+We model galactic temperature $T_g$ by summing the de Oliviera-Costa global sky model \citep{deOliveiraCosta:2008p2242} convolved by the PAPER beam as a function of lst and $T_{rcvr}$=320K as the system temperature of 550K (found in P14) minus the mean galactic temperature from that time period.
+The final result is plotted as a dotted line in Figure \ref{fig:noise} for before and after fringe rate filtering (t=40s and 788s)%XXX TODO fold this back into figure 1 so that it shows the correct noise estimate
+
+The final fringe-rate step is the last coherent average of the processing routine and represents the last large gain in sensitivity; all following averaging steps are done incoherently on the square of the visibilities where sensitivity is gained at a much slower rate.
+For this reason the noise curves (bottom three curves of Figure \ref{fig:noise})  merit closer inspection.
+At this sensitivity level we see a notable excess of power in the rms difference between baselines.
+Though the cause of this excess is unknown, it is suggestive of the excess covariance between redundant baselines which we remove in Section \ref{sec:power_spectrum} a hypothesis also supported by the high levels of residual foregrounds in the redshift 10.5 bin where the baseline difference rms is highest.
+
+\subsection{Power Spectra}
+
+In  Figure \ref{fig:pspecs} we show the  power spectra at different redshift bins --showing both the spherically averaged power spectrum plotted in $P_k$ units (left) and the final estimate of $\Delta^2$ -- and in Figure \ref{fig:slices} we see power spectra at different $k$ modes as a function of redshift  all plotted with $2\sigma$ error bars.
+In the redshift 7.7 bin the foreground measurements (the inner 5 data points) have been included for comparison with P14.
+The theoretical noise power spectrum is estimated by counting data points using the method of \cite{Pober:2014p10390}.
+At the lowest redshift (7.6) most of the power spectrum points are consistent with zero and error bars broadly reflective of the noise model.
+The number of noise-like bins decreases towards higher redshifts, until the redshift 10.5 where all bins appear to be dominated by systematics rather than noise.
+In all but the redshift 10.5 slice, the $P_k$ spectrum has a clear notch near the $k_\parallel=0$ where the broadband foreground filter has removed power.
+At redshift 10.5, the broadband filter amplitude is down 75\% from its peak at z=8.5
+
+Limitations: 
 \begin{enumerate}
-\item as a check, the redshift 7.7 is very similar 
-\item power spectrum at 10.3 is completely foreground contaminated. this appears to be consistent with the noise curves where the residual is much higher.
-\item Most of the data points lie above the theoretical noise curves. this because the noise model is a simple root(2Nbt) estimate.
-\item we could get a redshift bin at 7 if we had a prescription for including non-uniform weights.  
+\item time dependence of cal solutions
+\item missing data
 \end{enumerate}
@@
 Paciga 2013 GMRT marked with an 'x', Parsons 2014 PAPER limit marked with thin black, this work marked with thick blue diamonds.
-\label{fig:slices}}
+Points consistent with  \label{fig:slices}}
 \end{figure}
@@
 
-Here we have, using the same dataset extended the expanded the redshift coverage of \cite{Parsons:2013p9876} cover the range $10>z>7$.
-These measurements demonstrate a precision removal of foregrounds to one part in ten million across the entire redshift range of interest.
-Though, with only 32 antenna, this data lacks the sensitivity to exclude all but the most extreme models, it does demonstrate the ability of the wideband filtration method to affectively remove foregrounds to the precision needed to integrate 92 nights (1809 hours) to the thermal noise limit.
+Here we have extended the redshift coverage of \cite{Parsons:2013p9876} to cover the range $10.5>z>7.5$.
+These measurements demonstrate a precision removal of foregrounds to one part in ten million across a wide redshift.
+With only 32 antenna, this data lacks the sensitivity to exclude all but the most extreme models, however, it does demonstrate the ability of the wideband filtration method to affectively remove foregrounds to the precision needed to integrate a full season to the thermal noise limit.
 As future observations with PAPER  will add sensitivity primarily by increasing to 128 antennae this projects well for our ability to reach design sensitivity.

### 2014-04-29 ms.tex
updating figure captions and some new text in the discussion

@@
 \includegraphics[width=\textwidth]{figures/psa32_trms_with_and_without_fr_filtering_z.png}
-\caption{\label{fig:noise} Noise estimated by differencing adjacent frequencies (magenta) and between redundant baselines (blue), compared with an estimate of the theoretical noise level (dotted) indicating the rising significance of foregrounds as more data are added and sensitivity increases.
+\caption{\label{fig:noise} Root mean square (rms) noise before and after integrating to 10 minute LST bins (top and bottom sets of curves) indicating the rising significance of foregrounds as we take the last deep integration step.
+Noise is estimated by differencing adjacent frequencies (magenta) and between redundant baselines (blue), compared with an estimate of the theoretical noise level (dotted).
 The top three lines show noise after filtering foregrounds and binning into 40s long sidereal bins.
-At this noise level the frequency and baseline differences are roughly similar, both demonstrate the same small bumps of increased noise due to interference flagging.
+At this noise level the frequency and baseline differences are roughly similar, both demonstrate the same small bumps of increased noise due to interference flagging and are consistent with noise over much of the band.
 The bottom three lines show the noise level after integrating up to the maximum fringe rate of 776s.
-Here we see the interference bumps suppressed as more samples with different flagging are averaged together.
-We also see the difference between baselines become consistently larger than the difference between channels.
-In all cases the noise spectrum is characterized by an overall slope towards lower frequencies owing to the power law slope of the galactic spectrum which drives the system temperature.
-At the band edges noise rises dramatically, more so in the difference between baselines, then the spectral difference.
-This is, in part, due to the use of the Blackman-Harris window in the foreground filter is optimized for filtering foregrounds at the center of the band with minimal contamination of higher delay modes.
-We also note that the theoretical noise is consistently lower than the observed.
-This is most likely due to inaccuracies in the counting of data points included in the 92-day average, a key input into the theoretical noise limit.
-For this reason we estimate all error bars with a bootstrap method rather than by direct propagation.
-Correct propagation of error bars is a subject for future improvement.}
+The step between the top and bottom sets is the last coherent integration with noise largely decreasing by the expected factor $\sqrt{776/40}$ except in the difference between baselines (blue) which demonstrates a clear excess at all frequencies, particularly above redshift 10.
+With this last large jump in sensitivity we are now seeing the slight dominance of baseline covariance over the spectral derivative rms.
+}
 \end{figure}
@@
 \includegraphics[width=0.45\textwidth]{{figures/pspec_log_z_7.55}.png}
-\caption{\label{fig:pspecs} Left to right from top: power spectra centered on redshifts 10.3,8.5, 7.9, 7.7 and 7.6 (frequencies: 127, 149.5, 159.5 164.5, and 167 MHz) at an effective bandwidth of 10MHz.
-Though the redshift 7.7 bin overlaps adjacent points it has provided for comparison against P14 which targeted this redshift with this data and reduction pipeline.}
+\caption{\label{fig:pspecs} Left to right from top: power spectra centered on redshifts 10.3,8.5, 7.9, 7.7 and 7.6 (frequencies: 127, 149.5, 159.5 164.5, and 167 MHz) at an effective bandwidth of 10MHz or redshift range 0.5  }
 \end{figure}
@@
 
-In  Figure \ref{fig:pspecs} we show the  power spectra at different redshift bins --showing both the spherically averaged power spectrum plotted in $P_k$ units (left) and the final estimate of $\Delta^2$ -- and in Figure \ref{fig:slices} we see power spectra at different $k$ modes as a function of redshift  all plotted with $2\sigma$ error bars.
-In the redshift 7.7 bin the foreground measurements (the inner 5 data points) have been included for comparison with P14.
-The theoretical noise power spectrum is estimated by counting data points using the method of \cite{Pober:2014p10390}.
-At the lowest redshift (7.6) most of the power spectrum points are consistent with zero and error bars broadly reflective of the noise model.
-The number of noise-like bins decreases towards higher redshifts, until the redshift 10.5 where all bins appear to be dominated by systematics rather than noise.
-In all but the redshift 10.5 slice, the $P_k$ spectrum has a clear notch near the $k_\parallel=0$ where the broadband foreground filter has removed power.
-At redshift 10.5, the broadband filter amplitude is down 75\% from its peak at z=8.5
-
-Limitations: 
-\begin{enumerate}
-\item time dependence of cal solutions
-\item missing data
-\end{enumerate}
+In  Figure \ref{fig:pspecs} we show the  power spectra at different redshift bins --showing both the spherically averaged power spectrum plotted in $P_k$ units (left) and the final estimate of $\Delta^2$ -- and in Figure \ref{fig:slices} we see different $k$ mode slices as a function of redshift  all plotted with $2\sigma$ error bars.
+The theoretical noise power spectrum is estimated using the method described in \cite{Pober:2014p10390} assuming a system temperature of 550K and the observing scheme described in Section \ref{sec:observations}.
+Generally, both figures convey a picture of a power spectrum dominated by noise, with foregrounds encroaching from modes closest to the foreground contaminated modes.
+Foreground dominance appears to grow with increasing redshift, mirroring the rise in baseline-to-baseline rms we found in Figure \ref{fig:noise}.
+Recall that the wide-band foreground model we have subtracted is built by weighting the entire band by a Blackman-Harris window function.
+Foreground removal is much weaker at this frequency, with the model subtracted being centered around 150MHz, contributions from this point in the band are down-weighted by 75\%.
+This filter leads to the notch of missing power near $k_\parallel=0$  (except at z=7.68 where the foregrounds have been left in for comparison).
+Only at redshift 10.3 is this notch nearly non-existent.
+The relative distribution of foreground removal accuracy is also visible in the shape of $\Delta^2$ vs redshift at $k=0.1$ which is only a factor of two away from the light travel horizon.
+This bin, which is firmly well into the foreground dominated portion of $k$ space has much the same shape as the $T_rms$ curve in Figure \ref{fig:noise}, with a minimum near redshift 8.5 and a dramatic rise above redshift 10.
 
@@
 \caption{Power spectrum amplitude vs redshift at a selection of k modes.
-Right to left from top, k=0.1,0.2,0.3,0.4 hMpc$^-1$.
+Left to right from top, k=0.1,0.2,0.3,0.4 hMpc$^{-1}$.
 Paciga 2013 GMRT marked with an 'x', Parsons 2014 PAPER limit marked with thin black, this work marked with thick blue diamonds.
-Points consistent with  \label{fig:slices}}
+The k=0.1 hMpc$^{-1}$ bin (top-left), which samples the delay spectrum at only 2x the maximum horizon delay  is foreground dominated with  a redshift dependence similar to the  $T_{rms}$ residual in Figure \ref{fig:noise}.
+The rest are consistent with thermal noise.
+\label{fig:slices}}
 \end{figure}
@@
 
-Here we have extended the redshift coverage of \cite{Parsons:2013p9876} to cover the range $10.5>z>7.5$.
-These measurements demonstrate a precision removal of foregrounds to one part in ten million across a wide redshift.
+With this paper we have extended the redshift coverage of \cite{Parsons:2013p9876} to cover the range $10.5>z>7.6$.
+The amount of foreground contamination slowly increases with redshift, paralleling the level of incoherent noise observed  This highlights the limitation of a the simple Blackman-Harris weighting scheme.
+Namely that the model of
+
 With only 32 antenna, this data lacks the sensitivity to exclude all but the most extreme models, however, it does demonstrate the ability of the wideband filtration method to affectively remove foregrounds to the precision needed to integrate a full season to the thermal noise limit.

### 2014-04-30 ms.tex
fixed an unfinished sentence or two in the discussion and fleshed out the conclusions a little bit

@@
 \includegraphics[width=0.45\textwidth]{{figures/pspec_log_z_7.55}.png}
-\caption{\label{fig:pspecs} Left to right from top: power spectra centered on redshifts 10.3,8.5, 7.9, 7.7 and 7.6 (frequencies: 127, 149.5, 159.5 164.5, and 167 MHz) at an effective bandwidth of 10MHz or redshift range 0.5  }
+\caption{\label{fig:pspecs} Left to right from top: power spectra centered on redshifts 10.3,8.5, 7.9, 7.7 and 7.6 (frequencies: 127, 149.5, 159.5 164.5, and 167 MHz.
+All with an effective bandwidth of $\pm$10MHz which covers the redshift span $\Delta z\approx$1421MHz ($B/f^2$) which ranges from 0.8 at $z=$10.3 to 0.45 at $z=$7.4.
+}
 \end{figure}
@@
 
+%Across the entire band some $k$ modes appear to be dominated somewhat by non-noise systematics, these are not always clearly foregrounds because they appear at only certain $k$ modes rather than clearly emanating from small delays like in the z=10.3 power spectrum.
+
 \begin{figure}
@@
 With this paper we have extended the redshift coverage of \cite{Parsons:2013p9876} to cover the range $10.5>z>7.6$.
-The amount of foreground contamination slowly increases with redshift, paralleling the level of incoherent noise observed  This highlights the limitation of a the simple Blackman-Harris weighting scheme.
-Namely that the model of
+The amount of foreground contamination slowly increases with redshift, paralleling the level of incoherent noise observed  This highlights the limitation of the Blackman-Harris weighting scheme used to estimate the broadband foreground model which trades spectral coverage for isolation between delay modes.
+Accessing these regions nearer to the edge of the band, where there is less spectral information on which to base a foreground model will require some combination of methods for removing foregrounds that make maximal use of the full band without as much down-weighting, as well as a more precise method for removing foregrounds in the narrow band-power spectrum.
 
@@
 As future observations with PAPER  will add sensitivity primarily by increasing to 128 antennae this projects well for our ability to reach design sensitivity.
+The residual foregrounds analyzed here are all at or near the level of noise and expose the next level of removal required to access a wider redshift range.
 

### 2014-05-01 ms.tex
updating figures in pdf

@@
 \includegraphics[width=0.45\textwidth]{{figures/pspec_log_z_7.55}.png}
-\caption{\label{fig:pspecs} Left to right from top: power spectra centered on redshifts 10.3,8.5, 7.9, 7.7 and 7.6 (frequencies: 127, 149.5, 159.5 164.5, and 167 MHz.
-All with an effective bandwidth of $\pm$10MHz which covers the redshift span $\Delta z\approx$1421MHz ($B/f^2$) which ranges from 0.8 at $z=$10.3 to 0.45 at $z=$7.4.
+\caption{\label{fig:pspecs} Left to right from top: power spectra centered on redshifts 10.3,8.5, 7.9, 7.7 and 7.6 (frequencies: 127, 149.5, 159.5 164.5, and 167 MHz.)  All with an effective bandwidth of $\pm$10MHz  covering the redshift span $\Delta z\approx$1421MHz ($B/f^2$) which ranges from 0.8 at $z=$10.3 to 0.45 at $z=$7.4.
+The noise curve (blue dashes) calculated using the method described in \cite{Pober:2014p10390}  indicates the 1$\sigma$ confidence bounds on data points consisting purely of noise; 65\% of uncorrelated noise like data points will lie below the curve.
+Note that approximately every other $k$ bin is uncorrelated.
 }

### 2014-05-05 ms.tex
change y-axis on pspec so we can see early results. punching up discussion text

@@
 
-\title{The PAPER-32 power spectrum measured at several redshifts}
+\title{Multi-redshift limits on the 21cm power spectrum from PAPER}
 \author{
@@
 Pat Klima\altaffilmark{5},
-Adrian Liu\altaffilmark{2},
+Adrian Liu\altaffilmark{2,11},
 Dave H. E. MacMahon\altaffilmark{8}
@@
 \altaffiltext{10}{Cavendish Lab., Cambridge, UK}
+\altaffiltext{11}{Berkeley Center for Cosmological Physics, UC Berkeley, Berkeley,
+CA}
 
@@
 We present new observations from the Donald C.
-Backer Precision Array for Probing the Epoch of Reionization telescope which probe the redshift range of $6<z<9$, extending previously published single redshift results to cover the full range accessible to the instrument.
-The removal of foregrounds depends heavily on the spectral smoothness of both instrumental response and foregrounds.
-These spectra demonstrate a robust foreground removal of foreground signals to the thermal noise limit, leaving systematics on the scale of the noise.
-We find that these residual systematics are present throughout the observed frequency band are most likely some mixture residual foregrounds and instrumental systematics.
+Backer Precision Array for Probing the Epoch of Reionization (PAPER) telescope which place new lower limits on the HI power spectrum over the redshift range of $7.5<z<10.5$, extending previously published single redshift results to cover the full range accessible to the instrument.
+The epoch of reionization power spectrum is expected to evolve strongly with redshift, and it is this variation with cosmic history that will allow us to begin to place constraints on the physics of reionization.
+The primary obstacle to this goal are bright foregrounds and the attendant systematics associated with high dynamic range measurements.
+Here we use foreground removal techniques that take advantage of the large instrumental bandwidth to make a high dynamic range estimate of smooth foreground power.
+Power spectra at different points across the spectrum reveal the variable efficacy of this subtraction.
+Noise limited measurements of $\Delta^2$ at $k=$0.2hMpc$^{-1}$ reach as low as 29mK at redshift 7.5.
+Most spectra demonstrate a robust foreground removal of foreground signals to the thermal noise limit, leaving systematics on the scale of the noise.
+We find that in the case of the highest redshift power spectrum, foregrounds begin to dominate over the noise, reflecting the increased difficulty in filtering foregrounds near the edge of the instrumental band.
 \end{abstract}
@@
 \section{Introduction}
-The Epoch of Reionization, when the first stars ionized the pervasive cosmological Hydrogen in the last global phase change, is predicted to be observable in highly redshifted 21 cm radiation.
+The Epoch of Reionization, when the first stars ionized the pervasive cosmological Hydrogen in the last global phase change is predicted to be observable in highly redshifted 21 cm radiation.
 The Donald C.
 Backer Precision Array for Probing the Epoch of Reionization (PAPER, XXX)\footnote{\url{http:\\eor.berkeley.edu}} is a low frequency radio interferometer experiment dedicated to opening this window on the universe.
-Challenges include foregrounds which are brighter by several orders of magnitude and limited collecting areas of first generation instruments which necessitate long integration times.
+Challenges include foregrounds which are brighter by several orders of magnitude and long integration times necessitated by the limited collecting areas of first generation instruments.
 Direct observation of hydrogen before and during re-ionization is predicted to deliver a wealth of cosmological and astrophysical data, including the nature of the first stellar objects and the timing and rate of galaxy formation, reviews on the physics of reionization as well as theory about the nature of foregrounds maybe be found in \citet{Furlanetto:2006p2267,Morales:2010p8093,Pritchard:2012p9555}.
@@
 Here we report on deep integrations made with a 32 element array in 2011, first described in \cite{Parsons:2013p9876}, hereafter P14.
-In that work we described in detail out reduction method which enabled a deep limit on the power spectrum in the presence of bright foregrounds.
-Using this method we combined 92 nights of data to place upper limits on the power spectrum of HI in the early universe.
-However, this analysis was limited to a single redshift.
+In that work we described in detail our reduction method which enabled a deep limit on the power spectrum in the presence of bright foregrounds.
+Using this method we combined 92 nights (1800 hours) of data to place upper limits on the power spectrum of HI in the early universe.
+However, this analysis was limited to a single redshift ($z=$7.7).
 The epoch of reionization signal is expected to evolve strongly with redshift, and it is this variation with cosmic history that will allow us to begin to place constraints on the physics of reionization \cite{Pritchard:2008p8123,Pober:2014p10390}.
 Therefore, while a detection of the 21 cm signal at even a single frequency would be a tremendous breakthrough, analysis techniques must be developed to capitalize on the wide bandwidths of the current generations of high redshift 21cm telescopes.
-The primary obstacle to detection, the wide bandwidth Using the same data as P14, this paper presents new power spectrum upper limits covering the redshift range $10.5>z>7.2$.
-In section  \ref{sec:obs_meth} we summarize the observation and reduction methodology, present the new upper limits in \ref{sec:results} in \ref{sec:conclusion} offer some conclusions and discussion of future work.
+Using the same data as P14, this paper presents new power spectrum upper limits covering the redshift range $10.5>z>7.2$.
+In section  \ref{sec:obs_meth} we summarize the observation and reduction methodology, we present the new upper limits in \ref{sec:results}, and in \ref{sec:conclusion} we offer conclusions and discussion of future work.
 
@@
 \label{sec:observations}
-As the observations and methods in use are unchanged from those described in P14, we will offer a quick summary and refer the reader to \cite{Parsons:2013p9876} for a more in depth discussion.
+The work here follows the same basic procedure and uses the same underlying data set as P14.
+Here we provide a quick summary and refer the reader to \cite{Parsons:2013p9876} for a more in depth discussion.
 A general overview of the PAPER system can be found in \cite{Parsons:2010p6757}, calibration of the primary beam in \cite{Pober:2012p8800}, and imaging results in \cite{Jacobs:2011p8438,Jacobs:2013p10014,Stefan:2013p9926}.
 Sensitivity analysis described in \cite{Parsons:2012p9028} revealed that for the low gain elements employed by PAPER, a highly redundant ``grid'' type arrangement offers a significant sensitivity boost.
-In an interferometer each baseline samples a different mode of the power spectrum.
+In most interferometers each baseline samples a different mode of the power spectrum; the ideal case for reconstructing images.
 In a grid, there are many samples of each of these cosmological modes which can be averaged to a high sensitivity before being combined with other different $k$ modes.
-The PAPER South Africa 32 antenna deployment (PSA32) was arranged in a 4x8 grid.
-Here we include only the three shortest types of spacings where the reionization power is expected to be brightest.
-This selection includes those between adjacent columns within at least one row of each other, a selection containing 70 30m baselines.
+The PAPER South Africa 32 antenna deployment (PSA32) was arranged in a 4x8 grid with a column spacing of 30m and a row spacing of 4m.
+In our analysis, as in P14, we include only the three shortest types of spacings where the reionization power is expected to be brightest.
+This selection includes those between adjacent columns within at least one row of each other, a selection containing 70 $\sim$30m-long baselines.
 
-Observations spanning the band between 100 to 200-MHz ($13.1>z>6.1$) were recorded at a resolution of 48kHz and 10.7s  beginning Dec 7, 2011 and ending March 19, 2012 for a total of 92 nights.
+Observations spanning the band between 100 to 200-MHz ($13.1>z>6.1$) were recorded at a resolution of 48kHz and 10.7s  beginning Dec 7, 2011 and ending March 19, 2012 for a total of 92 nights or 1800 hours.
 WIthin this set we included observations in the LST range 1h - 9h where the sky dominated system temperature is at a minimum.
+Note that this LST range is slightly shorter than in P14 which extended to LST of 12 hours.
+These last three hours were found to contribute minimally and be dominated by bright galactic foreground emission and so have been excluded here.
 
@@
 
-In several stages throughout the analysis process we take a 2D Fourier transform of the visibility spectra $V(t,\nu)$ into ``delay'/fringe rate'' space.
+In several stages throughout the analysis process we take a 2D Fourier transform of the visibility spectra $V(\nu,t)$ into ``delay/fringe rate'' space where delay is the Fourier dual to frequency and similarly fringe-rate for time.
 In this space, smooth spectrum sources are physically localized to delays shorter than the light travel time length of the baseline (e.g.
 sources at the horizon, in the direction of the baseline vector) and fringe rates shorter than the sidereal rate.
-In this Fourier space sources are highly localized, deviations from a flat spectrum manifesting as a slight dispersion.
-In this space the spectrum sampling function, which is uneven due to flagging of interference takes the form of a convolution by a point-spread-function, much in the same way an imperfect sampling of the uv plane gives rise to the psf of an interferometer.
-To account for this we use a CLEAN like, iterative, peak finder and subtraction algorithm which is limited to finding peaks within the physically allowable ranges, and occasionally beyond, where necessary to account for spectral and temporal dispersion.
+In this Fourier space, sources are highly localized with deviations from a flat spectrum manifesting as a slight dispersion.
+The spectrum sampling function, which is uneven due to flagging of interference takes the form of a convolution by a point-spread-function in the same way an imperfect sampling of the uv plane gives rise to the angular psf of an interferometer.
+To account for this, we use a CLEAN like, iterative, peak-finder and subtraction algorithm which is limited to finding peaks within the physically allowable ranges.
 
@@
 This final step takes advantage of the redundant baselines to make an unbiased power spectrum estimate by cross-multiplying identical baselines and then averaging the power spectrum modes.
+The details of each application of the delay/fringe-rate transform will be laid out the following sections as we provide a brief walk-through of the processing pipeline.
 
@@
 %I know this isn't technically true, but I think its important that the downselect is obvious up front.
-   The visibilities are then compressed in the frequency and time directions by filtering delay modes above the horizon and highest frequency fringe rates for a 300m baseline.
+   The visibilities are then compressed in the frequency and time directions by removing delay modes above the horizon and highest frequency fringe rates for a 300m baseline.
 This filtering is done in tandem with a radio frequency interference (RFI) flagging step, using the residuals which have had bright sky-like signals removed to flag 4$\sigma$ deviations before feeding the flags back into another iteration of the compression step.
 This mitigates the effects of bright, narrow band, interference  being scattered by the delay filter to generate a cleaner averaged data set and results in a time and frequency bin of length 39.6s and width 492.61kHz.
-  This process reduces the data volume by a factor of \~20, or roughly an order of magnitude improvement on traditional time and frequency averaging which in this array would be limited to ~100kHz and 10s to avoid averaging away fringes.
+  This process reduces the data volume by a factor of $\sim$20, or roughly an order of magnitude improvement on traditional time and frequency averaging which in this array would be limited to ~100kHz and 10s to avoid averaging away fringes.
 
@@
 Here we have avoided calibrating each channel independently to preserve as much frequency variation as possible.
-These solutions vary little over the observing period, exhibiting less than 1\% rms variation.
+These solutions vary little over the three month observing period, exhibiting less than 1\% rms variation.
 A single solution derived for for the Dec 7 data set is used for the entire observing run.
@@
 We fit the two delay parameters to a model of Pictor A, Fornax A, and the Crab Nebula during a time when the sky is dominated by these three sources while marginalizing over the unknown apparent flux ratio between the three sources\footnote{Unknown in the sense of a joint uncertainty in source flux and primary beam pattern.}.
-With the delays in place we are now able form a beam on Pictor A and (for each channel) set the overall amplitude to the calibration value found in \cite{jacobs:2013b}.
+With the delays in place we are now able form a beam on Pictor A and (for each channel) set the overall amplitude to the calibration value of 381Jy found in \cite{jacobs:2013b}.
 
@@
 \caption{\label{fig:pspecs} Left to right from top: power spectra centered on redshifts 10.3,8.5, 7.9, 7.7 and 7.6 (frequencies: 127, 149.5, 159.5 164.5, and 167 MHz.)  All with an effective bandwidth of $\pm$10MHz  covering the redshift span $\Delta z\approx$1421MHz ($B/f^2$) which ranges from 0.8 at $z=$10.3 to 0.45 at $z=$7.4.
-The noise curve (blue dashes) calculated using the method described in \cite{Pober:2014p10390}  indicates the 1$\sigma$ confidence bounds on data points consisting purely of noise; 65\% of uncorrelated noise like data points will lie below the curve.
-Note that approximately every other $k$ bin is uncorrelated.
+The noise curve (blue dashes) is calculated using the method described in \cite{Pober:2014p10390}  and indicates the 1$\sigma$ confidence bounds on data points consisting purely of noise; 65\% of uncorrelated noise like data points will lie below the curve.
+Note that due to the weighting of the delay transform, adjacent $k$ bins are 50\% correlated.
+The black line is a fiducial model at 50\% ionization \citep{Lidz:2008p8251}.
+The redshift 7.68 bin has been reprocessed for comparison with the P14 result.
+Note that the LST range processed here is 3 hours less than P14 for a small reduction in sensitivity but large reduction in foreground.
 }
@@
 With this paper we have extended the redshift coverage of \cite{Parsons:2013p9876} to cover the range $10.5>z>7.6$.
-The amount of foreground contamination slowly increases with redshift, paralleling the level of incoherent noise observed  This highlights the limitation of the Blackman-Harris weighting scheme used to estimate the broadband foreground model which trades spectral coverage for isolation between delay modes.
-Accessing these regions nearer to the edge of the band, where there is less spectral information on which to base a foreground model will require some combination of methods for removing foregrounds that make maximal use of the full band without as much down-weighting, as well as a more precise method for removing foregrounds in the narrow band-power spectrum.
+The amount of foreground contamination slowly increases with redshift, paralleling the level of incoherent noise observed.
+This highlights a possible limitation of the scheme used to subtract the broadband foreground model which trades spectral coverage for isolation between delay modes.
+Accessing higher redshifts, where there is less spectral information on which to base a foreground model will require some combination of methods for removing foregrounds that make maximal use of the full band without as much down-weighting, as well as a more precise method for removing foregrounds in the narrow band-power spectrum.
 

### 2014-05-08 ms.tex
finishing second round rough draft edits

@@
 The Donald C.
-Backer Precision Array for Probing the Epoch of Reionization (PAPER, XXX)\footnote{\url{http:\\eor.berkeley.edu}} is a low frequency radio interferometer experiment dedicated to opening this window on the universe.
+Backer Precision Array for Probing the Epoch of Reionization (PAPER, \cite{Parsons:2010p6757})\footnote{\url{eor.berkeley.edu}} is a low frequency radio interferometer experiment dedicated to opening this window on the universe.
 Challenges include foregrounds which are brighter by several orders of magnitude and long integration times necessitated by the limited collecting areas of first generation instruments.
@@
 Telescopes seeking to measure this signal include the Giant Metre-wave Radio Telescope (GMRT; \cite{Paciga:2013p9943}), the Low Frequency Array (LOFAR; \cite{Yatawatta:2013p9699}) and the Murchison Widefield Array (MWA; \cite{Bowman:2013p9950} and \cite{Tingay:2013p9022}).
-PAPER is located in the Karoo desert at the Square Kilometer Array-South Africa cite.
+PAPER is located in the Karoo desert at the site of the South African portion of the future Square Kilometer Array.
 PAPER has doubled in size on a yearly basis since 2009, making observations with each array.
@@
 Here we report on deep integrations made with a 32 element array in 2011, first described in \cite{Parsons:2013p9876}, hereafter P14.
-In that work we described in detail our reduction method which enabled a deep limit on the power spectrum in the presence of bright foregrounds.
-Using this method we combined 92 nights (1800 hours) of data to place upper limits on the power spectrum of HI in the early universe.
-However, this analysis was limited to a single redshift ($z=$7.7).
+In P14 we described in detail our reduction method and gave deep new limits on the HI power spectrum in the presence of bright foregrounds at redshift 7.68 using 92 nights (1800 hours) of data.
 The epoch of reionization signal is expected to evolve strongly with redshift, and it is this variation with cosmic history that will allow us to begin to place constraints on the physics of reionization \cite{Pritchard:2008p8123,Pober:2014p10390}.
@@
 
-\section{Observations and reduction}
+\section{Observations}
 \label{sec:observations}
@@
 In our analysis, as in P14, we include only the three shortest types of spacings where the reionization power is expected to be brightest.
-This selection includes those between adjacent columns within at least one row of each other, a selection containing 70 $\sim$30m-long baselines.
+This selection includes those between adjacent columns and within at least one row of each other, a selection containing 70 $\sim$30m-long baselines.
 
-Observations spanning the band between 100 to 200-MHz ($13.1>z>6.1$) were recorded at a resolution of 48kHz and 10.7s  beginning Dec 7, 2011 and ending March 19, 2012 for a total of 92 nights or 1800 hours.
-WIthin this set we included observations in the LST range 1h - 9h where the sky dominated system temperature is at a minimum.
+Observations spanning the band between 100 to 200-MHz ($13.1>z>6.1$) were recorded at a resolution of 48kHz and 10.7s  beginning Dec 7, 2011 and ending March 19, 2012 (with some down-time for maintenance) after a total of 92 nights or 1800 hours.
+Within this set we included observations in the LST range 1h - 9h where the sky dominated system temperature is at a minimum.
 Note that this LST range is slightly shorter than in P14 which extended to LST of 12 hours.
-These last three hours were found to contribute minimally and be dominated by bright galactic foreground emission and so have been excluded here.
+These last three hours were found to contribute minimally to increasing sensitivity while being dominated by bright galactic foreground emission and so have been excluded here.
 
@@
 \label{sec:obs_meth}
-Here we summarize our data reduction steps, for detailed descriptions see P14.
+Here we summarize our data reduction steps, for detailed descriptions see section 3 of P14.
 
@@
 This final step takes advantage of the redundant baselines to make an unbiased power spectrum estimate by cross-multiplying identical baselines and then averaging the power spectrum modes.
+Here we are also using the approximation that over short ($\delta z < \sim 0.5$) the delay spectrum is a good approximation of a sample of the 3D power spectrum \citep{Parsons:2012p8896}.
+Thus as we step through the various iterations of delay spectrum we have the ability to isolate and subtract foregrounds in the native domain of the measured power spectrum.
+Note that isolation of foregrounds to a region below line-of-site $k_\parallel$s that increases with baseline length is also the much discussed "wedge" \cite{Liu:2014p10462,Liu:2014p10463,Pober:2013p9942,Morales:2012p8790,Vedantham:2012p10297,Datta:2010p8781}.
+As we only have a single length of baseline the wedge becomes a narrow slice indeed.
 The details of each application of the delay/fringe-rate transform will be laid out the following sections as we provide a brief walk-through of the processing pipeline.
+
+\subsection{Selection of Redshift Bins}
+The redshift bins for which we have computed power spectra --shown in Figure \ref{fig:bins}-- have been selected from the available bandwidth using two criteria: minimizing covariance between redshifts and avoiding missing data.
+We minimize covariance between adjacent redshift bins by limiting overlap of each bin to the spectrum at the outer half of the bin channel range which, as will be described in section \ref{sec:power_spectrum} is significantly down-weighted by the use of a Blackman-Harris window.
+The power spectrum method developed in P14 leverages the redundant baselines to estimate non-sky covariance and project out these contaminated modes.
+This inversion is particularly sensitive to missing data -the inverse of the covariance is not defined- so we also select only channels that have no missing data over the entire sidereal period (see the dotted line in \ref{fig:bins}).
+With these constraints we arrive at the redshifts 10.3, 8.5,7.9 and 7.5.
+For the purposes of comparison with P14 we also include the redshift 7.86 bin.
+
+The spectral width of the redshift bins is dictated by the competing needs of sensitivity and foreground reduction which both benefit from wider bandwidths and the likely rate of reionization evolution which is thought to be close to $\Delta z \approx 0.5$, which to a good approximation is given for bandwidth $B$ and frequency $f$ as $\Delta z=1421 B/f^2\approx 0.5$.
+To balance these two competing constraints we choose a single spectral bandwidth of 20Mhz, weighted by a Blackman-Harris window for an effective width of 10Mhz, or $\Delta z=0.5$ at $z=7.45$ ranging up to a $\Delta z = 0.86$ at redshift 10.3.
 
@@
    The visibilities are then compressed in the frequency and time directions by removing delay modes above the horizon and highest frequency fringe rates for a 300m baseline.
-This filtering is done in tandem with a radio frequency interference (RFI) flagging step, using the residuals which have had bright sky-like signals removed to flag 4$\sigma$ deviations before feeding the flags back into another iteration of the compression step.
-This mitigates the effects of bright, narrow band, interference  being scattered by the delay filter to generate a cleaner averaged data set and results in a time and frequency bin of length 39.6s and width 492.61kHz.
+This filtering\footnote{See Appendix A of P14} is done in tandem with a radio frequency interference (RFI) flagging step, using the residuals which have had bright sky-like signals removed to flag 4$\sigma$ deviations before feeding the flags back into another iteration of the compression step.
+This mitigates the effects of bright, narrow band, interference  being scattered into Reionization delay modes and results in a time and frequency bin of length 39.6s and width 492.61kHz.
   This process reduces the data volume by a factor of $\sim$20, or roughly an order of magnitude improvement on traditional time and frequency averaging which in this array would be limited to ~100kHz and 10s to avoid averaging away fringes.
@@
 \subsection{Calibration}
-We model the gain as a per-antenna amplitude, a phase slope -physically a single time delay- and a single passband for all antennae.
+We model the gain as a per-antenna amplitude and a phase slope -physically a single time delay- and single low order polynomial passband for all antennae.
 Because the array samples correlations redundantly, the relative calibration between antennae is numerically overdetermined and tractable as a linear algebra problem \cite{Liu:2010p10391}.
 As described in P14, we compute the ratio between redundant baselines, fit for a gain and phase slope and then algebraically solve for a per-antenna solution.
-Here we have avoided calibrating each channel independently to preserve as much frequency variation as possible.
+Using this method we have avoided calibrating each channel independently to preserve as much frequency variation as possible.
 These solutions vary little over the three month observing period, exhibiting less than 1\% rms variation.
@@
 The redundant solution contains two free delay parameters and an overall amplitude scale.
-We fit the two delay parameters to a model of Pictor A, Fornax A, and the Crab Nebula during a time when the sky is dominated by these three sources while marginalizing over the unknown apparent flux ratio between the three sources\footnote{Unknown in the sense of a joint uncertainty in source flux and primary beam pattern.}.
+We fit the two delay parameters to a model of Pictor A, Fornax A, and the Crab Nebula during a time when the sky is dominated by these three sources while marginalizing over the unknown\footnote{Unknown in the sense of a joint uncertainty in source flux and primary beam pattern.} apparent flux ratio between the three sources.
 With the delays in place we are now able form a beam on Pictor A and (for each channel) set the overall amplitude to the calibration value of 381Jy found in \cite{jacobs:2013b}.
@@
 no missing channels while maximizing coverage over the band.
-Redshift 7.7 is included for comparison with P14.
+Redshift 7.68 is included for comparison with P14.
 Channels with no missing data are indicated by the dotted line.
@@
 \end{figure}
-%XXX more, and a plot of those redundancy residuals
+
 \subsection{Foreground Filtering and Night Averaging}
@@
 Where during the previous compression step a liberal horizon of 300m (1800ns, much longer than the 30 meter baselines under study)  was used to calculate the window size, we now choose a window corresponding to the 30m baselines under study.
-The CLEAN model that is iteratively built is then subtracted from the data leaving residuals which we carry into the next stage.
+The broadband delay spectrum model is iteratively built then subtracted from the data leaving residuals which we carry into the next stage.
 Next, a four hour long running mean is subtracted.
 This removes excess correlation due to cross-talk in the analog signal chain.
-The residuals are then flagged once more for RFI before the 92 nights of data are averaged into 36.4 second long local sidereal time (LST) bins.
-During averaging we found that some LST bins were dominated by very bright data points due to some exceedingly bright outlier samples.
-As these were very rare we found that removing the 10\% brightest from every bin, a small median filter, was quite effective.
+The residuals are then flagged once more for RFI before the 92 nights of data are averaged into 36.4 second long local sidereal time (LST) bins, which as PAPER is a drift scanning instrument, are equivalent to bins in Right Ascension/ Declination.
+During averaging we found that some LST bins were dominated by a small number of exceedingly bright samples lying well outside the rest of the gaussian distributed data.
+To compensate we filter the 10\% brightest sources in each bin.
 The source of these outliers is not known,  a likely possibility is instability in the analog signal chain stimulated by weather or bright interference, a circumstance that has since been observed in later seasons.
@@
 Sky-like fringes on a 30m baseline rotate much slower than 40 seconds.
-Performing a final fringe-rate filter, this time limiting to fringe-rates expected on a 30m baseline (down from 300m in the last iteration) we arrive at a data-set averaged to 900s, the maximum possible while still maintaining coherency.
+Performing a final fringe-rate filter, limiting to fringe-rates expected on a 30m baseline (down from 300m in the last iteration), we arrive at a data-set averaged to 789s, the maximum possible while still maintaining fringe coherence.
 The root mean square of the residual signal (seen in Figure \ref{fig:noise}) at the end of this process is close to the 3mK level expected given the total integration time.
@@
 The residual correlations are all samples of the power spectrum.
-To estimate the final power spectrum and its uncertainty we compute the mean and variance of many random randomly-selected subsamples, sampling the dimensions: sidereal time, redundant baseline pair, and delay sign\footnote{As visibilities are complex, both the positive and negative delays  carry  information.}
+To estimate the final power spectrum and its uncertainty we compute the mean and variance of many random randomly-selected subsamples, sampling the dimensions: sidereal time, redundant baseline pair, and delay sign\footnote{As visibilities are complex, both the positive and negative delays  carry  information.
+Physically the two signs correspond to the two halves of the sky.}
 
@@
 \includegraphics[width=0.45\textwidth]{{figures/pspec_log_z_7.94}.png}
+\includegraphics[width=0.45\textwidth]{{figures/pspec_log_z_7.55}.png}
+\caption{\label{fig:pspecs} Left to right from top: power spectra centered on redshifts 10.3,8.5, 7.9, 7.68 and 7.6 (frequencies: 127, 149.5, 159.5 164.5, and 167 MHz.)  All with an effective bandwidth of $\pm$10MHz  covering the redshift span $\Delta z\approx$1421MHz ($B/f^2$) which ranges from 0.8 at $z=$10.3 to 0.45 at $z=$7.4.
+The noise curve (blue dashes) is calculated using the method described in \cite{Pober:2014p10390}  and indicates the 1$\sigma$ confidence bounds on data points consisting purely of noise; 65\% of uncorrelated noise like data points will lie below the curve.
+However, note the caveats that due to the weighting of the delay transform, adjacent $k$ bins are 50\% correlated and that the sensitivity estimate is probably accurate to within a factor of two.
+The black line is a fiducial model at 50\% ionization \citep{Lidz:2008p8251}.
+.
+}
+
+\end{figure}
+\begin{figure}
+\centering
 \includegraphics[width=0.45\textwidth]{{figures/pspec_log_z_7.68}.png}
-\includegraphics[width=0.45\textwidth]{{figures/pspec_log_z_7.55}.png}
-\caption{\label{fig:pspecs} Left to right from top: power spectra centered on redshifts 10.3,8.5, 7.9, 7.7 and 7.6 (frequencies: 127, 149.5, 159.5 164.5, and 167 MHz.)  All with an effective bandwidth of $\pm$10MHz  covering the redshift span $\Delta z\approx$1421MHz ($B/f^2$) which ranges from 0.8 at $z=$10.3 to 0.45 at $z=$7.4.
-The noise curve (blue dashes) is calculated using the method described in \cite{Pober:2014p10390}  and indicates the 1$\sigma$ confidence bounds on data points consisting purely of noise; 65\% of uncorrelated noise like data points will lie below the curve.
-Note that due to the weighting of the delay transform, adjacent $k$ bins are 50\% correlated.
-The black line is a fiducial model at 50\% ionization \citep{Lidz:2008p8251}.
-The redshift 7.68 bin has been reprocessed for comparison with the P14 result.
-Note that the LST range processed here is 3 hours less than P14 for a small reduction in sensitivity but large reduction in foreground.
-}
+\caption{\label{fig:P14compare} The redshift 7.68 bin has been reprocessed for comparison with the P14 result.
+Note that the LST range processed here is 3 hours less than P14 for a small reduction in sensitivity but large reduction in foreground }
 \end{figure}
+
 \section{Results}
@@
 In Figure \ref{fig:noise} we examine $T_{rms}$ as calculated by differencing between adjacent channels and between redundant baselines.
-We see that in the last step of coherent integration, the step between 40 second to 789 second integrations, the noise level decreases by the expected factor of $\sqrt{789/40}$.
-At both stages the noise level found by both methods is roughly consistent with the theoretical level, only deviating significantly at the edges where the effective signal to noise due to the Blackman Harris window drops precipitously.
+We see that in the last stage of coherent integration, the step between 40 second to 789 second integrations, the noise level decreases by the expected factor of $\sqrt{789/40}$.
+At both stages the noise level found by both methods is roughly consistent with the theoretical level, only deviating significantly at the edges where the effective signal to noise due to the decreasing passband gain, Blackman Harris window, and uneven sampling (see Fig \ref{fig:bins}) drops precipitously.
 
-The theoretical noise level is a straightforward calculation, but given the method employed here, should be given due description.
-Not every LST bin is observed an equal number of times nor is the noise level constant, as the galaxy moves overhead.
-the number of points in each bin is mostly determined by the observing schedule, which for PAPER is simply sunrise to sunset, but also by any flagging of data or observing outage (only 92 of the 103 days in the observing period have recorded data).
+The theoretical visibility rms noise level is a straightforward calculation, but given the method employed here, should be given due description.
+The noise level (eq \ref{eq:Trms}) is caused by the temperature of the sky (mostly due to galactic emission) which varies with LST and by the system noise, both on the scale of a few 100K.
+The noise of any long integration is also dependent on the number of observations in each LST.
+The number of points in each bin is mostly determined by the observing schedule, which for PAPER is sunrise to sunset, but also by any flagging of data or observing outage (only 92 of the 103 days in the observing period have recorded data).
 Unfortunately, technical problems limited the retention of the exact cumulative count of points averaged into each LST bin during the averaging process.
@@
 Some bins are observed only a few times, while a slim range in the LST range 6 - 9 hours is observed on every night.
-On average each bin is measured 64 times.
-Given this accounting of the number of samples $N(lst)$, we estimate the final rms temperature to be
+Over a number of lst bins $N_{lst}$ the average effective time per bin $N_{eff}$ is given by
 
 \[
-T_{rms}  = \frac{T_{g}(lst)}{\sqrt{2BtN(lst)}} + T_{rcvr}
+\frac{1}{\sqrt{N_{eff}}} = \sum_\textrm{lst=1}^\textrm{lst=9}{\frac{N_{lst}}{\sqrt{N(lst)}}}
+\]
+
+On average each bin is in this dataset is effectively measured 39 times. Given this accounting of the number of samples $N(lst)$, we estimate the final rms temperature to be
+
+\[
+T_{rms}  = \frac{T_{sky}(lst) + T_{rcvr}}{\sqrt{2BtN(lst)}} \approx \frac{T_{sky} + T_{rcvr}}{\sqrt{2BtN_{eff}}}
+\label{eq:Trms}
 \]
 where B and t are the 496kHz channel width and lst bin length, respectively.
-We model galactic temperature $T_g$ by summing the de Oliviera-Costa global sky model \citep{deOliveiraCosta:2008p2242} convolved by the PAPER beam as a function of lst and $T_{rcvr}$=320K as the system temperature of 550K (found in P14) minus the mean galactic temperature from that time period.
-The final result is plotted as a dotted line in Figure \ref{fig:noise} for before and after fringe rate filtering (t=40s and 788s)%XXX TODO fold this back into figure 1 so that it shows the correct noise estimate
+We model galactic temperature $T_{sky}$ by summing the \citep{deOliveiraCosta:2008p2242} global sky model  convolved by the PAPER beam as a function of lst averaged over the targeted lst range and $T_{rcvr}$=300K as the system temperature of 550K (found in P14) minus the mean galactic temperature from that time period (250K).
+The final result is plotted as a dotted line in Figure \ref{fig:noise} for before and after fringe rate filtering (t=40s and 789s)
 
@@
 Recall that the wide-band foreground model we have subtracted is built by weighting the entire band by a Blackman-Harris window function.
-Foreground removal is much weaker at this frequency, with the model subtracted being centered around 150MHz, contributions from this point in the band are down-weighted by 75\%.
 This filter leads to the notch of missing power near $k_\parallel=0$  (except at z=7.68 where the foregrounds have been left in for comparison).
 Only at redshift 10.3 is this notch nearly non-existent.
-The relative distribution of foreground removal accuracy is also visible in the shape of $\Delta^2$ vs redshift at $k=0.1$ which is only a factor of two away from the light travel horizon.
-This bin, which is firmly well into the foreground dominated portion of $k$ space has much the same shape as the $T_rms$ curve in Figure \ref{fig:noise}, with a minimum near redshift 8.5 and a dramatic rise above redshift 10.
+Contributions to the foreground model from 126Mhz ($z=$10.3) are down-weighted by $\sim$75\%.
+The $k=0.1$ slice (upper left of figure \ref{fig:slices}) is only a factor of $\sim$three away from the light travel horizon of $k=0.035$ and therefor most likely to be contaminated by low $k$ foregrounds.
+This $k$ mode is well above the noise and has a redshift  dependence  much the same shape as the $T_{rms}$ curve in Figure \ref{fig:noise}, with a minimum near redshift 8.5 and a dramatic rise above redshift 10.
 
 %Across the entire band some $k$ modes appear to be dominated somewhat by non-noise systematics, these are not always clearly foregrounds because they appear at only certain $k$ modes rather than clearly emanating from small delays like in the z=10.3 power spectrum.
+
+%Possible assessment of comparison to noise.  Opt/Mid/Pes noise estimates.  If Opt noise => signal, if Pes=>systematics
 
@@
 Left to right from top, k=0.1,0.2,0.3,0.4 hMpc$^{-1}$.
-Paciga 2013 GMRT marked with an 'x', Parsons 2014 PAPER limit marked with thin black, this work marked with thick blue diamonds.
+Parsons 2014 PAPER limit marked with thin black, this work marked with thick blue diamonds.
 The k=0.1 hMpc$^{-1}$ bin (top-left), which samples the delay spectrum at only 2x the maximum horizon delay  is foreground dominated with  a redshift dependence similar to the  $T_{rms}$ residual in Figure \ref{fig:noise}.
-The rest are consistent with thermal noise.
+The rest have amplitudes below the predicted noise level, but are not themselves perfectly consistent with noise, particularly at high redshifts.
+Though the points are all within the region predicted to be dominated by noise (grey) the variation between the redshifts and size of error bars suggests that a small residual signal is probably still present.
+The noise estimate incorporates a number of simplifications and could easily be off by a factor of 2-4.
 \label{fig:slices}}
 \end{figure}
+
+%\subsection{Future Work}
+%Further work is in progress to refine methods to include partially sampled data.
 
@@
 
-With this paper we have extended the redshift coverage of \cite{Parsons:2013p9876} to cover the range $10.5>z>7.6$.
-The amount of foreground contamination slowly increases with redshift, paralleling the level of incoherent noise observed.
-This highlights a possible limitation of the scheme used to subtract the broadband foreground model which trades spectral coverage for isolation between delay modes.
-Accessing higher redshifts, where there is less spectral information on which to base a foreground model will require some combination of methods for removing foregrounds that make maximal use of the full band without as much down-weighting, as well as a more precise method for removing foregrounds in the narrow band-power spectrum.
+With this paper we have extended the redshift coverage of \cite{Parsons:2013p9876} to cover the range $10.3>z>7.6$.
+Most of the power spectrum data points demonstrate a robust removal of foreground signals to the thermal noise limit, leaving systematics on the scale of the noise.
+Of particular note is the increased difficulty removing foregrounds at redshift 10.3.
+This appears to be largely due to the centrally weighted Blackman-Harris filter used to provide a high dynamic range delay transform in the broadband foreground filter.
+Future work will focus on more effectively removing foregrounds nearer to the edge of the band, opening more of the spectrum by being more tolerant of missing data, and better handling the removing of what residual foregrounds we do see with spectrum estimators which more robustly estimate power in the presence of residual foregrounds (e.g.
+\cite{Liu:2011p8763}).
 
-With only 32 antenna, this data lacks the sensitivity to exclude all but the most extreme models, however, it does demonstrate the ability of the wideband filtration method to affectively remove foregrounds to the precision needed to integrate a full season to the thermal noise limit.
+%With only 32 antenna, this data lacks the sensitivity to exclude all but the most extreme models, however, it does demonstrate the ability of the wideband filtration method to affectively remove foregrounds to the precision needed to integrate a full season to the thermal noise limit.
 As future observations with PAPER  will add sensitivity primarily by increasing to 128 antennae this projects well for our ability to reach design sensitivity.

### 2014-06-26 ms.tex
large block of edits from Pober, Liu, and Carilli

@@
 Richard F. Bradley\altaffilmark{4,5,6},
-Chris L.  Carilli\altaffilmark{7},
+Chris L.  Carilli\altaffilmark{7,10},
 David R. DeBoer\altaffilmark{8},
@@
 We present new observations from the Donald C.
-Backer Precision Array for Probing the Epoch of Reionization (PAPER) telescope which place new lower limits on the HI power spectrum over the redshift range of $7.5<z<10.5$, extending previously published single redshift results to cover the full range accessible to the instrument.
+Backer Precision Array for Probing the Epoch of Reionization (PAPER) telescope which place new deeper limits on the HI power spectrum over the redshift range of $7.5<z<10.5$, extending previously published single redshift results to cover the full range accessible to the instrument.
 The epoch of reionization power spectrum is expected to evolve strongly with redshift, and it is this variation with cosmic history that will allow us to begin to place constraints on the physics of reionization.
 The primary obstacle to this goal are bright foregrounds and the attendant systematics associated with high dynamic range measurements.
-Here we use foreground removal techniques that take advantage of the large instrumental bandwidth to make a high dynamic range estimate of smooth foreground power.
-Power spectra at different points across the spectrum reveal the variable efficacy of this subtraction.
-Noise limited measurements of $\Delta^2$ at $k=$0.2hMpc$^{-1}$ reach as low as 29mK at redshift 7.5.
+Here we use filtering techniques that take advantage of the large instrumental bandwidth to make a high dynamic range subtraction of foreground power.
+Power spectra at different points across the redshift range reveal the variable efficacy of this subtraction.
+Noise limited measurements of $\Delta$ at $k=$0.2hMpc$^{-1}$,z$=7.5$ reach as low as 58mK ($2-\sigma$).
 Most spectra demonstrate a robust foreground removal of foreground signals to the thermal noise limit, leaving systematics on the scale of the noise.
+To better understand the source of these residual systematics we examine different estimates of the noise in the data and compare with theoretical predictions.
 We find that in the case of the highest redshift power spectrum, foregrounds begin to dominate over the noise, reflecting the increased difficulty in filtering foregrounds near the edge of the instrumental band.
@@
 \section{Introduction}
-The Epoch of Reionization, when the first stars ionized the pervasive cosmological Hydrogen in the last global phase change is predicted to be observable in highly redshifted 21 cm radiation.
+The Epoch of Reionization, when the first luminous objects ionized the pervasive cosmological hydrogen is predicted to be observable in highly redshifted 21 cm radiation.
 The Donald C.
 Backer Precision Array for Probing the Epoch of Reionization (PAPER, \cite{Parsons:2010p6757})\footnote{\url{eor.berkeley.edu}} is a low frequency radio interferometer experiment dedicated to opening this window on the universe.
-Challenges include foregrounds which are brighter by several orders of magnitude and long integration times necessitated by the limited collecting areas of first generation instruments.
-Direct observation of hydrogen before and during re-ionization is predicted to deliver a wealth of cosmological and astrophysical data, including the nature of the first stellar objects and the timing and rate of galaxy formation, reviews on the physics of reionization as well as theory about the nature of foregrounds maybe be found in \citet{Furlanetto:2006p2267,Morales:2010p8093,Pritchard:2012p9555}.
+Challenges include foregrounds which are brighter by several orders of magnitude and long integration times necessitated by the limited collecting areas of first-generation instruments.
+Direct observation of hydrogen before and during re-ionization is predicted to deliver a wealth of cosmological and astrophysical data, including the nature of the first stellar objects and the timing and rate of galaxy formation.
+Reviews on the physics of reionization as well as theoretical expectations on the nature of foregrounds may be be found in \citet{Furlanetto:2006p2267,Morales:2010p8093,Pritchard:2012p9555}.
 
-Telescopes seeking to measure this signal include the Giant Metre-wave Radio Telescope (GMRT; \cite{Paciga:2013p9943}), the Low Frequency Array (LOFAR; \cite{Yatawatta:2013p9699}) and the Murchison Widefield Array (MWA; \cite{Bowman:2013p9950} and \cite{Tingay:2013p9022}).
-PAPER is located in the Karoo desert at the site of the South African portion of the future Square Kilometer Array.
-PAPER has doubled in size on a yearly basis since 2009, making observations with each array.
+Other telescopes seeking to measure this signal include the Giant Metre-wave Radio Telescope (GMRT; \cite{Paciga:2013p9943}), the Low Frequency Array (LOFAR\footnote{\url{www.lofar.org}}; \cite{Yatawatta:2013p9699}) and the Murchison Widefield Array (MWA\footnote{\url{mwatelescope.org}}; \cite{Bowman:2013p9950} and \cite{Tingay:2013p9022}).
 
-Here we report on deep integrations made with a 32 element array in 2011, first described in \cite{Parsons:2013p9876}, hereafter P14.
-In P14 we described in detail our reduction method and gave deep new limits on the HI power spectrum in the presence of bright foregrounds at redshift 7.68 using 92 nights (1800 hours) of data.
-The epoch of reionization signal is expected to evolve strongly with redshift, and it is this variation with cosmic history that will allow us to begin to place constraints on the physics of reionization \cite{Pritchard:2008p8123,Pober:2014p10390}.
-Therefore, while a detection of the 21 cm signal at even a single frequency would be a tremendous breakthrough, analysis techniques must be developed to capitalize on the wide bandwidths of the current generations of high redshift 21cm telescopes.
-Using the same data as P14, this paper presents new power spectrum upper limits covering the redshift range $10.5>z>7.2$.
-In section  \ref{sec:obs_meth} we summarize the observation and reduction methodology, we present the new upper limits in \ref{sec:results}, and in \ref{sec:conclusion} we offer conclusions and discussion of future work.
+PAPER is located in the Karoo desert at the site of the South African portion of the future Square Kilometer Array\footnote{\url{skatelescope.org}} and has doubled in size on a yearly basis since 2009, and science-grade observations have been made with each stage of the build-out.
+
+Here we report on deep integrations made with a 32 element array in 2011, first described in \cite{Parsons:2014p10499}, hereafter P14.
+Our data reduction method was described in detail in P14, and used this pipeline to give the deepest yet limits on the HI power spectrum in the presence of bright foregrounds at redshift 7.68 using 92 nights (1800 hours) of data.
+The Epoch of Reionization signal is expected to evolve strongly with redshift, and it is this variation with cosmic history that will allow us to begin to place constraints on the physics of reionization \citep{Pritchard:2008p8123,Pober:2014p10390}.
+Therefore, while a detection of the 21 cm signal at even a single frequency would be a tremendous breakthrough, analysis techniques must be developed to capitalize on the wide bandwidths of the current generation of high-redshift 21cm telescopes.
+Using the same data set as P14, this paper presents improved upper limits on the HI power spectrum over the redshift range $10.5>z>7.2$.
+In Section  \ref{sec:observations} we summarize the observations, in Section \ref{sec:obs_meth} review the reduction methodology, we present the new upper limits in \ref{sec:results}, and in Section \ref{sec:conclusion} we offer conclusions and discussion of future work.
 
@@
 The work here follows the same basic procedure and uses the same underlying data set as P14.
-Here we provide a quick summary and refer the reader to \cite{Parsons:2013p9876} for a more in depth discussion.
-A general overview of the PAPER system can be found in \cite{Parsons:2010p6757}, calibration of the primary beam in \cite{Pober:2012p8800}, and imaging results in \cite{Jacobs:2011p8438,Jacobs:2013p10014,Stefan:2013p9926}.
+Here we provide a quick summary and refer the reader to P14 for a more in depth discussion.
+A general overview of the PAPER system can be found in \cite{Parsons:2010p6757}, calibration of the primary beam in \cite{Pober:2012p8800}, and imaging results in \cite{Jacobs:2011p8438,jacobs:2013b,Stefan:2013p9926}.
 Sensitivity analysis described in \cite{Parsons:2012p9028} revealed that for the low gain elements employed by PAPER, a highly redundant ``grid'' type arrangement offers a significant sensitivity boost.
-In most interferometers each baseline samples a different mode of the power spectrum; the ideal case for reconstructing images.
-In a grid, there are many samples of each of these cosmological modes which can be averaged to a high sensitivity before being combined with other different $k$ modes.
-The PAPER South Africa 32 antenna deployment (PSA32) was arranged in a 4x8 grid with a column spacing of 30m and a row spacing of 4m.
+In most interferometers the locations of the antennae are optimized such that each baseline samples a different Fourier mode of the sky; this is the ideal case for reconstructing images where each mode contains different information.
+For a power spectrum measurement the key metric is sensitivity per mode, rather than number of modes.
+A grid configuration allows many samples of each cosmological mode, to be averaged to a high sensitivity before being combined with other Fourier modes.
+The PAPER South Africa 32 antenna deployment (PSA32) was arranged in a 4$\times$8 grid with a column spacing of 30m and a row spacing of 4m.
 In our analysis, as in P14, we include only the three shortest types of spacings where the reionization power is expected to be brightest.
 This selection includes those between adjacent columns and within at least one row of each other, a selection containing 70 $\sim$30m-long baselines.
+We will use these baselines to make a one dimensional estimate of the HI power spectrum.
 
-Observations spanning the band between 100 to 200-MHz ($13.1>z>6.1$) were recorded at a resolution of 48kHz and 10.7s  beginning Dec 7, 2011 and ending March 19, 2012 (with some down-time for maintenance) after a total of 92 nights or 1800 hours.
+Observations spanning the band between 100 to 200-MHz ($13.1>z>6.1$) were recorded at a resolution of 48kHz and 10.7s  beginning Dec 7, 2011 and ending March 19, 2012 (with some down-time for maintenance) giving a total of 92 nights or 1800 hours.
 Within this set we included observations in the LST range 1h - 9h where the sky dominated system temperature is at a minimum.
@@
 \label{sec:obs_meth}
-Here we summarize our data reduction steps, for detailed descriptions see section 3 of P14.
+Here we summarize our data reduction steps; for detailed descriptions see Section 3 of P14.
+In summary, we use 70 nearly identical baselines to make a 1D estimate along the spectral or line-of-sight direction of the reionization era HI power spectrum.
+All processing, save calibration, and  the final cross-multiplication step treats each baseline as independent.
+Foregrounds and interference are removed on a per-baseline basis with no a-priori sky model using signal processing techniques and a physical model of the array.
+In the final cross-multiplication step, the last layer of systematics is estimated and removed by projecting non-physical correlations between baselines.
 
 In several stages throughout the analysis process we take a 2D Fourier transform of the visibility spectra $V(\nu,t)$ into ``delay/fringe rate'' space where delay is the Fourier dual to frequency and similarly fringe-rate for time.
-In this space, smooth spectrum sources are physically localized to delays shorter than the light travel time length of the baseline (e.g.
-sources at the horizon, in the direction of the baseline vector) and fringe rates shorter than the sidereal rate.
+In this space, smooth spectrum sources are physically localized to delays shorter than the light travel time length of the baseline and fringe rates shorter than the sidereal rate.
+Sources at the horizon, in the direction of the baseline vector, have the longest delays, while fringe rates are highest where the celestial equator crosses the horizon.
+%TODO is that right?
  In this Fourier space, sources are highly localized with deviations from a flat spectrum manifesting as a slight dispersion.
-The spectrum sampling function, which is uneven due to flagging of interference takes the form of a convolution by a point-spread-function in the same way an imperfect sampling of the uv plane gives rise to the angular psf of an interferometer.
-To account for this, we use a CLEAN like, iterative, peak-finder and subtraction algorithm which is limited to finding peaks within the physically allowable ranges.
+The spectrum sampling function, which is uneven due to flagging of interference takes the form of a convolution by a point-spread-function (PSF) in the same way an imperfect sampling of the $uv$ plane gives rise to the angular PSF of an interferometer.
+If enough data is missing this PSF can cause smooth-spectrum sources to leak beyond the horizon.
+To account for this, we use a CLEAN like, iterative, peak-finder and subtraction algorithm which is limited to finding peaks within the physically allowable ranges of delay and/or fringe-rate In this case, the 1D ``CLEAN'' beam is the Fourier transform of the spectral or time sampling function \citep{Parsons:2009p7859}.
 
@@
 This final step takes advantage of the redundant baselines to make an unbiased power spectrum estimate by cross-multiplying identical baselines and then averaging the power spectrum modes.
-Here we are also using the approximation that over short ($\delta z < \sim 0.5$) the delay spectrum is a good approximation of a sample of the 3D power spectrum \citep{Parsons:2012p8896}.
 Thus as we step through the various iterations of delay spectrum we have the ability to isolate and subtract foregrounds in the native domain of the measured power spectrum.
-Note that isolation of foregrounds to a region below line-of-site $k_\parallel$s that increases with baseline length is also the much discussed "wedge" \cite{Liu:2014p10462,Liu:2014p10463,Pober:2013p9942,Morales:2012p8790,Vedantham:2012p10297,Datta:2010p8781}.
-As we only have a single length of baseline the wedge becomes a narrow slice indeed.
+Note that isolation of foregrounds to a region below line-of-sight $k_\parallel$s that increases with baseline length is also the much discussed ``wedge'' \cite{Liu:2014p10462,Liu:2014p10463,Thyagarajan:2013p10039,Pober:2013p9942,Trott:2012p10466,Morales:2012p8790,Parsons:2012p8896,Vedantham:2012p10297,Datta:2010p8781,Parsons:2009p7859}.
+As we only have a single length of baseline the wedge manifests as a single value of $k\sim k_\parallel$, below which foregrounds are expected to dominate.
 The details of each application of the delay/fringe-rate transform will be laid out the following sections as we provide a brief walk-through of the processing pipeline.
@@
 We minimize covariance between adjacent redshift bins by limiting overlap of each bin to the spectrum at the outer half of the bin channel range which, as will be described in section \ref{sec:power_spectrum} is significantly down-weighted by the use of a Blackman-Harris window.
-The power spectrum method developed in P14 leverages the redundant baselines to estimate non-sky covariance and project out these contaminated modes.
-This inversion is particularly sensitive to missing data -the inverse of the covariance is not defined- so we also select only channels that have no missing data over the entire sidereal period (see the dotted line in \ref{fig:bins}).
+Avoiding missing data is important because the power spectrum method, developed in P14, which leverages the redundant baselines to estimate non-sky covariance and project out these contaminated modes is particularly sensitive to missing data -the inverse of the covariance is not defined- so we also select only channels that have no missing data over the entire sidereal period (see the dotted line in \ref{fig:bins}).
 With these constraints we arrive at the redshifts 10.3, 8.5,7.9 and 7.5.
-For the purposes of comparison with P14 we also include the redshift 7.86 bin.
+For the purposes of comparison with P14 we also include the redshift 7.68 bin.
+\citet{Dillon:2013p10497}describes a pseudo-inverse method for handling the covariance introduced by missing data.
+Though not implemented here, we will use a version of their method in future analysis.
 
-The spectral width of the redshift bins is dictated by the competing needs of sensitivity and foreground reduction which both benefit from wider bandwidths and the likely rate of reionization evolution which is thought to be close to $\Delta z \approx 0.5$, which to a good approximation is given for bandwidth $B$ and frequency $f$ as $\Delta z=1421 B/f^2\approx 0.5$.
-To balance these two competing constraints we choose a single spectral bandwidth of 20Mhz, weighted by a Blackman-Harris window for an effective width of 10Mhz, or $\Delta z=0.5$ at $z=7.45$ ranging up to a $\Delta z = 0.86$ at redshift 10.3.
+The spectral width of the redshift bins is dictated by two competing needs.
+First, sensitivity and foreground reduction both benefit from wider bandwidths.
+On the other hand, power spectrum measurements which trace the evolution with redshift require smaller bandwidths.
+Often, the evolution scale is taken to be on order $\Delta z \approx 0.5$, which translates to a bandwidth limit $B<0.5 (f^2/1421)$ MHz.
+To balance these two competing constraints we choose a single spectral bandwidth of 20MHz, weighted by a Blackman-Harris window for an effective width of 10MHz, or $\Delta z=0.5$ at $z=7.45$ ranging up to a $\Delta z = 0.86$ at redshift 10.3.
 
@@
 %I know this isn't technically true, but I think its important that the downselect is obvious up front.
-   The visibilities are then compressed in the frequency and time directions by removing delay modes above the horizon and highest frequency fringe rates for a 300m baseline.
+   The visibilities are then compressed in the frequency and time directions by removing delay modes and highest frequency fringe rates corresponding to a 300m baseline (the longest baseline in the array).
 This filtering\footnote{See Appendix A of P14} is done in tandem with a radio frequency interference (RFI) flagging step, using the residuals which have had bright sky-like signals removed to flag 4$\sigma$ deviations before feeding the flags back into another iteration of the compression step.
-This mitigates the effects of bright, narrow band, interference  being scattered into Reionization delay modes and results in a time and frequency bin of length 39.6s and width 492.61kHz.
+This mitigates the effects of bright, narrow band, interference  being scattered into higher delay modes (where reionization lives) and results in a time and frequency bin of length 39.6s and width 492.61kHz.
   This process reduces the data volume by a factor of $\sim$20, or roughly an order of magnitude improvement on traditional time and frequency averaging which in this array would be limited to ~100kHz and 10s to avoid averaging away fringes.
@@
 \subsection{Calibration}
-We model the gain as a per-antenna amplitude and a phase slope -physically a single time delay- and single low order polynomial passband for all antennae.
-Because the array samples correlations redundantly, the relative calibration between antennae is numerically overdetermined and tractable as a linear algebra problem \cite{Liu:2010p10391}.
+We model the gain as a per-antenna amplitude and a phase slope -physically a single time delay- and single real, low order polynomial passband for all antennae.
+Because the array samples correlations redundantly, the relative calibration between antennae is numerically overdetermined and tractable as a linear algebra problem \citep{Liu:2010p10391}.
 As described in P14, we compute the ratio between redundant baselines, fit for a gain and phase slope and then algebraically solve for a per-antenna solution.
 Using this method we have avoided calibrating each channel independently to preserve as much frequency variation as possible.
-These solutions vary little over the three month observing period, exhibiting less than 1\% rms variation.
+These solutions vary little over the three month observing period, exhibiting less than 1\% r.m.s.
+variation.
 A single solution derived for for the Dec 7 data set is used for the entire observing run.
-Time and frequency variation of redundant solutions is explored in detail in Zheng et al (in prep).
+Time and frequency variation of redundant solutions is explored in general in \citet{Zheng:2014p10467}.
 
-Relative calibration is all that is necessary to form a power spectrum, however to set the flux scale, we must form a beam on a bright known point source.
-By itself, the redundant solution is completely independent of any sky model.
-The redundant solution contains two free delay parameters and an overall amplitude scale.
-We fit the two delay parameters to a model of Pictor A, Fornax A, and the Crab Nebula during a time when the sky is dominated by these three sources while marginalizing over the unknown\footnote{Unknown in the sense of a joint uncertainty in source flux and primary beam pattern.} apparent flux ratio between the three sources.
+Relative calibration is all that is necessary to form a power spectrum.
+To obtain a correctly normalized power spectrum, however, it is necessary to set a flux scale for which we need a flux measurement of a calibration standard.
+For this we form a beam on a known, bright calibration source.
+However, by itself the redundant calibration does not contain enough phase information to phase coherently to a sky location, there remain two free phase parameters which cannot be solved by redundancy alone.
+We find these by fitting a model of Pictor A, Fornax A, and the Crab Nebula during a time when the sky is dominated by these three sources while marginalizing over the unknown\footnote{Unknown in the sense of a joint uncertainty in source flux and primary beam pattern.} apparent flux ratio between the three sources.
 With the delays in place we are now able form a beam on Pictor A and (for each channel) set the overall amplitude to the calibration value of 381Jy found in \cite{jacobs:2013b}.
@@
 Redshift 7.68 is included for comparison with P14.
-Channels with no missing data are indicated by the dotted line.
+Channels with no missing data are indicated by the dotted line, visible at the edge of flagged channels.
 Each redshift bin is 20 MHz wide, but weighted by a Blackman-Harris window function which heavily down-weights the outer 10MHz for a Noise Equivalent Bandwidth of 10MHz.
@@
 %The success of the remainder of the processing stages hinges on the spectral stability of the data, which can easily be assessed by comparing redundant baselines.
-In figure XXX we see the rms of the ratios between redundant calibrations, before and after the calibration has been applied.
+In figure XXX we see the r.m.s.
+of the ratios between redundant calibrations, before and after the calibration has been applied.
 \begin{figure}
@@
 \caption{\label{fig:noise} Root mean square (rms) noise before and after integrating to 10 minute LST bins (top and bottom sets of curves) indicating the rising significance of foregrounds as we take the last deep integration step.
-Noise is estimated by differencing adjacent frequencies (magenta) and between redundant baselines (blue), compared with an estimate of the theoretical noise level (dotted).
+Noise is estimated by differencing adjacent frequencies (magenta) and between redundant baselines (blue), compared with an estimate of the theoretical noise level with a $T_{sys}$ of 550K (dotted).
 The top three lines show noise after filtering foregrounds and binning into 40s long sidereal bins.
@@
 Foregrounds are filtered from the calibrated data by removing all bright delay components with light travel times less then the baseline length.
-Where during the previous compression step a liberal horizon of 300m (1800ns, much longer than the 30 meter baselines under study)  was used to calculate the window size, we now choose a window corresponding to the 30m baselines under study.
+Where during the previous compression step a liberal horizon of 300m (1800ns, much longer than the 30 meter baselines under study)  was used to calculate the window size, we now choose a window corresponding to the 30m baselines under study with an extra buffer of 15ns to provide a slight buffer against the 1/B$\sim$12ns resolution of the delay spectrum.
 The broadband delay spectrum model is iteratively built then subtracted from the data leaving residuals which we carry into the next stage.
@@
 This removes excess correlation due to cross-talk in the analog signal chain.
-The residuals are then flagged once more for RFI before the 92 nights of data are averaged into 36.4 second long local sidereal time (LST) bins, which as PAPER is a drift scanning instrument, are equivalent to bins in Right Ascension/ Declination.
+The residuals are then flagged once more for RFI before the 92 nights of data are averaged into 36.4 second long local sidereal time (LST) bins, which as PAPER is a drift-scanning instrument, are equivalent to bins in Right Ascension/ Declination.
 During averaging we found that some LST bins were dominated by a small number of exceedingly bright samples lying well outside the rest of the gaussian distributed data.
-To compensate we filter the 10\% brightest sources in each bin.
-The source of these outliers is not known,  a likely possibility is instability in the analog signal chain stimulated by weather or bright interference, a circumstance that has since been observed in later seasons.
+To compensate we filter the 10\% brightest samples in each bin.
+The source of these outliers is not known,  a likely possibility being an instability in the analog signal chain stimulated by weather or bright interference, a circumstance that has since been observed in later seasons.
 
@@
 \label{sec:power_spectrum}
-The output of the forgoing is a single calibrated and foreground filtered sidereal night.
+The output of the above steps is a single calibrated and foreground filtered sidereal night of visibility data.
 The power spectrum is estimated in the delay spectrum of  a 10MHz bandwidth range centered on the redshift of interest.
@@
 
-The residual correlations are all samples of the power spectrum.
+The residual elements of the correlation matrix corresponding to cross-multiplication of matching delay bins between different baselines are all un-biased samples of the power spectrum.
 To estimate the final power spectrum and its uncertainty we compute the mean and variance of many random randomly-selected subsamples, sampling the dimensions: sidereal time, redundant baseline pair, and delay sign\footnote{As visibilities are complex, both the positive and negative delays  carry  information.
-Physically the two signs correspond to the two halves of the sky.}
+Physically the two signs correspond to the two halves of the sky.}.
+
+%\begin{figure}
+%\centering
+%\includegraphics[width=0.45\textwidth]{{figures/pspec_log_z_10.29}.png}
+%\includegraphics[width=0.45\textwidth]{{figures/pspec_log_z_8.54}.png}
+%\includegraphics[width=0.45\textwidth]{{figures/pspec_log_z_7.94}.png}
+%\includegraphics[width=0.45\textwidth]{{figures/pspec_log_z_7.55}.png}
+%\caption{\label{fig:pspecs} Left to right from top: power spectra centered on redshifts 10.3,8.5, 7.9, 7.68 and 7.55 (frequencies: 127, 149.5, 159.5 164.5, and 167 MHz.)  All with an effective bandwidth of $\pm$10MHz  covering the redshift span $\Delta z\approx$1421MHz ($B/f^2$) which ranges from 0.8 at $z=$10.3 to 0.45 at $z=$7.4.
+The noise curve (dashes) is calculated using the method described in \cite{Pober:2013p9581}  and indicates the 1$\sigma$ confidence bounds on data points consisting purely of noise; 65\% of uncorrelated noise like data points will lie below the curve.
+However, note the caveats that due to the weighting of the delay transform, adjacent $k$ bins are 50\% correlated.
+It is also clear that the sensitivity calculations under-estimate the sensitivity of the measurements by a small factor.
+See \cite{Pober:2014p10390,Pober:2013p9581} for a discussion of the approximations made in those calculations.
+The black line is a fiducial model at 50\% ionization \citep{Lidz:2008p8251}.
+GMRT points from \cite{Paciga:2013p9943} indicated with 'x's and MWA points (also using 32 antenna) are black diamonds.}
+%
+%\end{figure}
 
@@
 \centering
-\includegraphics[width=0.45\textwidth]{{figures/pspec_log_z_10.29}.png}
-\includegraphics[width=0.45\textwidth]{{figures/pspec_log_z_8.54}.png}
-\includegraphics[width=0.45\textwidth]{{figures/pspec_log_z_7.94}.png}
-\includegraphics[width=0.45\textwidth]{{figures/pspec_log_z_7.55}.png}
-\caption{\label{fig:pspecs} Left to right from top: power spectra centered on redshifts 10.3,8.5, 7.9, 7.68 and 7.6 (frequencies: 127, 149.5, 159.5 164.5, and 167 MHz.)  All with an effective bandwidth of $\pm$10MHz  covering the redshift span $\Delta z\approx$1421MHz ($B/f^2$) which ranges from 0.8 at $z=$10.3 to 0.45 at $z=$7.4.
-The noise curve (blue dashes) is calculated using the method described in \cite{Pober:2014p10390}  and indicates the 1$\sigma$ confidence bounds on data points consisting purely of noise; 65\% of uncorrelated noise like data points will lie below the curve.
-However, note the caveats that due to the weighting of the delay transform, adjacent $k$ bins are 50\% correlated and that the sensitivity estimate is probably accurate to within a factor of two.
+\includegraphics[width=0.9\textwidth]{figures/PAPER_32T_new_redshifts.png}
+\caption{ \label{fig:pspecs} PAPER power spectra at four redshifts.
+On the top, $P_k$ spectra (a simple units change from the raw delay spectrum) provide a useful diagnostic on foreground rejection, while on the bottom we plot in $\Delta^2$ cosmological units.
+Redshift increases left to right.
+All spectra have an effective bandwidth of $\pm$10MHz  covering the redshift span $\Delta z\approx$1421MHz ($B/f^2$) which ranges from 0.8 at $z=$10.3 to 0.45 at $z=$7.4.
+The noise curve (dashes) is calculated using the method described in \cite{Pober:2013p9581}  and indicates the 1$\sigma$ confidence bounds on data points consisting purely of noise; 65\% of uncorrelated noise like data points will lie below the curve.
+However, note the caveats that due to the weighting of the delay transform, adjacent $k$ bins are 50\% correlated.
+It is also clear that the sensitivity calculations under-estimate the sensitivity of the measurements by a small factor.
+See \cite{Pober:2014p10390,Pober:2013p9581} for a discussion of the approximations made in those calculations.
 The black line is a fiducial model at 50\% ionization \citep{Lidz:2008p8251}.
-.
+GMRT points from \cite{Paciga:2013p9943} indicated with 'x's and MWA points (also using 32 antenna) are black diamonds.
 }
+\end{figure}
 
-\end{figure}
 \begin{figure}
 \centering
-\includegraphics[width=0.45\textwidth]{{figures/pspec_log_z_7.68}.png}
+\includegraphics[width=0.9\textwidth]{{figures/pspec_log_z_7.68}.png}
 \caption{\label{fig:P14compare} The redshift 7.68 bin has been reprocessed for comparison with the P14 result.
-Note that the LST range processed here is 3 hours less than P14 for a small reduction in sensitivity but large reduction in foreground }
+Note that the LST range processed here is 3 hours less than P14 for a small reduction in sensitivity but large reduction in foreground.
+}
 \end{figure}
@@
 \label{sec:noise}
-The root-mean-square of the residual measures how well we've removed foregrounds.
+The root-mean-square of the filtered visibilities measures how well we've removed foregrounds.
 In Figure \ref{fig:noise} we examine $T_{rms}$ as calculated by differencing between adjacent channels and between redundant baselines.
@@
 
-The theoretical visibility rms noise level is a straightforward calculation, but given the method employed here, should be given due description.
-The noise level (eq \ref{eq:Trms}) is caused by the temperature of the sky (mostly due to galactic emission) which varies with LST and by the system noise, both on the scale of a few 100K.
+The theoretical visibility r.m.s. noise level is a straightforward calculation, but given the method employed here, should be given due description.
+The noise level in a long integration (Eq.
+\ref{eq:Trms}) is caused by the temperature of the sky (mostly due to galactic emission) which varies with LST and by the receiver noise, both on the scale of a few 100K.
 The noise of any long integration is also dependent on the number of observations in each LST.
@@
 Some bins are observed only a few times, while a slim range in the LST range 6 - 9 hours is observed on every night.
-Over a number of lst bins $N_{lst}$ the average effective time per bin $N_{eff}$ is given by
+Over a number of LST bins $N_\textrm{lst}$ the average effective time per bin $N_{eff}$ is given by
 
 \[
-\frac{1}{\sqrt{N_{eff}}} = \sum_\textrm{lst=1}^\textrm{lst=9}{\frac{N_{lst}}{\sqrt{N(lst)}}}
+\frac{1}{\sqrt{N_{eff}}} = \frac{1}{N_{\textrm{lsts}}} \sum_\textrm{lst=1}^\textrm{lst=9}{\frac{1}{\sqrt{N(lst)}}}
 \]
 
-On average each bin is in this dataset is effectively measured 39 times. Given this accounting of the number of samples $N(lst)$, we estimate the final rms temperature to be
+On average each bin is in this dataset is effectively measured 39.6 times. Given this accounting of the number of samples $N(lst)$, we estimate the final r.m.s. temperature to be
+
+%\[
+%T_{rms}  =\frac{1}{N_{\textrm{lsts}}}\sum_{\textrm{lst}} \frac{T_{sky}(lst) + T_{rcvr}}{\sqrt{2BtN(lst)}} \approx \frac{1}{\sqrt{2BtN_{eff}}} \left(\frac{1}{N_{\textrm{lsts}}}\sum_{\textrm{lst}}T_{sky}(lst) + T_{rcvr}\right)
+%\label{eq:Trms}
+%\]
 
 \[
-T_{rms}  = \frac{T_{sky}(lst) + T_{rcvr}}{\sqrt{2BtN(lst)}} \approx \frac{T_{sky} + T_{rcvr}}{\sqrt{2BtN_{eff}}}
+T_{rms}  =\frac{1}{N_{\textrm{lsts}}}\sum_{\textrm{lst}} \frac{T_{sky}(lst) + T_{rcvr}}{\sqrt{2BtN(lst)}} \approx \frac{(<T_{sky}> + T_{rcvr})}{\sqrt{2BtN_{eff}}}
 \label{eq:Trms}
 \]
-where B and t are the 496kHz channel width and lst bin length, respectively.
-We model galactic temperature $T_{sky}$ by summing the \citep{deOliveiraCosta:2008p2242} global sky model  convolved by the PAPER beam as a function of lst averaged over the targeted lst range and $T_{rcvr}$=300K as the system temperature of 550K (found in P14) minus the mean galactic temperature from that time period (250K).
-The final result is plotted as a dotted line in Figure \ref{fig:noise} for before and after fringe rate filtering (t=40s and 789s)
+
+where B and t are the 496kHz channel width and lst bin length, respectively and the approximation indicates that we have assumed average values for the number of samples and the time dependent sky component of the system temperature.
+In P14 the total system temperature ($T_{sky} + T_{rcvr}$) was estimated to be 550K using 20 nights of the 92 night set which covered LSTs 3-9.
+Using the global sky model  \citep{deOliveiraCosta:2008p2242} and a model of the PAPER primary beam we estimate the mean $T_{sky}$ during that 20 night period to be 250K, implying a receiver temperature of 300K.
+The mean temperature over the full 92 night period is only 4K higher then the mean from used to make the 550K estimate in P14.
+The P14 estimate of 550K for the system temperature is a good estimate for the full observing period.
+$T_{rms}$ using 550K and an effective sample count of 39.6  is plotted as a dotted line in Figure \ref{fig:noise} for before and after fringe rate filtering (t=40s and 789s)
 
@@
 For this reason the noise curves (bottom three curves of Figure \ref{fig:noise})  merit closer inspection.
-At this sensitivity level we see a notable excess of power in the rms difference between baselines.
-Though the cause of this excess is unknown, it is suggestive of the excess covariance between redundant baselines which we remove in Section \ref{sec:power_spectrum} a hypothesis also supported by the high levels of residual foregrounds in the redshift 10.5 bin where the baseline difference rms is highest.
+At this sensitivity level we see a notable excess of power in the r.m.s.
+difference between baselines.
+Though the cause of this excess is unknown, it is suggestive of the excess covariance between redundant baselines which we remove in Section \ref{sec:power_spectrum} a hypothesis also supported by the high levels of residual foregrounds in the redshift 10.5 bin where the baseline difference r.m.s.
+is highest.
 
 \subsection{Power Spectra}
+\label{sec:pspecs}
 
-In  Figure \ref{fig:pspecs} we show the  power spectra at different redshift bins --showing both the spherically averaged power spectrum plotted in $P_k$ units (left) and the final estimate of $\Delta^2$ -- and in Figure \ref{fig:slices} we see different $k$ mode slices as a function of redshift  all plotted with $2\sigma$ error bars.
-The theoretical noise power spectrum is estimated using the method described in \cite{Pober:2014p10390} assuming a system temperature of 550K and the observing scheme described in Section \ref{sec:observations}.
-Generally, both figures convey a picture of a power spectrum dominated by noise, with foregrounds encroaching from modes closest to the foreground contaminated modes.
-Foreground dominance appears to grow with increasing redshift, mirroring the rise in baseline-to-baseline rms we found in Figure \ref{fig:noise}.
-Recall that the wide-band foreground model we have subtracted is built by weighting the entire band by a Blackman-Harris window function.
-This filter leads to the notch of missing power near $k_\parallel=0$  (except at z=7.68 where the foregrounds have been left in for comparison).
-Only at redshift 10.3 is this notch nearly non-existent.
-Contributions to the foreground model from 126Mhz ($z=$10.3) are down-weighted by $\sim$75\%.
-The $k=0.1$ slice (upper left of figure \ref{fig:slices}) is only a factor of $\sim$three away from the light travel horizon of $k=0.035$ and therefor most likely to be contaminated by low $k$ foregrounds.
+In  Figure \ref{fig:pspecs} we show the  power spectra at different redshift bins --showing both the spherically averaged power spectrum ($P(k)$, left) and the r.m.s.
+fluctuation per $k$ interval where $\Delta^2(k)\def\frac{k^3}{2\pi^2}P(k)$.In figure \ref{fig:slices} we see different $k$ mode slices as a function of redshift,  all plotted with $2\sigma$ error bars and in Table \ref{tab:data} we list the data plotted in these slices.
+The theoretical noise power spectrum is estimated using the method described in\cite{Pober:2013p9581}, assuming a system temperature of 550K and the observing scheme described in Section \ref{sec:observations}.
+Generally, both figures convey a picture of a power spectrum dominated by noise, with foregrounds encroaching from inside and just outside the horizon to higher $k$ ``reionization'' modes.
+
+One easy way to see this leakage is by comparing the bootstrap error bar, the theoretical error and the power spectrum value.
+At the level of noise here, we expect all points outside the horizon to be consistent with zero.
+As a simple test we divide the points into ``Detections'' and ``Upper Limits'' (Det/ULim) in Table \ref{tab:data} for both kinds of error bar.
+As we can see, the larger theoretical error bars are mostly consistent with zero, while the same is true of a little less then half of the bootstrap error bars.
+This is consistent with a picture of systematics or foreground leakage near the thermal noise limit.
+
+Foreground dominance appears to grow with increasing redshift, mirroring the rise in baseline-to-baseline r.m.s.
+we found in Figure \ref{fig:noise}.
+Recall that the wide-band foreground model we have subtracted  (which causes the notch of missing power near $k_\parallel=0$ in Figure \ref{fig:pspecs})\footnote{except at z=7.68 where the foregrounds have been left in for comparison} is the same for all redshift slices and was built by weighting the entire band by a Blackman-Harris window function.
+Using this window, contributions to the foreground model from 126MHz ($z=$10.3) are down-weighted by $\sim$75\%.
+If the 126MHz sky or systematics are much different then 160Mhz, the foreground filter will be less effective.
+The $k=0.1$hMpc$^{-1}$ slice (upper left of figure \ref{fig:slices}) is only a factor of $\sim$three away from the light travel horizon of $k=0.035$ and therefore most likely to be contaminated by low $k$ foregrounds.
 This $k$ mode is well above the noise and has a redshift  dependence  much the same shape as the $T_{rms}$ curve in Figure \ref{fig:noise}, with a minimum near redshift 8.5 and a dramatic rise above redshift 10.
@@
 Though the points are all within the region predicted to be dominated by noise (grey) the variation between the redshifts and size of error bars suggests that a small residual signal is probably still present.
-The noise estimate incorporates a number of simplifications and could easily be off by a factor of 2-4.
+The theoretical noise estimate is likely to be off by a small amount (see figure \ref{fig:pspecs} and Section \ref{sec:pspecs}).
 \label{fig:slices}}
@@
 
-%\subsection{Future Work}
-%Further work is in progress to refine methods to include partially sampled data.
+\begin{deluxetable}{llrrrrr}
+\tablecolumns{7}
+\tablecaption{Power spectrum values}
+\tablehead{
+\colhead{k} & \colhead{redshift}  & \colhead{$\Delta^2$} &
+\colhead{bootstrap error } &\colhead{\parbox{6em}{noise model error} }& \colhead{\parbox{5em}{bootstrap significance}} & \colhead{\parbox{6em}{noise model significance}}\\
+
+\colhead{[hMpc$^{-1}$]} & \colhead{}  & \colhead{[mK$^2$]} &
+\colhead{ [mK$^2$]} &\colhead{\parbox{7em}{[mK$^2$]} }& \colhead{} & \colhead{}
+}
+\startdata
+0.1 & 10.29 & 43575.5512 & $\pm$ 5729.3347 & $\pm$ 8029.1131 & Det & Det \tabularnewline
+0.1 & 8.54 & 5701.6267 & $\pm$ 1168.7025 & $\pm$ 3671.887 & Det & Det \tabularnewline
+0.1 & 7.94 & 7695.373 & $\pm$ 2044.5824 & $\pm$ 3133.8433 & Det & Det \tabularnewline
+0.1 & 7.55 & 9461.8178 & $\pm$ 2761.5645 & $\pm$ 2393.1972 & Det & Det \tabularnewline
+0.2 & 10.29 & 11410.9746 & $\pm$ 5039.8271 & $\pm$ 27281.158 & Det & ULim \tabularnewline
+0.2 & 8.54 & 5855.8909 & $\pm$ 2195.8363 & $\pm$ 12275.7791 & Det & ULim \tabularnewline
+0.2 & 7.94 & 5143.6577 & $\pm$ 1799.7316 & $\pm$ 10221.3583 & Det & ULim \tabularnewline
+0.2 & 7.55 & 856.316 & $\pm$ 1490.2202 & $\pm$ 7636.2256 & ULim & ULim \tabularnewline
+0.3 & 10.29 & 34756.4662 & $\pm$ 16384.7719 & $\pm$ 64977.9446 & Det & ULim \tabularnewline
+0.3 & 8.54 & 12923.2987 & $\pm$ 5373.4928 & $\pm$ 29058.6627 & Det & ULim \tabularnewline
+0.3 & 7.94 & 5204.3757 & $\pm$ 4834.4927 & $\pm$ 24046.4004 & ULim & ULim \tabularnewline
+0.3 & 7.55 & 7628.4564 & $\pm$ 4810.3274 & $\pm$ 17762.1184 & Det & ULim \tabularnewline
+0.4 & 10.29 & 42909.9827 & $\pm$ 30695.0274 & $\pm$ 128355.9988 & ULim & ULim \tabularnewline
+0.4 & 8.54 & 24115.248 & $\pm$ 16585.6317 & $\pm$ 56571.7524 & Det & ULim \tabularnewline
+0.4 & 7.94 & 14730.7401 & $\pm$ 9930.2065 & $\pm$ 46725.4956 & Det & ULim \tabularnewline
+0.4 & 7.55 & -8601.8091 & $\pm$ 14191.4889 & $\pm$ 34339.7156 & ULim & ULim \tabularnewline
+\enddata
+\tablenotetext{a}{Det indicates a measurement and error-bar inconsistent with zero, ULim indicates consistency with zero at 2 $\sigma$.}
+\label{tab:data}
+\end{deluxetable}
 
@@
 
-With this paper we have extended the redshift coverage of \cite{Parsons:2013p9876} to cover the range $10.3>z>7.6$.
-Most of the power spectrum data points demonstrate a robust removal of foreground signals to the thermal noise limit, leaving systematics on the scale of the noise.
-Of particular note is the increased difficulty removing foregrounds at redshift 10.3.
-This appears to be largely due to the centrally weighted Blackman-Harris filter used to provide a high dynamic range delay transform in the broadband foreground filter.
-Future work will focus on more effectively removing foregrounds nearer to the edge of the band, opening more of the spectrum by being more tolerant of missing data, and better handling the removing of what residual foregrounds we do see with spectrum estimators which more robustly estimate power in the presence of residual foregrounds (e.g.
+Here we have extended the redshift coverage of the \cite{Parsons:2014p10499} PAPER results to cover the range $10.3>z>7.6$, when large scale reionization of hydrogen is thought to most likely have occurred.
+Most of the power spectrum data points demonstrate a removal of foreground signals to approximately the thermal noise limit, leaving systematics on the scale of the noise.
+Noise limited bins --those which are statistically consistent with zero and marked as ULim in Table \ref{tab:data}-- occur at several $k$ modes at redshifts 7.55, 7.94 and 10.29.
+The deepest point --at redshift 7.55 and $k=0.2$-- is a 2$\sigma$ upper limit of (38mK)$^2$.
+For comparison, the \citet{Lidz:2008p8251} fiducial model of the reionization power spectrum estimates an amplitude of (20mK)$^2$, for models with 50\% reionization at $z=$7.55.
+
+The observations reported here demonstrate a nearly noise limited integration over a full season but using only a quarter of final design baseline count.
+These  32 antenna observations from 2011 were followed by a deployment of 64 antennae in 2012 and a 128 in 2013.
+A second season of observing with the 128 antenna array is now under way.
+Because the additional baselines are redundant, they will add coherently; the factor of 4 increase in antenna count is expected to drop the sensitivity limit by a factor of 4 in (mK)$^2$.
+
+%(iii) third paragraph would then be on lowest freq bin and methods for improvement in analysis
+
+Surveying the effectiveness of foreground filtering across the band we note the increased difficulty in removing foregrounds at redshift 10.3.
+This appears to be largely due to the centrally-weighted Blackman-Harris filter used to provide a high dynamic range delay transform in the broadband foreground filter.
+This filter provides an extremely sharp edge and high leakage rejection to the delay spectrum at the horizon, but at the cost of down-weighting data in the outer quarters of the band.
+Future work will focus on more effectively removing foregrounds nearer to the edge of the band, opening more of the spectrum by being more tolerant of missing data, and better handling the removing of what residual foregrounds with spectrum estimators that more robustly estimate power in the presence of residual foregrounds (e.g.
 \cite{Liu:2011p8763}).

### 2014-07-02 ms.tex
adding in Jonnies conclustion comments and a second read through. getting ready to send v2 to the collab

@@
 Power spectra at different points across the redshift range reveal the variable efficacy of this subtraction.
-Noise limited measurements of $\Delta$ at $k=$0.2hMpc$^{-1}$,z$=7.5$ reach as low as 58mK ($2-\sigma$).
+Noise limited measurements of $\Delta^2$ at $k=$0.2hMpc$^{-1}$ and z$=7.5$ reach as low as (48mK)$^2$ ($2\sigma$).
 Most spectra demonstrate a robust foreground removal of foreground signals to the thermal noise limit, leaving systematics on the scale of the noise.
@@
 Therefore, while a detection of the 21 cm signal at even a single frequency would be a tremendous breakthrough, analysis techniques must be developed to capitalize on the wide bandwidths of the current generation of high-redshift 21cm telescopes.
-Using the same data set as P14, this paper presents improved upper limits on the HI power spectrum over the redshift range $10.5>z>7.2$.
+Using the same data set as P14, this paper presents improved upper limits on the HI power spectrum over the redshift range $10.7>z>7.2$.
 In Section  \ref{sec:observations} we summarize the observations, in Section \ref{sec:obs_meth} review the reduction methodology, we present the new upper limits in \ref{sec:results}, and in Section \ref{sec:conclusion} we offer conclusions and discussion of future work.
@@
 The work here follows the same basic procedure and uses the same underlying data set as P14.
-Here we provide a quick summary and refer the reader to P14 for a more in depth discussion.
-A general overview of the PAPER system can be found in \cite{Parsons:2010p6757}, calibration of the primary beam in \cite{Pober:2012p8800}, and imaging results in \cite{Jacobs:2011p8438,jacobs:2013b,Stefan:2013p9926}.
+Here we provide a quick summary and refer the reader to P14 for a more in-depth discussion.
+A general overview of the PAPER system can be found in \cite{Parsons:2010p6757}, calibration of the primary beam in \cite{Pober:2012p8800}, and imaging results in \cite{Jacobs:2011p8438,jacobs:2013b} and \citet{Stefan:2013p9926}.
 Sensitivity analysis described in \cite{Parsons:2012p9028} revealed that for the low gain elements employed by PAPER, a highly redundant ``grid'' type arrangement offers a significant sensitivity boost.
@@
 \label{sec:obs_meth}
-Here we summarize our data reduction steps; for detailed descriptions see Section 3 of P14.
+Here we summarize our data reduction steps; for more details see Section 3 of P14.
 In summary, we use 70 nearly identical baselines to make a 1D estimate along the spectral or line-of-sight direction of the reionization era HI power spectrum.
@@
 In the final cross-multiplication step, the last layer of systematics is estimated and removed by projecting non-physical correlations between baselines.
+\subsection{Delay and Fringe rate transforms or The Fourier Furor}
+\label{sec:transforms}
+In several stages throughout the analysis process we take a 2D Fourier transform of the visibility spectra $V(\nu,t)$ into ``delay/fringe rate'' space where delay is the Fourier dual to frequency and similarly fringe-rate for time.
+In this space, smooth spectrum sources are physically localized to delays shorter than the light travel time length of the baseline and fringe rates shorter than the angular sidereal rate perpendicular to the baseline vector.
+Sources at the horizon, in the direction of the baseline vector, have the longest delays, while fringe rates are highest where the celestial equator crosses the horizon.
 
-In several stages throughout the analysis process we take a 2D Fourier transform of the visibility spectra $V(\nu,t)$ into ``delay/fringe rate'' space where delay is the Fourier dual to frequency and similarly fringe-rate for time.
-In this space, smooth spectrum sources are physically localized to delays shorter than the light travel time length of the baseline and fringe rates shorter than the sidereal rate.
-Sources at the horizon, in the direction of the baseline vector, have the longest delays, while fringe rates are highest where the celestial equator crosses the horizon.
-%TODO is that right?
  In this Fourier space, sources are highly localized with deviations from a flat spectrum manifesting as a slight dispersion.
@@
 If enough data is missing this PSF can cause smooth-spectrum sources to leak beyond the horizon.
-To account for this, we use a CLEAN like, iterative, peak-finder and subtraction algorithm which is limited to finding peaks within the physically allowable ranges of delay and/or fringe-rate In this case, the 1D ``CLEAN'' beam is the Fourier transform of the spectral or time sampling function \citep{Parsons:2009p7859}.
+To account for this, we use a CLEAN like, iterative, peak-finder and subtraction algorithm which is limited to finding peaks within the physically allowable ranges of delay and/or fringe-rate.
+In this case, the 1D ``CLEAN'' beam is the Fourier transform of the spectral or time sampling function \citep{Parsons:2009p7859}.
 
-The data analysis pipeline essentially consists of iterative application of the delay/fringe rate transform process, with an ever tightening allowable number of modes, interleaved with stages of averaging (time, frequency, night), before finally computing a power spectrum.
+The data analysis pipeline essentially consists of iterative application of the delay or fringe-rate transform process, with an ever tightening allowable number of modes, interleaved with stages of averaging (time, frequency, night), before finally computing a power spectrum.
 This final step takes advantage of the redundant baselines to make an unbiased power spectrum estimate by cross-multiplying identical baselines and then averaging the power spectrum modes.
-Thus as we step through the various iterations of delay spectrum we have the ability to isolate and subtract foregrounds in the native domain of the measured power spectrum.
-Note that isolation of foregrounds to a region below line-of-sight $k_\parallel$s that increases with baseline length is also the much discussed ``wedge'' \cite{Liu:2014p10462,Liu:2014p10463,Thyagarajan:2013p10039,Pober:2013p9942,Trott:2012p10466,Morales:2012p8790,Parsons:2012p8896,Vedantham:2012p10297,Datta:2010p8781,Parsons:2009p7859}.
+By not combining baselines until the last step and by assuming that the line of site transform is well approximated by the delay transform we dramatically simplify the analysis procedure to a series of signal processing steps.
+It also obviates the need for precision sky and instrumental beam models which are required by imaging arrays.
+
+This isolation of foregrounds to a region below a line-of-sight $k_\parallel$ mode that increases with baseline length is also the much discussed ``wedge'' \cite{Liu:2014p10462,Liu:2014p10463,Thyagarajan:2013p10039,Pober:2013p9942,Trott:2012p10466,Morales:2012p8790,Parsons:2012p8896,Vedantham:2012p10297,Datta:2010p8781,Parsons:2009p7859}.
 As we only have a single length of baseline the wedge manifests as a single value of $k\sim k_\parallel$, below which foregrounds are expected to dominate.
@@
 The redshift bins for which we have computed power spectra --shown in Figure \ref{fig:bins}-- have been selected from the available bandwidth using two criteria: minimizing covariance between redshifts and avoiding missing data.
-We minimize covariance between adjacent redshift bins by limiting overlap of each bin to the spectrum at the outer half of the bin channel range which, as will be described in section \ref{sec:power_spectrum} is significantly down-weighted by the use of a Blackman-Harris window.
-Avoiding missing data is important because the power spectrum method, developed in P14, which leverages the redundant baselines to estimate non-sky covariance and project out these contaminated modes is particularly sensitive to missing data -the inverse of the covariance is not defined- so we also select only channels that have no missing data over the entire sidereal period (see the dotted line in \ref{fig:bins}).
-With these constraints we arrive at the redshifts 10.3, 8.5,7.9 and 7.5.
+We minimize covariance between adjacent redshift bins by limiting overlap of each bin to the spectrum at the outer half of the bin channel range which, as will be described in section \ref{sec:power_spectrum}, is significantly down-weighted by the use of a Blackman-Harris window.
+Avoiding missing data is important because the power spectrum method, developed in P14, which leverages the redundant baselines to estimate non-sky covariance and project out these contaminated modes is particularly sensitive to missing data -the inverse of the covariance is not defined- so we also select only channels that have no missing data over the entire sidereal period (see the dotted line in Figure \ref{fig:bins}).
+With these constraints we arrive at the redshifts 10.3, 8.5,7.9 and 7.45.
 For the purposes of comparison with P14 we also include the redshift 7.68 bin.
-\citet{Dillon:2013p10497}describes a pseudo-inverse method for handling the covariance introduced by missing data.
-Though not implemented here, we will use a version of their method in future analysis.
+\citet{Dillon:2013p10497} describes a pseudo-inverse method for handling the covariance introduced by missing data.
+Though not implemented here, this method is in the process of being adapted for use in highly redundant analysis \citep{Liu:2014p10462,Liu:2014p10463}.
 
@@
 However, by itself the redundant calibration does not contain enough phase information to phase coherently to a sky location, there remain two free phase parameters which cannot be solved by redundancy alone.
-We find these by fitting a model of Pictor A, Fornax A, and the Crab Nebula during a time when the sky is dominated by these three sources while marginalizing over the unknown\footnote{Unknown in the sense of a joint uncertainty in source flux and primary beam pattern.} apparent flux ratio between the three sources.
-With the delays in place we are now able form a beam on Pictor A and (for each channel) set the overall amplitude to the calibration value of 381Jy found in \cite{jacobs:2013b}.
+We find these by fitting a model of Pictor A, Fornax A, and the Crab Nebula during a time when the sky is dominated by these three sources while marginalizing over the unknown\footnote{Unknown in the sense of a joint uncertainty in source flux and primary beam pattern.
+Though the fluxes of these sources is in some cases fairly well constrained, fluxes at 150MHz are still fairly uncertain as is the PAPER beam in those directions.} apparent flux ratio between the three sources.
+With the delays in place we are now able form a beam on Pictor A and (for each channel) set the overall amplitude to the calibration value of 382 (f/150MHz)$^{-0.77}$Jy found in \cite{jacobs:2013b}.
 
@@
 This removes excess correlation due to cross-talk in the analog signal chain.
-The residuals are then flagged once more for RFI before the 92 nights of data are averaged into 36.4 second long local sidereal time (LST) bins, which as PAPER is a drift-scanning instrument, are equivalent to bins in Right Ascension/ Declination.
+The residuals are then flagged once more for RFI before the 92 nights of data are averaged into 36.4 second long local sidereal time (LST) bins, which as PAPER is a drift-scanning instrument, are equivalent to bins in Right Ascension (Declination is fixed at -30\arcdeg).
 During averaging we found that some LST bins were dominated by a small number of exceedingly bright samples lying well outside the rest of the gaussian distributed data.
 To compensate we filter the 10\% brightest samples in each bin.
-The source of these outliers is not known,  a likely possibility being an instability in the analog signal chain stimulated by weather or bright interference, a circumstance that has since been observed in later seasons.
+The source of these outliers is not known,  a likely possibility is an instability in the analog signal chain stimulated by weather or bright interference, a circumstance that has since been observed in later seasons.
 
@@
 Performing a final fringe-rate filter, limiting to fringe-rates expected on a 30m baseline (down from 300m in the last iteration), we arrive at a data-set averaged to 789s, the maximum possible while still maintaining fringe coherence.
-The root mean square of the residual signal (seen in Figure \ref{fig:noise}) at the end of this process is close to the 3mK level expected given the total integration time.
+The root mean square of the residual signal (seen in Figure \ref{fig:noise}) at the end of this process is close to the 3mK level expected given the total integration time and  system temperature.
 
@@
 This leaves us with 40 delay samples on each of 70 baselines which are divided into three redundant groups.
-Within these groups we compute the  correlation between each mode.
-The cross correlation of the same delay modes between different redundant baselines provides our estimate of the power spectrum.
-Cross correlation between different delay modes provides our measure of systematic bias.
-Sky-like correlations should be identical between all redundant baselines to within the level of the noise.
-In practice some correlations deviate significantly from the median.
+Within these groups we cross correlate delay spectra between different redundant baselines.
+The cross multiplication of the same delay modes between different redundant baselines provides an unbiased estimate of the power spectrum.
+These "sky-like" correlations should be identical between all redundant baselines to within the level of the noise, while all other cross multiplications between delay modes should not be correlated between different baselines.
+In practice  these non-sky-like modes do occasionally have significant power which leaks into the correlations which sample the power spectrum.
 These are removed by iteratively dividing the covariance into a model of systematics and a model of sky-like emission and then projecting out large residual modes.
@@
 \caption{\label{fig:P14compare} The redshift 7.68 bin has been reprocessed for comparison with the P14 result.
+Compare with figure 6 of P14.
 Note that the LST range processed here is 3 hours less than P14 for a small reduction in sensitivity but large reduction in foreground.
@@
 
+%\[
+%\frac{1}{\sqrt{N_{eff}}} = \frac{1}{N_{\textrm{lsts}}} \sum_\textrm{lst=1}^\textrm{lst=9}{\frac{1}{\sqrt{N(lst)}}}
+%\]
 \[
-\frac{1}{\sqrt{N_{eff}}} = \frac{1}{N_{\textrm{lsts}}} \sum_\textrm{lst=1}^\textrm{lst=9}{\frac{1}{\sqrt{N(lst)}}}
+\frac{1}{\sqrt{N_{eff}}}  = \left\langle \frac{1}{\sqrt{N(lst)}}  \right\rangle_{\textrm{lst}}
 \]
-
-On average each bin is in this dataset is effectively measured 39.6 times. Given this accounting of the number of samples $N(lst)$, we estimate the final r.m.s. temperature to be
+On average each bin is in this dataset is effectively measured 39.6 times. Given this accounting of the number of samples $N(lst)$, we estimate the r.m.s. temperature of the averaged data set to be
 
@@
 
-where B and t are the 496kHz channel width and lst bin length, respectively and the approximation indicates that we have assumed average values for the number of samples and the time dependent sky component of the system temperature.
+\noindent{where B and t are the 496kHz channel width and lst bin length, respectively and the approximation indicates that we have assumed average values for the number of samples and the time dependent sky component of the system temperature.
 In P14 the total system temperature ($T_{sky} + T_{rcvr}$) was estimated to be 550K using 20 nights of the 92 night set which covered LSTs 3-9.
 Using the global sky model  \citep{deOliveiraCosta:2008p2242} and a model of the PAPER primary beam we estimate the mean $T_{sky}$ during that 20 night period to be 250K, implying a receiver temperature of 300K.
-The mean temperature over the full 92 night period is only 4K higher then the mean from used to make the 550K estimate in P14.
-The P14 estimate of 550K for the system temperature is a good estimate for the full observing period.
-$T_{rms}$ using 550K and an effective sample count of 39.6  is plotted as a dotted line in Figure \ref{fig:noise} for before and after fringe rate filtering (t=40s and 789s)
+The mean temperature over the full 92 night period is only 4K higher; the estimate of 550K for the system temperature is a good estimate for the full observing period.
+$T_{rms}$ using 550K and an effective sample count of 39.6  is plotted as a dotted line in Figure \ref{fig:noise} for before and after fringe rate filtering (t=40s and 789s)}
 
@@
 
-In  Figure \ref{fig:pspecs} we show the  power spectra at different redshift bins --showing both the spherically averaged power spectrum ($P(k)$, left) and the r.m.s.
-fluctuation per $k$ interval where $\Delta^2(k)\def\frac{k^3}{2\pi^2}P(k)$.In figure \ref{fig:slices} we see different $k$ mode slices as a function of redshift,  all plotted with $2\sigma$ error bars and in Table \ref{tab:data} we list the data plotted in these slices.
-The theoretical noise power spectrum is estimated using the method described in\cite{Pober:2013p9581}, assuming a system temperature of 550K and the observing scheme described in Section \ref{sec:observations}.
+In  Figure \ref{fig:pspecs} we show the  power spectra at different redshift bins --showing both the spherically averaged power spectrum $P(k)$ and the volume independent  $\Delta^2(k)\equiv\frac{k^3}{2\pi^2}P(k)$.
+The $P(k)$ spectrum averages the three baseline types but preserves the positive and negative delays.
+In figure \ref{fig:slices} we see different $k$ mode slices as a function of redshift,  all plotted with $2\sigma$ error bars and in Table \ref{tab:data} we list the data plotted in these slices.
+
+The theoretical noise power spectrum (dashed line in figures \ref{fig:pspecs} and \ref{fig:P14compare} and grey region in figure \ref{fig:slices}) is estimated using the method described in \cite{Pober:2013p9581}, assuming a system temperature of 550K and the observing scheme described in Section \ref{sec:observations}.
+In all figures the noise levels are plotted for comparison with the power spectrum values rather than the error bars, i.e.
+65\% of uncorrelated points should lie below the line or within the grey region.
+In general the theoretical error bars are larger than the bootstrap errors.
+This is probably most clear in \ref{fig:slices} where, though most points are inside the error region, the bootstrap errors are only consistent with zero 25\% of the time.
+In general it would appear that the sensitivity calculations under-estimate the sensitivity of the measurements by a small factor.
+See \cite{Pober:2014p10390,Pober:2013p9581} for a discussion of the approximations made in those calculations.
+
 Generally, both figures convey a picture of a power spectrum dominated by noise, with foregrounds encroaching from inside and just outside the horizon to higher $k$ ``reionization'' modes.
-
 One easy way to see this leakage is by comparing the bootstrap error bar, the theoretical error and the power spectrum value.
 At the level of noise here, we expect all points outside the horizon to be consistent with zero.
-As a simple test we divide the points into ``Detections'' and ``Upper Limits'' (Det/ULim) in Table \ref{tab:data} for both kinds of error bar.
+As a simple sorting criteria we divide the points into ``Detections'' and ``Upper Limits'' (Det/ULim) in Table \ref{tab:data} for both theoretical and bootstrap error bar.
 As we can see, the larger theoretical error bars are mostly consistent with zero, while the same is true of a little less then half of the bootstrap error bars.
@@
 we found in Figure \ref{fig:noise}.
-Recall that the wide-band foreground model we have subtracted  (which causes the notch of missing power near $k_\parallel=0$ in Figure \ref{fig:pspecs})\footnote{except at z=7.68 where the foregrounds have been left in for comparison} is the same for all redshift slices and was built by weighting the entire band by a Blackman-Harris window function.
-Using this window, contributions to the foreground model from 126MHz ($z=$10.3) are down-weighted by $\sim$75\%.
-If the 126MHz sky or systematics are much different then 160Mhz, the foreground filter will be less effective.
+Recall that the wide-band foreground model we have subtracted  --which causes the notch of missing power\footnote{Note that this notch is not present in figure \ref{fig:P14compare} because the foregrounds have been left in for comparison with P14.} near $k_\parallel=0$ in Figure \ref{fig:pspecs}-- is the same for all redshift slices and was built by weighting the entire band by a Blackman-Harris window function.
+Using this window, contributions to the foreground model from 126MHz ($z=$10.3) are down-weighted $\sim$75\% compared to 150Mhz ($z=$8.47).
+If the 126MHz sky or systematics are much different then 150Mhz foreground residuals at that redshift will be higher.
 The $k=0.1$hMpc$^{-1}$ slice (upper left of figure \ref{fig:slices}) is only a factor of $\sim$three away from the light travel horizon of $k=0.035$ and therefore most likely to be contaminated by low $k$ foregrounds.
 This $k$ mode is well above the noise and has a redshift  dependence  much the same shape as the $T_{rms}$ curve in Figure \ref{fig:noise}, with a minimum near redshift 8.5 and a dramatic rise above redshift 10.
-
-%Across the entire band some $k$ modes appear to be dominated somewhat by non-noise systematics, these are not always clearly foregrounds because they appear at only certain $k$ modes rather than clearly emanating from small delays like in the z=10.3 power spectrum.
-
-%Possible assessment of comparison to noise.  Opt/Mid/Pes noise estimates.  If Opt noise => signal, if Pes=>systematics
 
@@
 
-Here we have extended the redshift coverage of the \cite{Parsons:2014p10499} PAPER results to cover the range $10.3>z>7.6$, when large scale reionization of hydrogen is thought to most likely have occurred.
+The power of the highly redshifted 21 cm as a cosmic probe lies in its ability to probe a 3D volume by observing at different frequencies.
+The present analysis extends the work of  \cite{Parsons:2014p10499}, which used a single-baseline delay spectrum analysis to place an upper limit on the 21 cm power at z = 7.68 (163 MHz).
+In this work, we take advantage of the wide bandwidth of the PAPER instrument to place limits on the power spectrum at a range of redshifts $10.7>z>7.2$, demonstrating that the algorithms used in P14 can exploit the full potential of the 21 cm line.
+
 Most of the power spectrum data points demonstrate a removal of foreground signals to approximately the thermal noise limit, leaving systematics on the scale of the noise.
 Noise limited bins --those which are statistically consistent with zero and marked as ULim in Table \ref{tab:data}-- occur at several $k$ modes at redshifts 7.55, 7.94 and 10.29.
-The deepest point --at redshift 7.55 and $k=0.2$-- is a 2$\sigma$ upper limit of (38mK)$^2$.
-For comparison, the \citet{Lidz:2008p8251} fiducial model of the reionization power spectrum estimates an amplitude of (20mK)$^2$, for models with 50\% reionization at $z=$7.55.
+The deepest point --at redshift 7.55 and $k=0.2$-- is a 2$\sigma$ upper limit of (48mK)$^2$.
+For comparison, the \citet{Lidz:2008p8251} fiducial model of the reionization power spectrum estimates an amplitude of (4.4mK)$^2$, for models with 50\% reionization at $z=$7.55.
 
-The observations reported here demonstrate a nearly noise limited integration over a full season but using only a quarter of final design baseline count.
-These  32 antenna observations from 2011 were followed by a deployment of 64 antennae in 2012 and a 128 in 2013.
+The fact that the best upper limit still comes from the z = 7.6 band does illustrate that this redshift corresponds to a somewhat special frequency for the PAPER instrument, where the combined contribution of system noise, RFI signals, and residual foregrounds (if any remain) are at a minimum.
+
+Further work will look to expand on this result on several fronts.
+Alternative algorithms to the wide-band CLEAN could potentially better remove the delay-space covariance introduced by RFI flags in frequency and are under investigation.
+Another active point of investigation is aimed at alternative windows to the Blackman-Harris used in the wide-band CLEAN (or wholly distinct algorithms) that could better remove foregrounds from the edges of the band while still limiting foreground bleed into the EoR window.
+
+Future work will also include observations from subsequent seasons with more antenna.
+The observations reported here demonstrate a nearly noise limited integration over a full season but used only a quarter of the final design antenna count.
+These  32 antenna observations from 2011 were followed by a full season of 64 antennae in 2012 and 128 in 2013.
 A second season of observing with the 128 antenna array is now under way.
-Because the additional baselines are redundant, they will add coherently; the factor of 4 increase in antenna count is expected to drop the sensitivity limit by a factor of 4 in (mK)$^2$.
+In going from  32 to 128 antenna the mK$^2$ sensitivity increases by the expected factor of 4 and by an additional factor of $\sim$2 after accounting for the substantial uv-plane redundancy \citep{Parsons:2012p9028}.
+
+The analysis of the system temperature presented in section 4.1 is one of the most thorough such investigations applied to the PAPER instrument.
+It confirms the effectiveness of the wideband CLEAN in removing foregrounds over a wide range of the band, while illustrating weaknesses in this approach in pushing towards the edges of PAPER's frequency range.
+Furthermore, after the application of the fringe rate-based time averaging, we see hints that baseline-to-baseline variations between redundantly spaced baselines is becoming the dominant systematic in the analysis.
+Techniques like the covariance removal introduced in P14 and also used in this work can differentiate some of this variation from the sky signal; however some combination of improved calibration techniques \citep{Zheng:2014p10467}, model building \citep{Sullivan:2012p9457}, and a better understanding of error covariance  \citep{Liu:2011p8763} may also prove valuable in reducing this systematic.
 
 %(iii) third paragraph would then be on lowest freq bin and methods for improvement in analysis
-
-Surveying the effectiveness of foreground filtering across the band we note the increased difficulty in removing foregrounds at redshift 10.3.
-This appears to be largely due to the centrally-weighted Blackman-Harris filter used to provide a high dynamic range delay transform in the broadband foreground filter.
-This filter provides an extremely sharp edge and high leakage rejection to the delay spectrum at the horizon, but at the cost of down-weighting data in the outer quarters of the band.
-Future work will focus on more effectively removing foregrounds nearer to the edge of the band, opening more of the spectrum by being more tolerant of missing data, and better handling the removing of what residual foregrounds with spectrum estimators that more robustly estimate power in the presence of residual foregrounds (e.g.
-\cite{Liu:2011p8763}).
 

### 2014-07-08 ms.tex
fixed a bug in the plotting of noise curves. things seem to make a lot more sense now

@@
 \caption{ \label{fig:pspecs} PAPER power spectra at four redshifts.
-On the top, $P_k$ spectra (a simple units change from the raw delay spectrum) provide a useful diagnostic on foreground rejection, while on the bottom we plot in $\Delta^2$ cosmological units.
+On the top, $P_k$ spectra (a simple units change from the raw delay spectrum) provide a useful diagnostic on foreground rejection, while on the bottom we plot in $\Delta^2$ cosmological units, error bars are 2$\sigma$.
 Redshift increases left to right.
@@
 The noise curve (dashes) is calculated using the method described in \cite{Pober:2013p9581}  and indicates the 1$\sigma$ confidence bounds on data points consisting purely of noise; 65\% of uncorrelated noise like data points will lie below the curve.
-However, note the caveats that due to the weighting of the delay transform, adjacent $k$ bins are 50\% correlated.
-It is also clear that the sensitivity calculations under-estimate the sensitivity of the measurements by a small factor.
+However, note the caveat that due to the weighting of the delay transform, adjacent $k$ bins are 50\% correlated.
 See \cite{Pober:2014p10390,Pober:2013p9581} for a discussion of the approximations made in those calculations.
@@
 \caption{\label{fig:P14compare} The redshift 7.68 bin has been reprocessed for comparison with the P14 result.
+See figure \ref{fig:pspecs} for a detailed description of curves.
 Compare with figure 6 of P14.

### 2014-07-30 ms.tex
on the verge of submission

@@
-%\documentclass[10pt,iop]{emulateapj}
-\documentclass[preprint]{aastex}
+%\documentclass[fleqn,10pt,iop]{emulateapj}
+\documentclass{aastex}
 
@@
 \usepackage{xcolor}
+
 %\citestyle{aa}
@@
 
-\shorttitle{EoR time}
+\shorttitle{PAPER Power Spectrum over Cosmic Time}
 \shortauthors{Jacobs et al.}
@@
 David F. Moore\altaffilmark{3},
-Jonathan C. Pober\altaffilmark{2},
+Jonathan C. Pober\altaffilmark{12},
 Irina I. Stefan\altaffilmark{10},
@@
 \altaffiltext{9}{Square Kilometer Array, South Africa Project, Cape Town, South Africa}
-\altaffiltext{10}{Cavendish Lab., Cambridge, UK}
-\altaffiltext{11}{Berkeley Center for Cosmological Physics, UC Berkeley, Berkeley,
-CA}
+\altaffiltext{10}{Cavendish Astrophysics Group, University of Cambridge, Cambridge, UK}
+\altaffiltext{11}{Berkeley Center for Cosmological Physics, UC Berkeley, Berkeley,CA}
+\altaffiltext{12}{Dept. of Physics, University of Washington, Seattle, WA}
 
@@
 The primary obstacle to this goal are bright foregrounds and the attendant systematics associated with high dynamic range measurements.
-Here we use filtering techniques that take advantage of the large instrumental bandwidth to make a high dynamic range subtraction of foreground power.
-Power spectra at different points across the redshift range reveal the variable efficacy of this subtraction.
-Noise limited measurements of $\Delta^2$ at $k=$0.2hMpc$^{-1}$ and z$=7.5$ reach as low as (48mK)$^2$ ($2\sigma$).
+Here we present an integration across a full season of data and find systematics just beginning to dominate, though well below any levels previously reported.
+This  500 hour integration is the longest such yet recorded and tests foreground removal to a theoretical dynamic range of 10,000 to 1.
+We use filtering techniques that take advantage of the large instrumental bandwidth to make a high dynamic range subtraction of foreground power by averaging together many low sensitivity measurements.
+.
+ Power spectra at different points across the redshift range reveal the variable efficacy of the foreground subtraction.
+Noise limited measurements of $\Delta^2$ at $k=$0.2hMpc$^{-1}$ and z$=7.55$ reach as low as (48mK)$^2$ ($1\sigma$).
 Most spectra demonstrate a robust foreground removal of foreground signals to the thermal noise limit, leaving systematics on the scale of the noise.
@@
 \section{Introduction}
-The Epoch of Reionization, when the first luminous objects ionized the pervasive cosmological hydrogen is predicted to be observable in highly redshifted 21 cm radiation.
+The Epoch of Reionization, when the first luminous objects ionized the pervasive cosmological hydrogen, is predicted to be observable in highly redshifted 21 cm radiation.
 The Donald C.
@@
 
-PAPER is located in the Karoo desert at the site of the South African portion of the future Square Kilometer Array\footnote{\url{skatelescope.org}} and has doubled in size on a yearly basis since 2009, and science-grade observations have been made with each stage of the build-out.
+PAPER is located in the Karoo desert at the site of the South African portion of the future Square Kilometer Array\footnote{\url{skatelescope.org}} and has doubled in size on a yearly basis since 2009; science-grade observations have been made with each stage of the build-out.
 
 Here we report on deep integrations made with a 32 element array in 2011, first described in \cite{Parsons:2014p10499}, hereafter P14.
-Our data reduction method was described in detail in P14, and used this pipeline to give the deepest yet limits on the HI power spectrum in the presence of bright foregrounds at redshift 7.68 using 92 nights (1800 hours) of data.
-The Epoch of Reionization signal is expected to evolve strongly with redshift, and it is this variation with cosmic history that will allow us to begin to place constraints on the physics of reionization \citep{Pritchard:2008p8123,Pober:2014p10390}.
+Our data reduction pipeline was described in detail in P14, where the same methods were used to give the deepest limits yet on the HI power spectrum in the presence of bright foregrounds at redshift 7.68.
+However, the Epoch of Reionization signal is expected to evolve strongly with redshift.
+In fact, it is this signature variation which will distinguish it from foregrounds and it is this variation with cosmic history that will allow us to begin to place constraints on the physics of reionization \citep{Pritchard:2008p8123,Pober:2014p10390}.
 Therefore, while a detection of the 21 cm signal at even a single frequency would be a tremendous breakthrough, analysis techniques must be developed to capitalize on the wide bandwidths of the current generation of high-redshift 21cm telescopes.
 Using the same data set as P14, this paper presents improved upper limits on the HI power spectrum over the redshift range $10.7>z>7.2$.
-In Section  \ref{sec:observations} we summarize the observations, in Section \ref{sec:obs_meth} review the reduction methodology, we present the new upper limits in \ref{sec:results}, and in Section \ref{sec:conclusion} we offer conclusions and discussion of future work.
+In Section  \ref{sec:observations} we summarize the observations, in Section \ref{sec:obs_meth} review the reduction methodology, we present the new upper limits in Section \ref{sec:results}, and in Section \ref{sec:conclusion} we offer conclusions and discussion of future work.
 
@@
 A general overview of the PAPER system can be found in \cite{Parsons:2010p6757}, calibration of the primary beam in \cite{Pober:2012p8800}, and imaging results in \cite{Jacobs:2011p8438,jacobs:2013b} and \citet{Stefan:2013p9926}.
-Sensitivity analysis described in \cite{Parsons:2012p9028} revealed that for the low gain elements employed by PAPER, a highly redundant ``grid'' type arrangement offers a significant sensitivity boost.
-In most interferometers the locations of the antennae are optimized such that each baseline samples a different Fourier mode of the sky; this is the ideal case for reconstructing images where each mode contains different information.
+Sensitivity analysis described in \cite{Parsons:2012p9028} revealed that for the low gain elements employed by PAPER, a highly redundant ``grid'' type arrangement offers a significant sensitivity boost extending the performance of the compact designs motivated by \citet{Morales:2004p2494,Bowman:2006p1887,Lidz:2008p8251}.
+In most interferometers the locations of the antennas are optimized such that each baseline samples a different Fourier mode of the sky; this is the ideal case for reconstructing images where each mode contains different information.
 For a power spectrum measurement the key metric is sensitivity per mode, rather than number of modes.
@@
 This selection includes those between adjacent columns and within at least one row of each other, a selection containing 70 $\sim$30m-long baselines.
-We will use these baselines to make a one dimensional estimate of the HI power spectrum.
+We will use these baselines to make a one dimensional estimate --spectral line of sight only-- of the HI power spectrum.
 
-Observations spanning the band between 100 to 200-MHz ($13.1>z>6.1$) were recorded at a resolution of 48kHz and 10.7s  beginning Dec 7, 2011 and ending March 19, 2012 (with some down-time for maintenance) giving a total of 92 nights or 1800 hours.
+Observations spanning the band between 100 to 200-MHz ($13.1>z>6.1$) were recorded at a resolution of 48kHz and 10.7s  beginning Dec 7, 2011 and ending March 19, 2012 (with some down-time for maintenance) giving a total of 513 hours over 92 nights.
 Within this set we included observations in the LST range 1h - 9h where the sky dominated system temperature is at a minimum.
-Note that this LST range is slightly shorter than in P14 which extended to LST of 12 hours.
+Note that this LST range is slightly shorter than in P14, which extended to LST of 12 hours.
 These last three hours were found to contribute minimally to increasing sensitivity while being dominated by bright galactic foreground emission and so have been excluded here.
@@
 In the final cross-multiplication step, the last layer of systematics is estimated and removed by projecting non-physical correlations between baselines.
-\subsection{Delay and Fringe rate transforms or The Fourier Furor}
+\subsection{Delay/Fringe-Rate Filtering: Averaging and Foreground Removal}
 \label{sec:transforms}
 In several stages throughout the analysis process we take a 2D Fourier transform of the visibility spectra $V(\nu,t)$ into ``delay/fringe rate'' space where delay is the Fourier dual to frequency and similarly fringe-rate for time.
-In this space, smooth spectrum sources are physically localized to delays shorter than the light travel time length of the baseline and fringe rates shorter than the angular sidereal rate perpendicular to the baseline vector.
+In this space, smooth spectrum sources are physically localized to delays shorter than the light travel time length of the baseline and fringe rates shorter than the sidereal rotation rate of the tip of the east-west component of the baseline vector.
 Sources at the horizon, in the direction of the baseline vector, have the longest delays, while fringe rates are highest where the celestial equator crosses the horizon.
@@
 The spectrum sampling function, which is uneven due to flagging of interference takes the form of a convolution by a point-spread-function (PSF) in the same way an imperfect sampling of the $uv$ plane gives rise to the angular PSF of an interferometer.
-If enough data is missing this PSF can cause smooth-spectrum sources to leak beyond the horizon.
+If enough data is missing this PSF can cause smooth-spectrum sources to leak beyond the ``horizon'', the light travel time limit.
 To account for this, we use a CLEAN like, iterative, peak-finder and subtraction algorithm which is limited to finding peaks within the physically allowable ranges of delay and/or fringe-rate.
@@
 This final step takes advantage of the redundant baselines to make an unbiased power spectrum estimate by cross-multiplying identical baselines and then averaging the power spectrum modes.
-By not combining baselines until the last step and by assuming that the line of site transform is well approximated by the delay transform we dramatically simplify the analysis procedure to a series of signal processing steps.
+By not combining baselines until the last step and by assuming that the line of sight transform is well approximated by the delay transform we dramatically simplify the analysis procedure to a series of signal processing steps.
 It also obviates the need for precision sky and instrumental beam models which are required by imaging arrays.
 
-This isolation of foregrounds to a region below a line-of-sight $k_\parallel$ mode that increases with baseline length is also the much discussed ``wedge'' \cite{Liu:2014p10462,Liu:2014p10463,Thyagarajan:2013p10039,Pober:2013p9942,Trott:2012p10466,Morales:2012p8790,Parsons:2012p8896,Vedantham:2012p10297,Datta:2010p8781,Parsons:2009p7859}.
-As we only have a single length of baseline the wedge manifests as a single value of $k\sim k_\parallel$, below which foregrounds are expected to dominate.
-The details of each application of the delay/fringe-rate transform will be laid out the following sections as we provide a brief walk-through of the processing pipeline.
+This isolation of foregrounds to a region below a line-of-sight $k_\parallel$ mode that increases with baseline length is also the much discussed ``wedge'' \citep{PhysRevD.90.023018,PhysRevD.90.023019,Thyagarajan:2013p10039,Pober:2013p9942,Trott:2012p10466,Morales:2012p8790,Parsons:2012p8896,Vedantham:2012p10297,Datta:2010p8781,Parsons:2009p7859}.
+As we only have three baseline types of nearly the same length the wedge manifests as a single value of $k\sim k_\parallel$, below which foregrounds are expected to dominate.
+The details of each application of the delay/fringe-rate transform will be laid out in the following sections as we provide a brief walk-through of the processing pipeline.
 
 \subsection{Selection of Redshift Bins}
-The redshift bins for which we have computed power spectra --shown in Figure \ref{fig:bins}-- have been selected from the available bandwidth using two criteria: minimizing covariance between redshifts and avoiding missing data.
+The redshift bins for which we have computed power spectra --shown in Figure \ref{fig:bins}-- have been selected from the available bandwidth using two criteria: minimizing covariance between redshifts and avoiding missing spectral data.
 We minimize covariance between adjacent redshift bins by limiting overlap of each bin to the spectrum at the outer half of the bin channel range which, as will be described in section \ref{sec:power_spectrum}, is significantly down-weighted by the use of a Blackman-Harris window.
-Avoiding missing data is important because the power spectrum method, developed in P14, which leverages the redundant baselines to estimate non-sky covariance and project out these contaminated modes is particularly sensitive to missing data -the inverse of the covariance is not defined- so we also select only channels that have no missing data over the entire sidereal period (see the dotted line in Figure \ref{fig:bins}).
+Avoiding missing data is important because the power spectrum method, developed in P14, which leverages the redundant baselines to estimate non-sky covariance and project out these contaminated modes is particularly sensitive to missing data -the inverse of the covariance is not defined- so we also select only redshift ranges with channels that have no missing data over the entire sidereal period (see the dotted line in Figure \ref{fig:bins}).
 With these constraints we arrive at the redshifts 10.3, 8.5,7.9 and 7.45.
@@
 \citet{Dillon:2013p10497} describes a pseudo-inverse method for handling the covariance introduced by missing data.
-Though not implemented here, this method is in the process of being adapted for use in highly redundant analysis \citep{Liu:2014p10462,Liu:2014p10463}.
+Though not implemented here, this method is in the process of being adapted for use in highly redundant analysis \citep{PhysRevD.90.023018,PhysRevD.90.023019}.
 
@@
 \subsection{Initial Averaging}
-  First, the raw data are down-selected to just the 70 30m long baselines described in Section \ref{sec:observations}.
+  First, the raw data are down-selected to just the 70 $\sim$30m long baselines described in Section \ref{sec:observations}.
 %I know this isn't technically true, but I think its important that the downselect is obvious up front.
@@
 This filtering\footnote{See Appendix A of P14} is done in tandem with a radio frequency interference (RFI) flagging step, using the residuals which have had bright sky-like signals removed to flag 4$\sigma$ deviations before feeding the flags back into another iteration of the compression step.
-This mitigates the effects of bright, narrow band, interference  being scattered into higher delay modes (where reionization lives) and results in a time and frequency bin of length 39.6s and width 492.61kHz.
+This mitigates the effects of bright, narrow band, interference  being scattered into higher delay modes (where only the reionization signal is expected to be found) and results in a time and frequency bin of length 40s and width 492.61kHz.
   This process reduces the data volume by a factor of $\sim$20, or roughly an order of magnitude improvement on traditional time and frequency averaging which in this array would be limited to ~100kHz and 10s to avoid averaging away fringes.
@@
 \subsection{Calibration}
-We model the gain as a per-antenna amplitude and a phase slope -physically a single time delay- and single real, low order polynomial passband for all antennae.
-Because the array samples correlations redundantly, the relative calibration between antennae is numerically overdetermined and tractable as a linear algebra problem \citep{Liu:2010p10391}.
+We model the gain as a per-antenna amplitude and a phase slope -physically a single time delay- and single real, low order polynomial passband for all antennas.
+Because the array samples correlations redundantly, the relative calibration between antenna is numerically overdetermined and tractable as a linear algebra problem \citep{Liu:2010p10391}.
 As described in P14, we compute the ratio between redundant baselines, fit for a gain and phase slope and then algebraically solve for a per-antenna solution.
@@
 variation.
-A single solution derived for for the Dec 7 data set is used for the entire observing run.
+A single solution derived for the Dec 7 data set is used for the entire observing run.
 Time and frequency variation of redundant solutions is explored in general in \citet{Zheng:2014p10467}.
 
-Relative calibration is all that is necessary to form a power spectrum.
-To obtain a correctly normalized power spectrum, however, it is necessary to set a flux scale for which we need a flux measurement of a calibration standard.
-For this we form a beam on a known, bright calibration source.
-However, by itself the redundant calibration does not contain enough phase information to phase coherently to a sky location, there remain two free phase parameters which cannot be solved by redundancy alone.
+Relative calibration allows the gains of different antennas to be calibrated against one another.
+However, to obtain a correctly normalized power spectrum, it is necessary to set a flux scale for which we need a flux measurement of a calibration standard.
+For this we form a beamform on a known, bright calibration source.
+By itself, the redundant calibration does not contain enough phase information to phase coherently to a sky location, there remain two free phase parameters which cannot be solved by redundancy alone.
 We find these by fitting a model of Pictor A, Fornax A, and the Crab Nebula during a time when the sky is dominated by these three sources while marginalizing over the unknown\footnote{Unknown in the sense of a joint uncertainty in source flux and primary beam pattern.
@@
 
- \begin{figure}
+ \begin{figure*}
 \centering
@@
 }
-\end{figure} 
+\end{figure*} 
 %The success of the remainder of the processing stages hinges on the spectral stability of the data, which can easily be assessed by comparing redundant baselines.
@@
 \centering
-\includegraphics[width=\textwidth]{figures/psa32_trms_with_and_without_fr_filtering_z.png}
-\caption{\label{fig:noise} Root mean square (rms) noise before and after integrating to 10 minute LST bins (top and bottom sets of curves) indicating the rising significance of foregrounds as we take the last deep integration step.
+\includegraphics[width=\columnwidth]{figures/psa32_trms_with_and_without_fr_filtering_z.png}
+\caption{\label{fig:noise} Root mean square (rms) noise before and after integrating to 13 minute LST bins (top and bottom sets of curves) indicating the rising significance of foregrounds as we take the last deep integration step.
 Noise is estimated by differencing adjacent frequencies (magenta) and between redundant baselines (blue), compared with an estimate of the theoretical noise level with a $T_{sys}$ of 550K (dotted).
@@
 This removes excess correlation due to cross-talk in the analog signal chain.
-The residuals are then flagged once more for RFI before the 92 nights of data are averaged into 36.4 second long local sidereal time (LST) bins, which as PAPER is a drift-scanning instrument, are equivalent to bins in Right Ascension (Declination is fixed at -30\arcdeg).
+The residuals are then flagged once more for RFI before the 92 nights of data are averaged into 40 second long local sidereal time (LST) bins, which as PAPER is a drift-scanning instrument, are equivalent to bins in Right Ascension (Declination is fixed at -30\arcdeg).
 During averaging we found that some LST bins were dominated by a small number of exceedingly bright samples lying well outside the rest of the gaussian distributed data.
@@
 \label{sec:power_spectrum}
-The output of the above steps is a single calibrated and foreground filtered sidereal night of visibility data.
+The output of the above steps is eight sidereal hours of calibrated, foreground filtered visibility data averaged over 92 nights.
 The power spectrum is estimated in the delay spectrum of  a 10MHz bandwidth range centered on the redshift of interest.
@@
 The cross multiplication of the same delay modes between different redundant baselines provides an unbiased estimate of the power spectrum.
-These "sky-like" correlations should be identical between all redundant baselines to within the level of the noise, while all other cross multiplications between delay modes should not be correlated between different baselines.
+These ``sky-like'' correlations should be identical between all redundant baselines to within the level of the noise, while all other cross multiplications between delay modes should not be correlated between different baselines.
 In practice  these non-sky-like modes do occasionally have significant power which leaks into the correlations which sample the power spectrum.
@@
 This is done by dividing the baselines into different groups such that all cross-multiplications are done without introducing noise bias.
-For more see Appendix of C of P14.
+For more see Appendix C of P14.
 
@@
 
-%\begin{figure}
-%\centering
-%\includegraphics[width=0.45\textwidth]{{figures/pspec_log_z_10.29}.png}
-%\includegraphics[width=0.45\textwidth]{{figures/pspec_log_z_8.54}.png}
-%\includegraphics[width=0.45\textwidth]{{figures/pspec_log_z_7.94}.png}
-%\includegraphics[width=0.45\textwidth]{{figures/pspec_log_z_7.55}.png}
-%\caption{\label{fig:pspecs} Left to right from top: power spectra centered on redshifts 10.3,8.5, 7.9, 7.68 and 7.55 (frequencies: 127, 149.5, 159.5 164.5, and 167 MHz.)  All with an effective bandwidth of $\pm$10MHz  covering the redshift span $\Delta z\approx$1421MHz ($B/f^2$) which ranges from 0.8 at $z=$10.3 to 0.45 at $z=$7.4.
-The noise curve (dashes) is calculated using the method described in \cite{Pober:2013p9581}  and indicates the 1$\sigma$ confidence bounds on data points consisting purely of noise; 65\% of uncorrelated noise like data points will lie below the curve.
-However, note the caveats that due to the weighting of the delay transform, adjacent $k$ bins are 50\% correlated.
-It is also clear that the sensitivity calculations under-estimate the sensitivity of the measurements by a small factor.
-See \cite{Pober:2014p10390,Pober:2013p9581} for a discussion of the approximations made in those calculations.
-The black line is a fiducial model at 50\% ionization \citep{Lidz:2008p8251}.
-GMRT points from \cite{Paciga:2013p9943} indicated with 'x's and MWA points (also using 32 antenna) are black diamonds.}
-%
-%\end{figure}
-
-\begin{figure}
+\begin{figure*}
 \centering
@@
 \caption{ \label{fig:pspecs} PAPER power spectra at four redshifts.
-On the top, $P_k$ spectra (a simple units change from the raw delay spectrum) provide a useful diagnostic on foreground rejection, while on the bottom we plot in $\Delta^2$ cosmological units, error bars are 2$\sigma$.
+On the top, $P_k$ spectra (a simple units change from the raw delay spectrum) provide a useful diagnostic on foreground rejection, while on the bottom we plot in $\Delta^2$ cosmological units with 1 $\sigma$ error bars.
 Redshift increases left to right.
@@
 The black line is a fiducial model at 50\% ionization \citep{Lidz:2008p8251}.
-GMRT points from \cite{Paciga:2013p9943} indicated with 'x's and MWA points (also using 32 antenna) are black diamonds.
+GMRT points from \cite{Paciga:2013p9943} indicated with 'x's and MWA points (also using 32 antenna) are black diamonds \cite{Dillon:2014p9788}.
 }
-\end{figure}
+\end{figure*}
 
@@
 \centering
-\includegraphics[width=0.9\textwidth]{{figures/pspec_log_z_7.68}.png}
+\includegraphics[width=0.9\columnwidth]{{figures/pspec_log_z_7.68}.png}
 \caption{\label{fig:P14compare} The redshift 7.68 bin has been reprocessed for comparison with the P14 result.
-See figure \ref{fig:pspecs} for a detailed description of curves.
-Compare with figure 6 of P14.
+See Figure \ref{fig:pspecs} for a detailed description of curves.
+Compare with Figure 6 of P14.
 Note that the LST range processed here is 3 hours less than P14 for a small reduction in sensitivity but large reduction in foreground.
+The data shown here was processed exactly as the data in Figure \ref{fig:pspecs}, the foregrounds have been filtered.
+However the power spectrum shown in P14 added back in the narrow band delay spectrum points (the central five bins inside the horizon which is indicated by dashed lines) for comparison purposes we have added back in these points from that paper.
+They are marked with '+'s.
 }
@@
 %\]
-\[
+\begin{equation}
 \frac{1}{\sqrt{N_{eff}}}  = \left\langle \frac{1}{\sqrt{N(lst)}}  \right\rangle_{\textrm{lst}}
-\]
+\end{equation}
 On average each bin is in this dataset is effectively measured 39.6 times. Given this accounting of the number of samples $N(lst)$, we estimate the r.m.s. temperature of the averaged data set to be
@@
 
-\[
-T_{rms}  =\frac{1}{N_{\textrm{lsts}}}\sum_{\textrm{lst}} \frac{T_{sky}(lst) + T_{rcvr}}{\sqrt{2BtN(lst)}} \approx \frac{(<T_{sky}> + T_{rcvr})}{\sqrt{2BtN_{eff}}}
+\begin{equation}
+T_{rms}  =\frac{1}{N_{\textrm{lsts}}}\sum_{\textrm{lst}} \frac{T_{sky}(lst) + T_{rcvr}}{\sqrt{2BtN(lst)}} \approx \frac{(\langle T_{sky}\rangle + T_{rcvr})}{\sqrt{2BtN_{eff}}}
 \label{eq:Trms}
-\]
+\end{equation}
 
@@
 In P14 the total system temperature ($T_{sky} + T_{rcvr}$) was estimated to be 550K using 20 nights of the 92 night set which covered LSTs 3-9.
-Using the global sky model  \citep{deOliveiraCosta:2008p2242} and a model of the PAPER primary beam we estimate the mean $T_{sky}$ during that 20 night period to be 250K, implying a receiver temperature of 300K.
+Using the global sky model  \citep{deOliveiraCosta:2008p2242} and a model of the PAPER primary beam, we estimate the mean $T_{sky}$ during that 20 night period to be 250K, implying a receiver temperature of 300K.
 The mean temperature over the full 92 night period is only 4K higher; the estimate of 550K for the system temperature is a good estimate for the full observing period.
@@
 The final fringe-rate step is the last coherent average of the processing routine and represents the last large gain in sensitivity; all following averaging steps are done incoherently on the square of the visibilities where sensitivity is gained at a much slower rate.
-For this reason the noise curves (bottom three curves of Figure \ref{fig:noise})  merit closer inspection.
+For this reason the noise curves of the fringe-rate averaged data set (bottom set of curves of Figure \ref{fig:noise})  merit closer inspection.
 At this sensitivity level we see a notable excess of power in the r.m.s.
 difference between baselines.
-Though the cause of this excess is unknown, it is suggestive of the excess covariance between redundant baselines which we remove in Section \ref{sec:power_spectrum} a hypothesis also supported by the high levels of residual foregrounds in the redshift 10.5 bin where the baseline difference r.m.s.
+Though the cause of this excess is unknown, it is suggestive of the excess covariance between redundant baselines which we remove in section \ref{sec:power_spectrum}, a hypothesis also supported by the high levels of residual foregrounds in the redshift 10.5 bin where the baseline difference r.m.s.
 is highest.
@@
 
-In  Figure \ref{fig:pspecs} we show the  power spectra at different redshift bins --showing both the spherically averaged power spectrum $P(k)$ and the volume independent  $\Delta^2(k)\equiv\frac{k^3}{2\pi^2}P(k)$.
-The $P(k)$ spectrum averages the three baseline types but preserves the positive and negative delays.
-In figure \ref{fig:slices} we see different $k$ mode slices as a function of redshift,  all plotted with $2\sigma$ error bars and in Table \ref{tab:data} we list the data plotted in these slices.
+In  Figure \ref{fig:pspecs} we show the  power spectra at different redshift bins --showing both the spherically averaged power spectrum $P(k)$ and the volume weighted  $\Delta^2(k)\equiv\frac{k^3}{2\pi^2}P(k)$.
+The $P(k)$ spectrum averages the three baseline types but preserves the positive and negative delays; it is essentially just the average delay spectrum scaled to temperature and cosmological units.
+Note that due to the choice of window function, the adjacent $k$ bins correlate at the  50\% level.
+In Figure \ref{fig:slices} we see different $k$ mode slices as a function of redshift,  all plotted with $1\sigma$ error bars derived from the bootstrap process and compared with the theoretical noise level.
+Table \ref{tab:data} we list the data plotted in these slices.
 
-The theoretical noise power spectrum (dashed line in figures \ref{fig:pspecs} and \ref{fig:P14compare} and grey region in figure \ref{fig:slices}) is estimated using the method described in \cite{Pober:2013p9581}, assuming a system temperature of 550K and the observing scheme described in Section \ref{sec:observations}.
+The theoretical noise power spectrum (dashed line in Figures \ref{fig:pspecs} and \ref{fig:P14compare} and grey region in Figure \ref{fig:slices}) is estimated using the method described in \cite{Pober:2013p9581}, assuming a system temperature of 550K and the observing scheme described in Section \ref{sec:observations}.
 In all figures the noise levels are plotted for comparison with the power spectrum values rather than the error bars, i.e.
-65\% of uncorrelated points should lie below the line or within the grey region.
-In general the theoretical error bars are larger than the bootstrap errors.
-This is probably most clear in \ref{fig:slices} where, though most points are inside the error region, the bootstrap errors are only consistent with zero 25\% of the time.
+68\% of uncorrelated, noise-dominated points should lie below the line or within the grey region.
+Points where the error bar does not cross zero are inconsistent with noise at $>68$\% these are listed as ``detections'' in Table \ref{tab:data}.
+
+If our measurements were completely noise dominated, $68\%$ of the data points ought to include zero within their error bars.
+In Table \ref{tab:data}, we divide the points into ``Detections" (power spectrum values smaller than $1\sigma$ errors), and ``Upper limits" (measured values larger).
+With only one exception, this scheme yields the same results whether we use the theoretical errors or the bootstrap errors.
+Only a minority, $\sim$20\%, of points are consistent with noise at the 1 $\sigma$ level, and considering all the data points together, the k modes above 0.2 in Table \ref{tab:data} are inconsistent with zero at 6.6 $\sigma$.
+This quantifies the positive bias that is visually apparent in Figures \ref{fig:pspecs} and \ref{fig:slices}.
+
+%In general the theoretical error bars are similar to the bootstrap error bars.
+This is probably most clear in Figure \ref{fig:slices} which is plotted on a linear scale and along the redshift axis where correlations between points should be minimal where, though most points are slightly outside the error region, the bootstrap errors do extend inside the noise dominated region.
 In general it would appear that the sensitivity calculations under-estimate the sensitivity of the measurements by a small factor.
@@
 
-Generally, both figures convey a picture of a power spectrum dominated by noise, with foregrounds encroaching from inside and just outside the horizon to higher $k$ ``reionization'' modes.
+Most of the power spectrum points have an excess power at about the 2$\sigma$ level.
+As these measurements have a sensitivity level which is still two orders of magnitude away from theoretical reionization levels we consider the likelihood that these detections are evidence of reionization to be small.
+Far more likely is that we are limited by a residual foreground or systematic arising from approximations in the data analysis.
+The statistical significance of the residual --2$\sigma$ after averaging over the entire season-- is low enough that it is not possible to break it down along likely axes like sidereal time or time of day to probe its origin.
+Such studies must await higher sensitivity measurements with more antennas.
+However, there are several additional features that seem to suggest an excess originating in foregrounds rather than a HI signal
+
+%  That figure also provides an additional clue in the
+
+%Generally, both figures suggest a power spectrum containing instrumental noise and foregrounds entering at comparable levels, with the latter possibly leaking from within the horizon to modes of higher $k$.
+To see this and other systematic influences that may be present, one can compare the bootstrap error bars, the theoretical error bars, and the power spectrum values.
+After a length of observation similar to what will be necessary to complete a detection level integration with the completed 128 element array, there are clear systematics, though at a low enough level that only a limited analysis of their origins is possible.
+
+%Generally, both figures convey a picture of a power spectrum dominated by noise, with foregrounds encroaching from inside and just outside the horizon to higher $k$ ``reionization'' modes.
 One easy way to see this leakage is by comparing the bootstrap error bar, the theoretical error and the power spectrum value.
 At the level of noise here, we expect all points outside the horizon to be consistent with zero.
-As a simple sorting criteria we divide the points into ``Detections'' and ``Upper Limits'' (Det/ULim) in Table \ref{tab:data} for both theoretical and bootstrap error bar.
-As we can see, the larger theoretical error bars are mostly consistent with zero, while the same is true of a little less then half of the bootstrap error bars.
+As a simple sorting criteria we divide the points into ``Detections'' and ``Upper Limits'' (Det/ULim) in Table \ref{tab:data} for both theoretical and bootstrap error bars, finding agreement for all but one data point.
+A minority $\sim$20\% of points are consistent with noise distributed about zero at 1 $\sigma$.
+Taken together, all the points are inconsistent with zero at 6.6 $\sigma$.
+%As we can see, the larger theoretical error bars are mostly consistent with zero, while the same is true of a little less then half of the bootstrap error bars.
 This is consistent with a picture of systematics or foreground leakage near the thermal noise limit.
 
-Foreground dominance appears to grow with increasing redshift, mirroring the rise in baseline-to-baseline r.m.s.
-we found in Figure \ref{fig:noise}.
-Recall that the wide-band foreground model we have subtracted  --which causes the notch of missing power\footnote{Note that this notch is not present in figure \ref{fig:P14compare} because the foregrounds have been left in for comparison with P14.} near $k_\parallel=0$ in Figure \ref{fig:pspecs}-- is the same for all redshift slices and was built by weighting the entire band by a Blackman-Harris window function.
-Using this window, contributions to the foreground model from 126MHz ($z=$10.3) are down-weighted $\sim$75\% compared to 150Mhz ($z=$8.47).
-If the 126MHz sky or systematics are much different then 150Mhz foreground residuals at that redshift will be higher.
-The $k=0.1$hMpc$^{-1}$ slice (upper left of figure \ref{fig:slices}) is only a factor of $\sim$three away from the light travel horizon of $k=0.035$ and therefore most likely to be contaminated by low $k$ foregrounds.
+The predicted signature of the HI signal during reionization is one of a relatively short burst of power at small $k$ values as the ionized bubbles are briefly at their largest, before the signal from the ever increasing ionized medium drops sharply \citep{Pritchard:2008p8123}.
+In contrast, this observed residual is relatively constant with redshift while rising at both ends of the band, mimicking the RMS noise curve (Figure \ref{fig:noise}).
+This suggests that the accuracy of the foreground filter decreases towards the edges of the band.
+Recall that the wide-band foreground model we have subtracted was built by weighting the entire band by a Blackman-Harris window function which provides a much higher isolation of foregrounds inside the horizon.
+The tradeoff that is made is that data in the outer parts of the spectrum are heavily down weighted.
+For example data at 126MHz are down-weighted $\sim$75\% compared to 150Mhz.
+This amounts to making an assumption about the spectral smoothness of the data, namely that 126MHz can be mostly modeled by extrapolation from 150MHz.
+The precision of this subtraction can be judged by examining the foreground residuals inside the horizon in the $P(k)$ power spectra plotted in the upper part of Figure \ref{fig:pspecs}\footnote{Note that this notch is not present in Figure \ref{fig:P14compare} because the foregrounds have been left in for comparison with P14.}.
+The relative depth of the redshift 10.5 P(k) foreground residual is substantially shallower then at 7.55 and is generally correlated with the amount of residual found at higher $k$s.
+
+The residual level is also illustrated in the $k$ slices of Figure \ref{fig:slices}.  
+ The shortest $k$ mode (upper left of Figure \ref{fig:slices}) is the nearest to the light travel horizon and therefore most likely to be contaminated by foreground leakage.
 This $k$ mode is well above the noise and has a redshift dependence much the same shape as the $T_{rms}$ curve in Figure \ref{fig:noise}, with a minimum near redshift 8.5 and a dramatic rise above redshift 10.
+Longer $k$ modes are all generally positively biased, though none higher than 3 $\sigma$.
 
-\begin{figure}
+\begin{figure*}
 \centering
@@
 The k=0.1 hMpc$^{-1}$ bin (top-left), which samples the delay spectrum at only 2x the maximum horizon delay  is foreground dominated with  a redshift dependence similar to the  $T_{rms}$ residual in Figure \ref{fig:noise}.
-The rest have amplitudes below the predicted noise level, but are not themselves perfectly consistent with noise, particularly at high redshifts.
-Though the points are all within the region predicted to be dominated by noise (grey) the variation between the redshifts and size of error bars suggests that a small residual signal is probably still present.
-The theoretical noise estimate is likely to be off by a small amount (see figure \ref{fig:pspecs} and Section \ref{sec:pspecs}).
+Most have amplitudes at or slightly above the noise level (grey region), particularly at high redshifts suggesting that most are systematic limited.
+Calculating the theoretical noise requires making several approximations (see Figure \ref{fig:pspecs} and Section \ref{sec:pspecs}).
 \label{fig:slices}}
-\end{figure}
 
-\begin{deluxetable}{llrrrrr}
+\end{figure*}
+
+\begin{deluxetable*}{llrrrrr}
 \tablecolumns{7}
-\tablecaption{Power spectrum values}
+\tablecaption{Power spectrum values.
+Most are inconsistent with noise at $\sim$2$\sigma$, indicating that this full season of integration sees systematics at about twice the level of the noise.
+In sum the points above $k=0.2$ are inconsistent with zero at 6.6 $\sigma$.
+On average the theoretical and bootstrap estimates of the noise agree to within 15\%.}
 \tablehead{
@@
 \colhead{bootstrap error } &\colhead{\parbox{6em}{noise model error} }& \colhead{\parbox{5em}{bootstrap significance}} & \colhead{\parbox{6em}{noise model significance}}\\
-
 \colhead{[hMpc$^{-1}$]} & \colhead{}  & \colhead{[mK$^2$]} &
@@
 \startdata
-0.1 & 10.29 & 43575.5512 & $\pm$ 5729.3347 & $\pm$ 8029.1131 & Det & Det \tabularnewline
-0.1 & 8.54 & 5701.6267 & $\pm$ 1168.7025 & $\pm$ 3671.887 & Det & Det \tabularnewline
-0.1 & 7.94 & 7695.373 & $\pm$ 2044.5824 & $\pm$ 3133.8433 & Det & Det \tabularnewline
-0.1 & 7.55 & 9461.8178 & $\pm$ 2761.5645 & $\pm$ 2393.1972 & Det & Det \tabularnewline
-0.2 & 10.29 & 11410.9746 & $\pm$ 5039.8271 & $\pm$ 27281.158 & Det & ULim \tabularnewline
-0.2 & 8.54 & 5855.8909 & $\pm$ 2195.8363 & $\pm$ 12275.7791 & Det & ULim \tabularnewline
-0.2 & 7.94 & 5143.6577 & $\pm$ 1799.7316 & $\pm$ 10221.3583 & Det & ULim \tabularnewline
-0.2 & 7.55 & 856.316 & $\pm$ 1490.2202 & $\pm$ 7636.2256 & ULim & ULim \tabularnewline
-0.3 & 10.29 & 34756.4662 & $\pm$ 16384.7719 & $\pm$ 64977.9446 & Det & ULim \tabularnewline
-0.3 & 8.54 & 12923.2987 & $\pm$ 5373.4928 & $\pm$ 29058.6627 & Det & ULim \tabularnewline
-0.3 & 7.94 & 5204.3757 & $\pm$ 4834.4927 & $\pm$ 24046.4004 & ULim & ULim \tabularnewline
-0.3 & 7.55 & 7628.4564 & $\pm$ 4810.3274 & $\pm$ 17762.1184 & Det & ULim \tabularnewline
-0.4 & 10.29 & 42909.9827 & $\pm$ 30695.0274 & $\pm$ 128355.9988 & ULim & ULim \tabularnewline
-0.4 & 8.54 & 24115.248 & $\pm$ 16585.6317 & $\pm$ 56571.7524 & Det & ULim \tabularnewline
-0.4 & 7.94 & 14730.7401 & $\pm$ 9930.2065 & $\pm$ 46725.4956 & Det & ULim \tabularnewline
-0.4 & 7.55 & -8601.8091 & $\pm$ 14191.4889 & $\pm$ 34339.7156 & ULim & ULim \tabularnewline
+0.1 & 10.29 & 43575.6 & $\pm$ 5729.3 & $\pm$ 676.4 & Det & Det \tabularnewline
+0.1 & 8.54 & 5701.6 & $\pm$ 1168.7 & $\pm$ 279.7 & Det & Det \tabularnewline
+0.1 & 7.94 & 7695.4 & $\pm$ 2044.6 & $\pm$ 229.3 & Det & Det \tabularnewline
+0.1 & 7.55 & 9461.8 & $\pm$ 2761.6 & $\pm$ 153.3 & Det & Det \tabularnewline
+0.2 & 10.29 & 11411.0 & $\pm$ 5039.8 & $\pm$ 5326.1 & Det & Det \tabularnewline
+0.2 & 8.54 & 5855.9 & $\pm$ 2195.8 & $\pm$ 2184.4 & Det & Det \tabularnewline
+0.2 & 7.94 & 5143.7 & $\pm$ 1799.7 & $\pm$ 1785.6 & Det & Det \tabularnewline
+0.2 & 7.55 & 856.3 & $\pm$ 1490.2 & $\pm$ 1190.0 & ULim & ULim \tabularnewline
+0.3 & 10.29 & 34756.5 & $\pm$ 16384.8 & $\pm$ 17922.7 & Det & Det \tabularnewline
+0.3 & 8.54 & 12923.3 & $\pm$ 5373.5 & $\pm$ 7339.6 & Det & Det \tabularnewline
+0.3 & 7.94 & 5204.4 & $\pm$ 4834.5 & $\pm$ 5996.1 & Det & ULim \tabularnewline
+0.3 & 7.55 & 7628.5 & $\pm$ 4810.3 & $\pm$ 3993.6 & Det & Det \tabularnewline
+0.4 & 10.29 & 42910.0 & $\pm$ 30695.0 & $\pm$ 42439.6 & Det & Det \tabularnewline
+0.4 & 8.54 & 24115.2 & $\pm$ 16585.6 & $\pm$ 17370.2 & Det & Det \tabularnewline
+0.4 & 7.94 & 14730.7 & $\pm$ 9930.2 & $\pm$ 14187.9 & Det & Det \tabularnewline
+0.4 & 7.55 & -8601.8 & $\pm$ 14191.5 & $\pm$ 9447.4 & ULim & ULim \tabularnewline
 \enddata
-\tablenotetext{a}{Det indicates a measurement and error-bar inconsistent with zero, ULim indicates consistency with zero at 2 $\sigma$.}
+\tablenotetext{a}{All error bars are 1 $\sigma$.}
+\tablenotetext{b}{Det indicates a measurement and error-bar inconsistent with zero, ULim indicates consistency with zero at 1 $\sigma$.}
 \label{tab:data}
-\end{deluxetable}
+\end{deluxetable*}
 
@@
 The present analysis extends the work of  \cite{Parsons:2014p10499}, which used a single-baseline delay spectrum analysis to place an upper limit on the 21 cm power at z = 7.68 (163 MHz).
-In this work, we take advantage of the wide bandwidth of the PAPER instrument to place limits on the power spectrum at a range of redshifts $10.7>z>7.2$, demonstrating that the algorithms used in P14 can exploit the full potential of the 21 cm line.
+In this work, we take advantage of the wide bandwidth of the PAPER instrument to place limits on the power spectrum at a range of redshifts $10.7>z>7.2$.
 
-Most of the power spectrum data points demonstrate a removal of foreground signals to approximately the thermal noise limit, leaving systematics on the scale of the noise.
-Noise limited bins --those which are statistically consistent with zero and marked as ULim in Table \ref{tab:data}-- occur at several $k$ modes at redshifts 7.55, 7.94 and 10.29.
-The deepest point --at redshift 7.55 and $k=0.2$-- is a 2$\sigma$ upper limit of (48mK)$^2$.
+Most of the power spectrum data points demonstrate a removal of foreground signals to approximately twice the thermal noise limit.
+The deepest point --at redshift 7.55 and $k=0.2$-- is a 1$\sigma$ upper limit of (48mK)$^2$.
 For comparison, the \citet{Lidz:2008p8251} fiducial model of the reionization power spectrum estimates an amplitude of (4.4mK)$^2$, for models with 50\% reionization at $z=$7.55.
+Though most points are apparent detections of residual foregrounds they are still at a much deeper limit then previously found at these redshifts.
 
 The fact that the best upper limit still comes from the z = 7.6 band does illustrate that this redshift corresponds to a somewhat special frequency for the PAPER instrument, where the combined contribution of system noise, RFI signals, and residual foregrounds (if any remain) are at a minimum.
+The success of the filtering style foreground removal heavily leverages the surrounding spectral coverage, and is somewhat less effective at removing foregrounds across the entire band.
 
@@
 
-Future work will also include observations from subsequent seasons with more antenna.
+Future work will also include observations from subsequent seasons with more antennas.
 The observations reported here demonstrate a nearly noise limited integration over a full season but used only a quarter of the final design antenna count.
-These  32 antenna observations from 2011 were followed by a full season of 64 antennae in 2012 and 128 in 2013.
+These  32 antenna observations from 2011 were followed by a full season of 64 antennas in 2012 and 128 in 2013.
 A second season of observing with the 128 antenna array is now under way.
@@
 It confirms the effectiveness of the wideband CLEAN in removing foregrounds over a wide range of the band, while illustrating weaknesses in this approach in pushing towards the edges of PAPER's frequency range.
-Furthermore, after the application of the fringe rate-based time averaging, we see hints that baseline-to-baseline variations between redundantly spaced baselines is becoming the dominant systematic in the analysis.
+Furthermore, after the application of the fringe rate-based time averaging, we see hints that baseline-to-baseline variations between redundantly spaced baselines are becoming the dominant systematic in the analysis.
 Techniques like the covariance removal introduced in P14 and also used in this work can differentiate some of this variation from the sky signal; however some combination of improved calibration techniques \citep{Zheng:2014p10467}, model building \citep{Sullivan:2012p9457}, and a better understanding of error covariance  \citep{Liu:2011p8763} may also prove valuable in reducing this systematic.
 
-%(iii) third paragraph would then be on lowest freq bin and methods for improvement in analysis
+\section{Acknowledgements}
+The PAPER project is supported through the NSF-AST program (award \#1129258).
+Computing resources were provided by a grant from Mt.
+Cuba Astronomical Foundation.
+J.C.P.
+is supported by an NSF Astronomy and Astrophysics Fellowship under award AST \#1302774.
+We thank Paul Sutter and Jonathan Pritchard for helpful discussions.
 
-%With only 32 antenna, this data lacks the sensitivity to exclude all but the most extreme models, however, it does demonstrate the ability of the wideband filtration method to affectively remove foregrounds to the precision needed to integrate a full season to the thermal noise limit.
-As future observations with PAPER  will add sensitivity primarily by increasing to 128 antennae this projects well for our ability to reach design sensitivity.
-The residual foregrounds analyzed here are all at or near the level of noise and expose the next level of removal required to access a wider redshift range.
-
-%\begin{align}
-%\label{eq:BrightnessFluctuations}
-% \Delta T &\approx \frac{T_S - T_{CMB}}{1 +z} \tau_{21} \nonumber\\
-% &\approx T_0(z) \left( \frac{T_S - T_{CMB}}{T_S} \right) x_{HI} (1 + \delta) 
-%\end{align}
-%where $T_S$ is the 21cm line spin temperature, $x_{HI}$ is the local ionization fraction and $\delta$ is the mass over-density.
-All, save $T_0$, are position dependent.
-$T_0$ encodes the global temperature evolution due to cosmological expansion
-%\begin{align}
-% T_0(z) & = 23~{\rm mK}~\left( \frac{\Omega_b h^2}{0.02} \right) \left[ \left(
-% \frac{0.15}{\Omega_m h^2} \right) \left( \frac{1+z}{10} \right) \right]^{1/2}\nonumber\\
-% & = 25~{\rm mK}~\left( \frac{1+z}{10} \right)^{1/2}
-%\label{eq:Prefactor}
-%\end{align}
-%\citep[see, e.g., ][]{zaldarriaga_et_al2004,furlanetto_et_al2006}, where we have used the
-%Planck 2013 parameters \citep{planck_et_al2013}.
-
-\bibliography{library}
+\bibliography{library,aliu}
 

### 2014-07-30 generate_AAS.py
adding script to create journal submission

@@ -1,0 +1,69 @@
+#parse through the tex
+#find figures
+#replace filename with .eps
+#create/cleanup target directory
+#convert list of files
+#copy other necessary files
+import re,os
+
+#target = 'arxiv_submission'
+project=os.path.basename(os.getcwd())
+target = 'apj_submission_1'
+
+if target.startswith('arxiv'):
+    tex = open('ms-arxiv.tex').read()
+    plots = re.findall('\\includegraphics.*\{([^}]+)\}',tex)
+else:
+    tex = open('ms.tex').read()
+#    plots = re.findall('\{figures\/(\S*)\}',tex)
+    plots = re.findall('\\includegraphics.*\]{(.*)\}',tex)
+
+#clean out the '}'s from the figure paths
+l = ['a','b','c','d']
+eps_plots = [None]*len(plots)
+print len(eps_plots)
+plot_filenames = [None]*len(plots)
+for i in xrange(len(plots)):
+    print plots[i]
+    plot_filenames[i] = plots[i].replace('}','').replace('{','')
+    if i in [0,1,2,3]:
+        eps_plots[i] = 'f{n}.eps'.format(n=i+1)
+    if i>3:
+        eps_plots[i] = 'f5{letter}.eps'.format(letter=l[i-4])
+
+
+try:
+    os.mkdir(target)
+except(OSError):
+    os.system('rm -rf %s'%target)
+    os.mkdir(target)
+#eps_plots = [p[:-4]+'.eps' for p in plots]
+#Copy the plots to the new directory and point at the new figures in the tex
+for i,plot in enumerate(plots):
+    print plot,plot_filenames[i],'->',eps_plots[i]
+    if target.startswith('arxiv'):
+        print "cp ",plot_filenames[i],target
+        os.system('cp %s %s/'%(plot_filenames[i],target))
+        tex = tex.replace(plot,plot)
+        print plot
+    else:
+        print "cp ",plot_filenames[i],target
+        os.system('convert %s %s/%s'%(plot_filenames[i],target,eps_plots[i]))
+        print 'convert %s %s/%s'%(plot_filenames[i],target,eps_plots[i])
+        print 'inserting ',eps_plots[i].replace('.eps','')
+        tex = tex.replace(plot,eps_plots[i].replace('.eps',''))
+
+#replace bibliography 
+#\bibliography{library}
+bib = open('ms.bbl').read()
[…]

### 2014-08-04 ms.tex
compiles for aastex
revised Trms section. now with more clarity

@@
-%\documentclass[fleqn,10pt,iop]{emulateapj}
-\documentclass{aastex}
+\documentclass[fleqn,10pt,iop]{emulateapj}
+%\documentclass[preprint2]{aastex}
 
@@
 
+\tabletypesize{\scriptsize}
+
 	% End definitions
@@
 Daniel C. Jacobs\altaffilmark{1},
+Jonathan C. Pober\altaffilmark{12},
 Aaron R. Parsons\altaffilmark{2,8},
@@
 David F. Moore\altaffilmark{3},
-Jonathan C. Pober\altaffilmark{12},
 Irina I. Stefan\altaffilmark{10},
@@
 \begin{abstract}
-We present new observations from the Donald C.
-Backer Precision Array for Probing the Epoch of Reionization (PAPER) telescope which place new deeper limits on the HI power spectrum over the redshift range of $7.5<z<10.5$, extending previously published single redshift results to cover the full range accessible to the instrument.
 The epoch of reionization power spectrum is expected to evolve strongly with redshift, and it is this variation with cosmic history that will allow us to begin to place constraints on the physics of reionization.
-The primary obstacle to this goal are bright foregrounds and the attendant systematics associated with high dynamic range measurements.
-Here we present an integration across a full season of data and find systematics just beginning to dominate, though well below any levels previously reported.
-This  500 hour integration is the longest such yet recorded and tests foreground removal to a theoretical dynamic range of 10,000 to 1.
-We use filtering techniques that take advantage of the large instrumental bandwidth to make a high dynamic range subtraction of foreground power by averaging together many low sensitivity measurements.
-.
- Power spectra at different points across the redshift range reveal the variable efficacy of the foreground subtraction.
+The primary obstacle to the measurement of the EoR power spectrum is bright foreground emission.
+We present an analysis of observations from the Donald C.
+Backer Precision Array for Probing the Epoch of Reionization (PAPER) telescope which place new limits on the HI power spectrum over the redshift range of $7.5<z<10.5$, extending previously published single redshift results to cover the full range accessible to the instrument.
+To suppress foregrounds, we use filtering techniques that take advantage of the large instrumental bandwidth to isolate and suppress foreground leakage into the interesting regions of $k$-space.
+Our 500 hour integration is the longest such yet recorded and demonstrates this method to a dynamic range of $10^4$.
+Power spectra at different points across the redshift range reveal the variable efficacy of the foreground isolation.
 Noise limited measurements of $\Delta^2$ at $k=$0.2hMpc$^{-1}$ and z$=7.55$ reach as low as (48mK)$^2$ ($1\sigma$).
-Most spectra demonstrate a robust foreground removal of foreground signals to the thermal noise limit, leaving systematics on the scale of the noise.
-To better understand the source of these residual systematics we examine different estimates of the noise in the data and compare with theoretical predictions.
-We find that in the case of the highest redshift power spectrum, foregrounds begin to dominate over the noise, reflecting the increased difficulty in filtering foregrounds near the edge of the instrumental band.
+We demonstrate that the size of the error bars in our power spectrum measurement as generated by a bootstrap method is consistent with the fluctuations due to thermal noise.
+Relative to this thermal noise, most spectra exhibit and excess of power at a few sigma.
+The likely sources of this excess include residual foreground leakage, particularly at the highest redshift, and unflagged RFI.
+We conclude by discussing data reduction improvements that promise to remove much of this excess.
 \end{abstract}
@@
 At this noise level the frequency and baseline differences are roughly similar, both demonstrate the same small bumps of increased noise due to interference flagging and are consistent with noise over much of the band.
-The bottom three lines show the noise level after integrating up to the maximum fringe rate of 776s.
-The step between the top and bottom sets is the last coherent integration with noise largely decreasing by the expected factor $\sqrt{776/40}$ except in the difference between baselines (blue) which demonstrates a clear excess at all frequencies, particularly above redshift 10.
+The bottom three lines show the noise level after integrating up to the maximum fringe rate of 789s.
+The step between the top and bottom sets is the last coherent integration with noise largely decreasing by the expected factor $\sqrt{789/40}$ except in the difference between baselines (blue) which demonstrates a clear excess at all frequencies, particularly above redshift 10.
 With this last large jump in sensitivity we are now seeing the slight dominance of baseline covariance over the spectral derivative rms.
@@
 \label{sec:noise}
-The root-mean-square of the filtered visibilities measures how well we've removed foregrounds.
+The root-mean-square of differenced visibilities measures how well we've removed foregrounds.
 In Figure \ref{fig:noise} we examine $T_{rms}$ as calculated by differencing between adjacent channels and between redundant baselines.
-We see that in the last stage of coherent integration, the step between 40 second to 789 second integrations, the noise level decreases by the expected factor of $\sqrt{789/40}$.
-At both stages the noise level found by both methods is roughly consistent with the theoretical level, only deviating significantly at the edges where the effective signal to noise due to the decreasing passband gain, Blackman Harris window, and uneven sampling (see Fig \ref{fig:bins}) drops precipitously.
+The square of this difference is then averaged over the full LST range probed by the power spectra (1h-9h).
+Note that this operation combines all LST modes as though they were repeated measurements of the same sky when in practice the power spectrum values are averaged in quadrature over different LST bins.
+This provides a lower bound on the sensitivity possible in the data set and a sensitive probe on foreground residuals or other sources of excess power.
+Using all the data in this way, we expect to illuminate rare outliers that effect the bottom line.
 
-The theoretical visibility r.m.s. noise level is a straightforward calculation, but given the method employed here, should be given due description.
-The noise level in a long integration (Eq.
-\ref{eq:Trms}) is caused by the temperature of the sky (mostly due to galactic emission) which varies with LST and by the receiver noise, both on the scale of a few 100K.
-The noise of any long integration is also dependent on the number of observations in each LST.
-The number of points in each bin is mostly determined by the observing schedule, which for PAPER is sunrise to sunset, but also by any flagging of data or observing outage (only 92 of the 103 days in the observing period have recorded data).
-Unfortunately, technical problems limited the retention of the exact cumulative count of points averaged into each LST bin during the averaging process.
-Here we have chosen to use the first order estimate based on the observing schedule to estimate the number of observations in each LST.
-Some bins are observed only a few times, while a slim range in the LST range 6 - 9 hours is observed on every night.
-Over a number of LST bins $N_\textrm{lst}$ the average effective time per bin $N_{eff}$ is given by
+\subsubsection{The Theoretical RMS Noise Level}
 
-%\[
-%\frac{1}{\sqrt{N_{eff}}} = \frac{1}{N_{\textrm{lsts}}} \sum_\textrm{lst=1}^\textrm{lst=9}{\frac{1}{\sqrt{N(lst)}}}
-%\]
-\begin{equation}
-\frac{1}{\sqrt{N_{eff}}}  = \left\langle \frac{1}{\sqrt{N(lst)}}  \right\rangle_{\textrm{lst}}
-\end{equation}
-On average each bin is in this dataset is effectively measured 39.6 times. Given this accounting of the number of samples $N(lst)$, we estimate the r.m.s. temperature of the averaged data set to be
-
-%\[
-%T_{rms}  =\frac{1}{N_{\textrm{lsts}}}\sum_{\textrm{lst}} \frac{T_{sky}(lst) + T_{rcvr}}{\sqrt{2BtN(lst)}} \approx \frac{1}{\sqrt{2BtN_{eff}}} \left(\frac{1}{N_{\textrm{lsts}}}\sum_{\textrm{lst}}T_{sky}(lst) + T_{rcvr}\right)
-%\label{eq:Trms}
-%\]
+The expected noise level over this time range depends on the number of samples in each lst bin and the system temperature in that bin.
+System noise comes in part from the bright radio sky and part from amplifier noise in the receiver.
+As well as the channel width $B$ and the integration time $t$
 
 \begin{equation}
-T_{rms}  =\frac{1}{N_{\textrm{lsts}}}\sum_{\textrm{lst}} \frac{T_{sky}(lst) + T_{rcvr}}{\sqrt{2BtN(lst)}} \approx \frac{(\langle T_{sky}\rangle + T_{rcvr})}{\sqrt{2BtN_{eff}}}
+T_{rms}  =\frac{1}{N_{\textrm{lsts}}}\sum_{\textrm{lst}} \frac{T_{sky}(lst) + T_{rcvr}}{\sqrt{2BtN(lst)}} 
 \label{eq:Trms}
@@
 
-\noindent{where B and t are the 496kHz channel width and lst bin length, respectively and the approximation indicates that we have assumed average values for the number of samples and the time dependent sky component of the system temperature.
-In P14 the total system temperature ($T_{sky} + T_{rcvr}$) was estimated to be 550K using 20 nights of the 92 night set which covered LSTs 3-9.
-Using the global sky model  \citep{deOliveiraCosta:2008p2242} and a model of the PAPER primary beam, we estimate the mean $T_{sky}$ during that 20 night period to be 250K, implying a receiver temperature of 300K.
-The mean temperature over the full 92 night period is only 4K higher; the estimate of 550K for the system temperature is a good estimate for the full observing period.
-$T_{rms}$ using 550K and an effective sample count of 39.6  is plotted as a dotted line in Figure \ref{fig:noise} for before and after fringe rate filtering (t=40s and 789s)}
+Where $T_{sky}(lst)$ is modeled by the convolution of the global sky model \citep{deOliveiraCosta:2008p2242} with the PAPER beam \citep{Pober:2012p8800} which for the LST range here averages to 250 $(\nu/160)^(-2.6)$~K.
 
+One of the strongest influences on the sensitivity of wide-field, low frequency instruments like PAPER or the MWA is the number of samples local sidereal time bin available in an observing year.
+After removing data containing Sun and the Galactic center, both of which are sources of strong foreground emission, the amount of time required to record a deep $\sim$1000 hour observation can stretch into multiple years.
+Some LST bins in this range will never be sampled more than a few times per year.
+This observation, which ran from December to March, measured samples over half the night sky.
+However only the range 6h42 to 9h was measured every night to reach ``full'' sensitivity of the three-month-long observing season.
+Integrating equation \ref{eq:Trms} over this sampling function we find a $T_{rms}$ noise level equivalent to 40 nights of integration.
+
+Using the variance of a 20 day subset, P14 found the system temperature for this LST range to be 550~K at 160~MHz, while the sky model suggests a $T_{sky}$ of 250~K.
+This level of sky temperature is confirmed by the fit of the theoretical noise level in Figure \ref{fig:noise}.
+The sky portion of the noise has a significant slope.
+Ascribing more then 250~K of the total noise to the sky results in a significantly steeper theoretical $T_{rms}$ in Figure \ref{fig:noise}.
+Taken together this evidence points to a receiver noise temperature of 300~K.
+
+In summary we find that the noise level of this data set is well modeled by a 550~K system temperature which is nearly equal parts sky and receiver and when integrated over our lst sampling function is equivalent to 40 days of data.
+
+\subsubsection{Analysis of the RMS Noise Level}
+
+Here we have estimated the residual noise after foreground filtering and LST averaging, both before and after the last ``coherent'' integration step, when fringe rate filtering increases the effective integration time from 40 to 789 seconds.
 The final fringe-rate step is the last coherent average of the processing routine and represents the last large gain in sensitivity; all following averaging steps are done incoherently on the square of the visibilities where sensitivity is gained at a much slower rate.
-For this reason the noise curves of the fringe-rate averaged data set (bottom set of curves of Figure \ref{fig:noise})  merit closer inspection.
-At this sensitivity level we see a notable excess of power in the r.m.s.
-difference between baselines.
-Though the cause of this excess is unknown, it is suggestive of the excess covariance between redundant baselines which we remove in section \ref{sec:power_spectrum}, a hypothesis also supported by the high levels of residual foregrounds in the redshift 10.5 bin where the baseline difference r.m.s.
+The noise level of the 40s average (top curves of Figure \ref{fig:noise}) is consistent between the two methods of noise estimation and with the predicted level.
+
+After the last stage of coherent integration (from the top set of curves to the bottom set), where the integration increases from 40 to 789 second integrations.
+At this deepest level of integration the noise levels rise slightly above the expected decrease of $\sqrt{789/40}$.
+In both cases the noise curves deviate significantly from the theory at the edges of the band.
+This is mainly due to our neglect of variable sensitivity across the band (e.g.
+see Figure \ref{fig:bins}) in calculating the theoretical curve; note that the band edges still integrate down between the two steps.
+However it does not integrate down perfectly, residual levels are apparent across the entire band with larger excesses at the band edges.
+The largest excess is noted in the rms difference between redundant baselines.
+Lack of agreement between supposedly redundant baselines is suggestive of a breakdown in our assumption of spectrally and temporally smooth calibration.
+It is also suggestive of the excess covariance between redundant baselines which we remove in section \ref{sec:power_spectrum}, a hypothesis also supported by the high levels of residual foregrounds in the redshift 10.5 bin where the baseline difference r.m.s.
 is highest.
@@
 With only one exception, this scheme yields the same results whether we use the theoretical errors or the bootstrap errors.
-Only a minority, $\sim$20\%, of points are consistent with noise at the 1 $\sigma$ level, and considering all the data points together, the k modes above 0.2 in Table \ref{tab:data} are inconsistent with zero at 6.6 $\sigma$.
-This quantifies the positive bias that is visually apparent in Figures \ref{fig:pspecs} and \ref{fig:slices}.
+Under this scheme, we would expect a noise dominated signal to yield around eight ``upper limits'' with an equal number of positive and negative data points.
+Here we find only two to three upper limits while rest we classify as detections.
+In general, most detections are within 1 - 2 $\sigma$ but are all positively biased,  with the significance rising steeply at high redshifts and low $k$s.%, the quadratic sum of all errors above $k=0.1$ is 6.6 $\sigma$ indicating that the excess is largely coherent across the data set.
 
-%In general the theoretical error bars are similar to the bootstrap error bars.
-This is probably most clear in Figure \ref{fig:slices} which is plotted on a linear scale and along the redshift axis where correlations between points should be minimal where, though most points are slightly outside the error region, the bootstrap errors do extend inside the noise dominated region.
-In general it would appear that the sensitivity calculations under-estimate the sensitivity of the measurements by a small factor.
-See \cite{Pober:2014p10390,Pober:2013p9581} for a discussion of the approximations made in those calculations.
+\subsection{The Excess Signal}
 
-Most of the power spectrum points have an excess power at about the 2$\sigma$ level.
-As these measurements have a sensitivity level which is still two orders of magnitude away from theoretical reionization levels we consider the likelihood that these detections are evidence of reionization to be small.
+As these measurements have a maximum possible sensitivity level which is still two orders of magnitude away from theoretical reionization levels we consider the likelihood that these detections are evidence of reionization to be small.
 Far more likely is that we are limited by a residual foreground or systematic arising from approximations in the data analysis.
@@
 Such studies must await higher sensitivity measurements with more antennas.
-However, there are several additional features that seem to suggest an excess originating in foregrounds rather than a HI signal
+However, there are several additional features that seem to suggest an excess not caused by high redshift HI.
 
-%  That figure also provides an additional clue in the
-
-%Generally, both figures suggest a power spectrum containing instrumental noise and foregrounds entering at comparable levels, with the latter possibly leaking from within the horizon to modes of higher $k$.
-To see this and other systematic influences that may be present, one can compare the bootstrap error bars, the theoretical error bars, and the power spectrum values.
-After a length of observation similar to what will be necessary to complete a detection level integration with the completed 128 element array, there are clear systematics, though at a low enough level that only a limited analysis of their origins is possible.
-
-%Generally, both figures convey a picture of a power spectrum dominated by noise, with foregrounds encroaching from inside and just outside the horizon to higher $k$ ``reionization'' modes.
-One easy way to see this leakage is by comparing the bootstrap error bar, the theoretical error and the power spectrum value.
-At the level of noise here, we expect all points outside the horizon to be consistent with zero.
-As a simple sorting criteria we divide the points into ``Detections'' and ``Upper Limits'' (Det/ULim) in Table \ref{tab:data} for both theoretical and bootstrap error bars, finding agreement for all but one data point.
-A minority $\sim$20\% of points are consistent with noise distributed about zero at 1 $\sigma$.
-Taken together, all the points are inconsistent with zero at 6.6 $\sigma$.
-%As we can see, the larger theoretical error bars are mostly consistent with zero, while the same is true of a little less then half of the bootstrap error bars.
-This is consistent with a picture of systematics or foreground leakage near the thermal noise limit.
-
+The spectral signature of the measured excess is far more similar to a foreground residual then to theoretical HI models.
 The predicted signature of the HI signal during reionization is one of a relatively short burst of power at small $k$ values as the ionized bubbles are briefly at their largest, before the signal from the ever increasing ionized medium drops sharply \citep{Pritchard:2008p8123}.
-In contrast, this observed residual is relatively constant with redshift while rising at both ends of the band, mimicking the RMS noise curve (Figure \ref{fig:noise}).
+In contrast, this observed residual is relatively constant with redshift while rising at both ends of the band, mimicking the RMS noise curve in Figure \ref{fig:noise}.
 This suggests that the accuracy of the foreground filter decreases towards the edges of the band.
@@
 This amounts to making an assumption about the spectral smoothness of the data, namely that 126MHz can be mostly modeled by extrapolation from 150MHz.
-The precision of this subtraction can be judged by examining the foreground residuals inside the horizon in the $P(k)$ power spectra plotted in the upper part of Figure \ref{fig:pspecs}\footnote{Note that this notch is not present in Figure \ref{fig:P14compare} because the foregrounds have been left in for comparison with P14.}.
+The precision of this subtraction can be judged by examining residuals inside the horizon in the $P(k)$ power spectra plotted in the upper part of Figure \ref{fig:pspecs}\footnote{Note that this notch is not present in Figure \ref{fig:P14compare} because the foregrounds have been left in for comparison with P14.}.
 The relative depth of the redshift 10.5 P(k) foreground residual is substantially shallower then at 7.55 and is generally correlated with the amount of residual found at higher $k$s.
@@
 This $k$ mode is well above the noise and has a redshift dependence much the same shape as the $T_{rms}$ curve in Figure \ref{fig:noise}, with a minimum near redshift 8.5 and a dramatic rise above redshift 10.
-Longer $k$ modes are all generally positively biased, though none higher than 3 $\sigma$.
+These points are also the only ones where error bars significantly deviate from their predicted thermal values.
+As we see in Table \ref{tab:data}, going by predicted noise alone points closest to the horizon have error bars in excess of 20 $\sigma$ while the bootstrap error bars register a much larger uncertainty.
+This hints at a more complicated origin for the excess.
+If these points were purely foregrounds, the error bars would be noise dominated, the fact that the excess is not much larger then its variation suggests a source of error not accounted in our analysis.
+Likely culprits include incomplete flagging of interference, instrumental glitches, or time of frequency dependent calibration variation which introduces differences between redundant baselines.
+All of which have are part of ongoing analysis efforts which are briefly discussed in the conclusions.
+All of these effects have the effect of adding time and frequency dependence to the bright foregrounds limiting the effectiveness of the filter.
+Thus, just like foregrounds, the strong excess drops rapidly towards higher $k$ modes.
+Above $k$ of 0.1 hMpc$^{-1}$  the power spectra are generally positively biased, though most close to 2 $\sigma$ or less using either theoretical and bootstrap error bars.
 
@@
 
-\begin{deluxetable*}{llrrrrr}
-\tablecolumns{7}
-\tablecaption{Power spectrum values.
-Most are inconsistent with noise at $\sim$2$\sigma$, indicating that this full season of integration sees systematics at about twice the level of the noise.
-In sum the points above $k=0.2$ are inconsistent with zero at 6.6 $\sigma$.
-On average the theoretical and bootstrap estimates of the noise agree to within 15\%.}
+\begin{deluxetable}{llccccccc}
+\tablecolumns{9}
+\tablecaption{PAPER Power spectrum values using 32 antenna and a full season of integration.}
 \tablehead{
-\colhead{k} & \colhead{redshift}  & \colhead{$\Delta^2$} &
-\colhead{bootstrap error } &\colhead{\parbox{6em}{noise model error} }& \colhead{\parbox{5em}{bootstrap significance}} & \colhead{\parbox{6em}{noise model significance}}\\
-\colhead{[hMpc$^{-1}$]} & \colhead{}  & \colhead{[mK$^2$]} &
-\colhead{ [mK$^2$]} &\colhead{\parbox{7em}{[mK$^2$]} }& \colhead{} & \colhead{}
+\colhead{k} & 
+\colhead{redshift}  & 
+\colhead{$\Delta^2$} &
+\colhead{\parbox{6em}{theoretical error} }&
+\colhead{\parbox{6em}{theoretical significance}} &
+\colhead{\parbox{6em}{theoretical detection}} &
+\colhead{\parbox{6em}{bootstrap error }} &
+\colhead{\parbox{6em}{bootstrap significance}} &
+ \colhead{\parbox{5em}{bootstrap detection}} \\
+\colhead{[hMpc$^{-1}$]} & 
+\colhead{}  & 
+\colhead{[mK$^2$]} &
+\colhead{ $\sigma$} & 
+\colhead{}&
+\colhead{\parbox{7em}{[mK$^2$]} }& 
+\colhead{[mK$^2$]}&
+\colhead{$\sigma$} & 
+\colhead{}
 }
 \startdata
-0.1 & 10.29 & 43575.6 & $\pm$ 5729.3 & $\pm$ 676.4 & Det & Det \tabularnewline
-0.1 & 8.54 & 5701.6 & $\pm$ 1168.7 & $\pm$ 279.7 & Det & Det \tabularnewline
-0.1 & 7.94 & 7695.4 & $\pm$ 2044.6 & $\pm$ 229.3 & Det & Det \tabularnewline
-0.1 & 7.55 & 9461.8 & $\pm$ 2761.6 & $\pm$ 153.3 & Det & Det \tabularnewline
-0.2 & 10.29 & 11411.0 & $\pm$ 5039.8 & $\pm$ 5326.1 & Det & Det \tabularnewline
-0.2 & 8.54 & 5855.9 & $\pm$ 2195.8 & $\pm$ 2184.4 & Det & Det \tabularnewline
-0.2 & 7.94 & 5143.7 & $\pm$ 1799.7 & $\pm$ 1785.6 & Det & Det \tabularnewline
-0.2 & 7.55 & 856.3 & $\pm$ 1490.2 & $\pm$ 1190.0 & ULim & ULim \tabularnewline
-0.3 & 10.29 & 34756.5 & $\pm$ 16384.8 & $\pm$ 17922.7 & Det & Det \tabularnewline
-0.3 & 8.54 & 12923.3 & $\pm$ 5373.5 & $\pm$ 7339.6 & Det & Det \tabularnewline
-0.3 & 7.94 & 5204.4 & $\pm$ 4834.5 & $\pm$ 5996.1 & Det & ULim \tabularnewline
-0.3 & 7.55 & 7628.5 & $\pm$ 4810.3 & $\pm$ 3993.6 & Det & Det \tabularnewline
-0.4 & 10.29 & 42910.0 & $\pm$ 30695.0 & $\pm$ 42439.6 & Det & Det \tabularnewline
-0.4 & 8.54 & 24115.2 & $\pm$ 16585.6 & $\pm$ 17370.2 & Det & Det \tabularnewline
-0.4 & 7.94 & 14730.7 & $\pm$ 9930.2 & $\pm$ 14187.9 & Det & Det \tabularnewline
-0.4 & 7.55 & -8601.8 & $\pm$ 14191.5 & $\pm$ 9447.4 & ULim & ULim \tabularnewline
+0.1 & 10.29 & 43600 & $\pm$   700 & 64.4 & Det & $\pm$  5700 & 7.6 & Det \tabularnewline
+0.1 & 8.54 &  5700 & $\pm$   300 & 20.4 & Det & $\pm$  1200 & 4.9 & Det \tabularnewline
+0.1 & 7.94 &  7700 & $\pm$   200 & 33.6 & Det & $\pm$  2000 & 3.8 & Det \tabularnewline
+0.1 & 7.55 &  9500 & $\pm$   200 & 61.7 & Det & $\pm$  2800 & 3.4 & Det \tabularnewline
+0.2 & 10.29 & 11400 & $\pm$  5300 & 2.1 & Det & $\pm$  5000 & 2.3 & Det \tabularnewline
+0.2 & 8.54 &  5900 & $\pm$  2200 & 2.7 & Det & $\pm$  2200 & 2.7 & Det \tabularnewline
+0.2 & 7.94 &  5100 & $\pm$  1800 & 2.9 & Det & $\pm$  1800 & 2.9 & Det \tabularnewline
+0.2 & 7.55 &   900 & $\pm$  1200 & 0.7 & ULim & $\pm$  1500 & 0.6 & ULim \tabularnewline
+0.3 & 10.29 & 34800 & $\pm$ 17900 & 1.9 & Det & $\pm$ 16400 & 2.1 & Det \tabularnewline
+0.3 & 8.54 & 12900 & $\pm$  7300 & 1.8 & Det & $\pm$  5400 & 2.4 & Det \tabularnewline
+0.3 & 7.94 &  5200 & $\pm$  6000 & 0.9 & ULim & $\pm$  4800 & 1.1 & Det \tabularnewline
+0.3 & 7.55 &  7600 & $\pm$  4000 & 1.9 & Det & $\pm$  4800 & 1.6 & Det \tabularnewline
+0.4 & 10.29 & 42900 & $\pm$ 42400 & 1.0 & Det & $\pm$ 30700 & 1.4 & Det \tabularnewline
+0.4 & 8.54 & 24100 & $\pm$ 17400 & 1.4 & Det & $\pm$ 16600 & 1.5 & Det \tabularnewline
+0.4 & 7.94 & 14700 & $\pm$ 14200 & 1.0 & Det & $\pm$  9900 & 1.5 & Det \tabularnewline
+0.4 & 7.55 & -8600 & $\pm$  9400 & -0.9 & ULim & $\pm$ 14200 & -0.6 & ULim \tabularnewline
 \enddata
@@
 \label{tab:data}
-\end{deluxetable*}
+\end{deluxetable}
 
@@
 Techniques like the covariance removal introduced in P14 and also used in this work can differentiate some of this variation from the sky signal; however some combination of improved calibration techniques \citep{Zheng:2014p10467}, model building \citep{Sullivan:2012p9457}, and a better understanding of error covariance  \citep{Liu:2011p8763} may also prove valuable in reducing this systematic.
+Ongoing analysis of more sensitive data sets suggests that the excess seen here can be minimized further with some combination of: fringe rate filters that are better matched to the sky and time dependent calibration (Ali et al, in prep).
 

### 2014-10-21 ms.tex
removing comments for post to arxiv

@@
-\documentclass[fleqn,10pt,iop]{emulateapj}
-%\documentclass[preprint2]{aastex}
+%\documentclass[fleqn,10pt,iop]{emulateapj}
+\documentclass[preprint2]{aastex}
 
@@
 
-%
-%This signal is though to be the richest data set on the sky!  Many telescopes are searching for the power spectral signature of 
-%HI.
-Coarsely, models of 21cm emission can be distilled into two parameters.
-The redshift at which point the universe was 50\% ionized ($z_i$) and the transition time from mostly neutral to mostly ionized $dz$.
-Thus full coverage of the spectrum is essential.
-
-%This paper extends the result ARP2013a to cover the redshift range XXX.
-
 \section{Observations}
@@
   First, the raw data are down-selected to just the 70 $\sim$30m long baselines described in Section \ref{sec:observations}.  
-%I know this isn't technically true, but I think its important that the downselect is obvious up front.
    The visibilities are then compressed in the frequency and time directions by removing delay modes and highest frequency fringe rates corresponding to a 300m baseline (the longest baseline in the array).
@@
 \end{figure*}
-%The success of the remainder of the processing stages hinges on the spectral stability of the data, which can easily be assessed by comparing redundant baselines.
-In figure XXX we see the r.m.s.
-of the ratios between redundant calibrations, before and after the calibration has been applied.
+
 \begin{figure}
@@
 
-%This step is the first point at which signal loss could occur.
-To measure the loss we add white noise into the data at a level sufficient to double the system temperature (to make it easily detectable on the output), subtract  and then measure the fractional difference post filtration.
-We recover XXX\% of the injected signal.
-
-%delay transform, covariance projection (lossless, and lossy), bin and bootstrap average,
 \subsection{Power Spectrum}
@@
 Physically the two signs correspond to the two halves of the sky.}.
+For simplicity, measurements are averaged with equal weights.
 
@@
 The theoretical noise power spectrum (dashed line in Figures \ref{fig:pspecs} and \ref{fig:P14compare} and grey region in Figure \ref{fig:slices}) is estimated using the method described in \cite{Pober:2013p9581}, assuming a system temperature of 550K and the observing scheme described in Section \ref{sec:observations}.
+In general the theoretical error bars are smaller than those derived by the bootstrap.
+This is most likely due to the use of uniform weighting when combining power spectrum measurements, rather then the inverse variance weights assumed in the sensitivity calculation.
+
  In all figures the noise levels are plotted for comparison with the power spectrum values rather than the error bars, i.e.
@@
 Under this scheme, we would expect a noise dominated signal to yield around eight ``upper limits'' with an equal number of positive and negative data points.
-Here we find only two to three upper limits while rest we classify as detections.
+Here we find only two upper limits while the rest we classify as low significance detections of excess power.
 In general, most detections are within 1 - 2 $\sigma$ but are all positively biased,  with the significance rising steeply at high redshifts and low $k$s.%, the quadratic sum of all errors above $k=0.1$ is 6.6 $\sigma$ indicating that the excess is largely coherent across the data set.
 
-\subsection{The Excess Signal}
+\subsection{The Excess Power}
 
@@
 \colhead{[mK$^2$]} &
+\colhead{[mK$^2$]} &
 \colhead{ $\sigma$} & 
@@
 \colhead{\parbox{7em}{[mK$^2$]} }& 
-\colhead{[mK$^2$]}&
 \colhead{$\sigma$} & 
@@
 \startdata
-0.1 & 10.29 & 43600 & $\pm$   700 & 64.4 & Det & $\pm$  5700 & 7.6 & Det \tabularnewline
-0.1 & 8.54 &  5700 & $\pm$   300 & 20.4 & Det & $\pm$  1200 & 4.9 & Det \tabularnewline
-0.1 & 7.94 &  7700 & $\pm$   200 & 33.6 & Det & $\pm$  2000 & 3.8 & Det \tabularnewline
-0.1 & 7.55 &  9500 & $\pm$   200 & 61.7 & Det & $\pm$  2800 & 3.4 & Det \tabularnewline
-0.2 & 10.29 & 11400 & $\pm$  5300 & 2.1 & Det & $\pm$  5000 & 2.3 & Det \tabularnewline
-0.2 & 8.54 &  5900 & $\pm$  2200 & 2.7 & Det & $\pm$  2200 & 2.7 & Det \tabularnewline
-0.2 & 7.94 &  5100 & $\pm$  1800 & 2.9 & Det & $\pm$  1800 & 2.9 & Det \tabularnewline
-0.2 & 7.55 &   900 & $\pm$  1200 & 0.7 & ULim & $\pm$  1500 & 0.6 & ULim \tabularnewline
-0.3 & 10.29 & 34800 & $\pm$ 17900 & 1.9 & Det & $\pm$ 16400 & 2.1 & Det \tabularnewline
-0.3 & 8.54 & 12900 & $\pm$  7300 & 1.8 & Det & $\pm$  5400 & 2.4 & Det \tabularnewline
-0.3 & 7.94 &  5200 & $\pm$  6000 & 0.9 & ULim & $\pm$  4800 & 1.1 & Det \tabularnewline
-0.3 & 7.55 &  7600 & $\pm$  4000 & 1.9 & Det & $\pm$  4800 & 1.6 & Det \tabularnewline
-0.4 & 10.29 & 42900 & $\pm$ 42400 & 1.0 & Det & $\pm$ 30700 & 1.4 & Det \tabularnewline
-0.4 & 8.54 & 24100 & $\pm$ 17400 & 1.4 & Det & $\pm$ 16600 & 1.5 & Det \tabularnewline
-0.4 & 7.94 & 14700 & $\pm$ 14200 & 1.0 & Det & $\pm$  9900 & 1.5 & Det \tabularnewline
-0.4 & 7.55 & -8600 & $\pm$  9400 & -0.9 & ULim & $\pm$ 14200 & -0.6 & ULim \tabularnewline
+0.1 & 10.29 & 43600 & $\pm$   400 & 105.6 & Det & $\pm$  5700 & 7.6 & Det \tabularnewline
+0.1 & 8.54 &  5700 & $\pm$   200 & 33.0 & Det & $\pm$  1200 & 4.9 & Det \tabularnewline
+0.1 & 7.94 &  7700 & $\pm$   100 & 56.2 & Det & $\pm$  2000 & 3.8 & Det \tabularnewline
+0.1 & 7.55 &  9500 & $\pm$   100 & 122.7 & Det & $\pm$  2800 & 3.4 & Det \tabularnewline
+0.2 & 10.29 & 11400 & $\pm$  3300 & 3.5 & Det & $\pm$  5000 & 2.3 & Det \tabularnewline
+0.2 & 8.54 &  5900 & $\pm$  1400 & 4.3 & Det & $\pm$  2200 & 2.7 & Det \tabularnewline
+0.2 & 7.94 &  5100 & $\pm$  1100 & 4.8 & Det & $\pm$  1800 & 2.9 & Det \tabularnewline
+0.2 & 7.55 &   900 & $\pm$   600 & 1.4 & Det & $\pm$  1500 & 0.6 & ULim \tabularnewline
+0.3 & 10.29 & 34800 & $\pm$ 11000 & 3.2 & Det & $\pm$ 16400 & 2.1 & Det \tabularnewline
+0.3 & 8.54 & 12900 & $\pm$  4600 & 2.8 & Det & $\pm$  5400 & 2.4 & Det \tabularnewline
+0.3 & 7.94 &  5200 & $\pm$  3600 & 1.4 & Det & $\pm$  4800 & 1.1 & Det \tabularnewline
+0.3 & 7.55 &  7600 & $\pm$  2000 & 3.8 & Det & $\pm$  4800 & 1.6 & Det \tabularnewline
+0.4 & 10.29 & 42900 & $\pm$ 26000 & 1.7 & Det & $\pm$ 30700 & 1.4 & Det \tabularnewline
+0.4 & 8.54 & 24100 & $\pm$ 10800 & 2.2 & Det & $\pm$ 16600 & 1.5 & Det \tabularnewline
+0.4 & 7.94 & 14700 & $\pm$  8500 & 1.7 & Det & $\pm$  9900 & 1.5 & Det \tabularnewline
+0.4 & 7.55 & -8600 & $\pm$  4800 & -1.8 & Det & $\pm$ 14200 & -0.6 & ULim \tabularnewline
 \enddata
@@
 Another active point of investigation is aimed at alternative windows to the Blackman-Harris used in the wide-band CLEAN (or wholly distinct algorithms) that could better remove foregrounds from the edges of the band while still limiting foreground bleed into the EoR window.
+Weighting of the final power spectrum average is also a subject for further improvement.
+In this analysis we have used uniform weights; this is also the default assumption of most sensitivity calculations, including the one used here.
+Uniform weighting is optimal if the final measurements are limited by something other than noise, while inverse variance is appropriate for purely noise limited measurements.
+Given how close the error bars are to noise dominated, future analysis will investigate using different estimates of the variance for optimal weighting.
 

### 2014-10-21 generate_AAS.py
updating AAS script

@@ -8,15 +8,13 @@ import re,os
 
 #target = 'arxiv_submission'
 project=os.path.basename(os.getcwd())
-target = 'apj_submission_1'
-
+#target = 'apj_submission_1'
+target= 'arxiv_submission_1'
 if target.startswith('arxiv'):
     tex = open('ms-arxiv.tex').read()
-    plots = re.findall('\\includegraphics.*\{([^}]+)\}',tex)
 else:
     tex = open('ms.tex').read()
-#    plots = re.findall('\{figures\/(\S*)\}',tex)
-    plots = re.findall('\\includegraphics.*\]{(.*)\}',tex)
+plots = re.findall('\\includegraphics.*\]{(.*)\}',tex)
 
 #clean out the '}'s from the figure paths
 l = ['a','b','c','d']
@@ -44,8 +42,7 @@ for i,plot in enumerate(plots):
     if target.startswith('arxiv'):
         print "cp ",plot_filenames[i],target
         os.system('cp %s %s/'%(plot_filenames[i],target))
-        tex = tex.replace(plot,plot)
-        print plot
+        tex = tex.replace(plot,plot.replace('figures/',''))
     else:
         print "cp ",plot_filenames[i],target
         os.system('convert %s %s/%s'%(plot_filenames[i],target,eps_plots[i]))
@@ -61,9 +58,9 @@ out = open('%s/ms.tex'%target,'w')
 out.write(tex)
 #os.system('cp figures/tab1.txt figures/tab2.txt %s'%target)
 if target.startswith('arxiv'):
-    os.system('tar -cjf %s.tar.bz2 %s/*'%(target,target))
+    os.system('tar -cjf %s.tar.bz2 -C %s .'%(target,target))
 else:
-    os.system('tar -czf {project}_{target}.tar.bz2 {target}/*'.format(target=target,project=project))
+    os.system('tar -czf {project}_{target}.tar.bz2 -C {target} .'.format(target=target,project=project))
 
 
 

### 2014-10-31 ms.tex
finished reviewer comments, sent to collab

@@
 Relative to this thermal noise, most spectra exhibit and excess of power at a few sigma.
-The likely sources of this excess include residual foreground leakage, particularly at the highest redshift, and unflagged RFI.
+The likely sources of this excess include residual foreground leakage, particularly at the highest redshift, unflagged RFI, and calibration errors.
 We conclude by discussing data reduction improvements that promise to remove much of this excess.
@@
 Therefore, while a detection of the 21 cm signal at even a single frequency would be a tremendous breakthrough, analysis techniques must be developed to capitalize on the wide bandwidths of the current generation of high-redshift 21cm telescopes.
+
 Using the same data set as P14, this paper presents improved upper limits on the HI power spectrum over the redshift range $10.7>z>7.2$.
+The power spectrum error bars, calculated both from theory and using a bootstrap, are broadly consistent with a detection of power at the 1$\sigma$ level which can be interpreted as some combination of residual foregrounds, analysis errors and reionization.
+To determine which components are the mostly likely causes we compare the power spectrum excess measurements of the system noise found by differencing adjacent channels and redundant baselines.
+
  In Section  \ref{sec:observations} we summarize the observations, in Section \ref{sec:obs_meth} review the reduction methodology, we present the new upper limits in Section \ref{sec:results}, and in Section \ref{sec:conclusion} we offer conclusions and discussion of future work.
@@
 In most interferometers the locations of the antennas are optimized such that each baseline samples a different Fourier mode of the sky; this is the ideal case for reconstructing images where each mode contains different information.
-For a power spectrum measurement the key metric is sensitivity per mode, rather than number of modes.
-A grid configuration allows many samples of each cosmological mode, to be averaged to a high sensitivity before being combined with other Fourier modes.
+For a power spectrum measurement the key sensitivity metric is sensitivity per mode, rather than number of modes.
+A grid configuration allows many samples of each cosmological mode, to be averaged to a high sensitivity before being combined with other Fourier modes; though at the cost of losing foreground information.
+
 The PAPER South Africa 32 antenna deployment (PSA32) was arranged in a 4$\times$8 grid with a column spacing of 30m and a row spacing of 4m.
@@
 As we only have three baseline types of nearly the same length the wedge manifests as a single value of $k\sim k_\parallel$, below which foregrounds are expected to dominate.
+
+A grid configuration is sufficiently different from a traditional interferometer antenna layout that 
+there are many tradeoffs to consider.
+In this configuration foregrounds cannot be imaged, and therefor can only be filtered or blindly subtracted.
+Experimentally, the loss of foreground imaging is offset by an enhanced experimental precision.
+The grid configuration makes repeated measurements of the \emph{same} modes with \emph{independent} antennas, thus the per-antenna gains are a tractable linear algebra problem that does not depend on a sky model \citep{Liu:2010p10391}.
+This same redundancy also enables the identification of systematically deviant samples, allows us to avoid noise bias, and provides another axis along which to bootstrap.
+These are believed to provide the greatest handle on known systematics.
+This solution has a small number of free parameters which must be calibrated to a model of the sky to obtain a power spectrum calibrated to a standard flux scale.
+Flux and overall bandpass calibration is the area most impacted by the small number of sampled spatial modes.
+In practice, the use of synthesis rotation allows for adequate coverage of the brightest source in the sky which we use as our flux standard (see section \ref{sec:calibration}.
+
  The details of each application of the delay/fringe-rate transform will be laid out in the following sections as we provide a brief walk-through of the processing pipeline.
@@
 \subsection{Calibration}
+\label{sec:calibration}
 We model the gain as a per-antenna amplitude and a phase slope -physically a single time delay- and single real, low order polynomial passband for all antennas.
 Because the array samples correlations redundantly, the relative calibration between antenna is numerically overdetermined and tractable as a linear algebra problem \citep{Liu:2010p10391}.
+Importantly this solution does not require a model of the sky or instrument (beyond the assumption of redundancy).
 As described in P14, we compute the ratio between redundant baselines, fit for a gain and phase slope and then algebraically solve for a per-antenna solution.
@@
 However, to obtain a correctly normalized power spectrum, it is necessary to set a flux scale for which we need a flux measurement of a calibration standard.
-For this we form a beamform on a known, bright calibration source.
+For this we form a beam on a known, bright calibration source.
 By itself, the redundant calibration does not contain enough phase information to phase coherently to a sky location, there remain two free phase parameters which cannot be solved by redundancy alone.
@@
 With the delays in place we are now able form a beam on Pictor A and (for each channel) set the overall amplitude to the calibration value of 382 (f/150MHz)$^{-0.77}$Jy found in \cite{jacobs:2013b}.
+The accuracy of this step is ultimately limited by the accuracy of the beam-forming on Pictor to get a spectrum.
+The principle sources of error include beam modeling error and incomplete uv coverage.
+Correction for both was explored in detail by \citet{jacobs:2013b} used PAPER data to measure the complete Pictor A track through the primary beam.
+The light curve matched the beam model well over the full hour angle range, suggesting that the beam model is well understood.
+When integrated over the full track, the uv coverage was found to be very uniform.
+The noise weighted average was found to yield a 6\% flux error.
+This error in the overall offset of all power spectra is sub-dominant to the observed power spectrum error bars.
+%which showed that in measuring Pictor over the full range of hour angles one integrates over many beam pointings and uv-samples.
+Comparison of multiple sources with catalog values found on overall standard deviation of \~ 6\%.
 
@@
 This leaves us with 40 delay samples on each of 70 baselines which are divided into three redundant groups.
-Within these groups we cross correlate delay spectra between different redundant baselines.
-The cross multiplication of the same delay modes between different redundant baselines provides an unbiased estimate of the power spectrum.
-These ``sky-like'' correlations should be identical between all redundant baselines to within the level of the noise, while all other cross multiplications between delay modes should not be correlated between different baselines.
-In practice  these non-sky-like modes do occasionally have significant power which leaks into the correlations which sample the power spectrum.
-These are removed by iteratively dividing the covariance into a model of systematics and a model of sky-like emission and then projecting out large residual modes.
-This is done by dividing the baselines into different groups such that all cross-multiplications are done without introducing noise bias.
-For more see Appendix C of P14.
+Each redundant baseline samples the same mode on the sky but with different noise.
+The cross correlation between baselines gives an estimate of the power spectrum without any additional noise bias.
+In practice we found that not all cross power spectra agreed as they should were the result dominated by the sky; some instrumental artifacts remained.
+Though only a small number of modes are directly contaminating a larger number are indirectly affected by the delay spectrum point spread function.
 
-The residual elements of the correlation matrix corresponding to cross-multiplication of matching delay bins between different baselines are all un-biased samples of the power spectrum.
-To estimate the final power spectrum and its uncertainty we compute the mean and variance of many random randomly-selected subsamples, sampling the dimensions: sidereal time, redundant baseline pair, and delay sign\footnote{As visibilities are complex, both the positive and negative delays  carry  information.
-Physically the two signs correspond to the two halves of the sky.}.
+This contamination is removed by iteratively subtracting an average of a bootstrap subset of baselines and then building a model of the residual covariance.
+This is then divided out of the data covariance.
+Within this process the redundant samples are subdivided into four groups such that all cross-multiplications are done between the groups to avoid introducing noise bias into the removed covariance.
+(For more see Appendix C of P14.)  The estimated cross-power delay spectrum is recorded for each bootstrap.
+The final power spectrum and its uncertainty are the mean and variance of randomly selected subsets of these bootstraps.
+This final bootstrap samples the dimensions: sidereal time, redundant baseline pair, delay sign\footnote{As visibilities are complex, both the positive and negative delays  carry  information.
+Physically the two signs correspond to the two halves of the sky.}, and mean-sky bootstrap.
 For simplicity, measurements are averaged with equal weights.
@@
 \includegraphics[width=0.9\textwidth]{figures/PAPER_32T_new_redshifts.png}
-\caption{ \label{fig:pspecs} PAPER power spectra at four redshifts.
+\caption{ \label{fig:pspecs} PAPER power spectra at four redshifts with error bars derived via bootstrap subsampling.
 On the top, $P_k$ spectra (a simple units change from the raw delay spectrum) provide a useful diagnostic on foreground rejection, while on the bottom we plot in $\Delta^2$ cosmological units with 1 $\sigma$ error bars.
@@
 The root-mean-square of differenced visibilities measures how well we've removed foregrounds.
-In Figure \ref{fig:noise} we examine $T_{rms}$ as calculated by differencing between adjacent channels and between redundant baselines.
-The square of this difference is then averaged over the full LST range probed by the power spectra (1h-9h).
-Note that this operation combines all LST modes as though they were repeated measurements of the same sky when in practice the power spectrum values are averaged in quadrature over different LST bins.
-This provides a lower bound on the sensitivity possible in the data set and a sensitive probe on foreground residuals or other sources of excess power.
-Using all the data in this way, we expect to illuminate rare outliers that effect the bottom line.
+In Figure \ref{fig:noise} we examine $T_{rms}$ as calculated by differencing between adjacent channels and between redundant baselines, squaring and averaging in LST.
+The channel to channel difference is expected to be dominated by noise and residual foregrounds.
+The baseline to baseline difference will contain all of these as well as any additional variance due to calibration errors.
+These errors will all manifest as an excess rms compared to the theoretical noise level.
+We can also examine the rms at different sensitivity levels, for example before and after long term averaging.
+Comparing different integration lengths we can see if its averaging down properly.
 
@@
 \subsection{The Excess Power}
-
+\label{sec:excess power}
 As these measurements have a maximum possible sensitivity level which is still two orders of magnitude away from theoretical reionization levels we consider the likelihood that these detections are evidence of reionization to be small.
@@
 Recall that the wide-band foreground model we have subtracted was built by weighting the entire band by a Blackman-Harris window function which provides a much higher isolation of foregrounds inside the horizon.
-The tradeoff that is made is that data in the outer parts of the spectrum are heavily down weighted.
-For example data at 126MHz are down-weighted $\sim$75\% compared to 150Mhz.
-This amounts to making an assumption about the spectral smoothness of the data, namely that 126MHz can be mostly modeled by extrapolation from 150MHz.
+The tradeoff is that data in the outer parts of the spectrum are heavily down weighted.
+For, in the subtracted foreground model, data at 126MHz are down-weighted $\sim$75\% compared to 150Mhz.
+This amounts to making an assumption about the spectral evolution of the data, namely that 126MHz can be mostly modeled by extrapolation from 150MHz.
 The precision of this subtraction can be judged by examining residuals inside the horizon in the $P(k)$ power spectra plotted in the upper part of Figure \ref{fig:pspecs}\footnote{Note that this notch is not present in Figure \ref{fig:P14compare} because the foregrounds have been left in for comparison with P14.}.
@@
 If these points were purely foregrounds, the error bars would be noise dominated, the fact that the excess is not much larger then its variation suggests a source of error not accounted in our analysis.
-Likely culprits include incomplete flagging of interference, instrumental glitches, or time of frequency dependent calibration variation which introduces differences between redundant baselines.
+Likely culprits include incomplete flagging of interference, instrumental glitches, time and frequency dependent calibration variation which introduces differences between redundant baselines, and imperfect instrumental redundancy (e.g.
+variation of primary beams between antennas or deviations from redundant antenna placement).
 All of which have are part of ongoing analysis efforts which are briefly discussed in the conclusions.
-All of these effects have the effect of adding time and frequency dependence to the bright foregrounds limiting the effectiveness of the filter.
+Most of these effects have the effect of adding time and frequency dependence to the bright foregrounds limiting the effectiveness of the filter.
 Thus, just like foregrounds, the strong excess drops rapidly towards higher $k$ modes.
-Above $k$ of 0.1 hMpc$^{-1}$  the power spectra are generally positively biased, though most close to 2 $\sigma$ or less using either theoretical and bootstrap error bars.
+Above $k$ of 0.1 hMpc$^{-1}$  the power spectra are generally positively biased, though most close to 1 $\sigma$.
 

### 2014-11-10 ms.tex
revised 4.3 to be more clear about what we think the excess is

@@
 there are many tradeoffs to consider.
-In this configuration foregrounds cannot be imaged, and therefor can only be filtered or blindly subtracted.
+In this configuration foregrounds cannot be imaged, and therefor can only be filtered or subtracted by forward modeling.
 Experimentally, the loss of foreground imaging is offset by an enhanced experimental precision.
@@
 \label{sec:excess power}
-As these measurements have a maximum possible sensitivity level which is still two orders of magnitude away from theoretical reionization levels we consider the likelihood that these detections are evidence of reionization to be small.
-Far more likely is that we are limited by a residual foreground or systematic arising from approximations in the data analysis.
-The statistical significance of the residual --2$\sigma$ after averaging over the entire season-- is low enough that it is not possible to break it down along likely axes like sidereal time or time of day to probe its origin.
+%As these measurements have a maximum possible sensitivity level which is still two orders of magnitude away from theoretical reionization levels, we consider the likelihood that these detections are evidence of reionization to be small.
+Far more likely, is that we are limited by a residual foreground or systematic arising from approximations in the data analysis.
+  The statistical significance of the residual --$<2\sigma$ for most points-- is low enough that it is not possible to investigate possible sources of systematic bias --e.g.
+dependence on sidereal time, time of day, or other dimensions.
 Such studies must await higher sensitivity measurements with more antennas.
-However, there are several additional features that seem to suggest an excess not caused by high redshift HI.
+Without data sensitive enough to dissect the residuals we cannot identify their origin, but we can make a couple observations.
 
-The spectral signature of the measured excess is far more similar to a foreground residual then to theoretical HI models.
-The predicted signature of the HI signal during reionization is one of a relatively short burst of power at small $k$ values as the ionized bubbles are briefly at their largest, before the signal from the ever increasing ionized medium drops sharply \citep{Pritchard:2008p8123}.
-In contrast, this observed residual is relatively constant with redshift while rising at both ends of the band, mimicking the RMS noise curve in Figure \ref{fig:noise}.
+The redshift 7.68 bin reproduces the excess seen in P14 where it was observed that a complete lack of X-Ray heating from early stars would be required to produce a signal two orders of magnitude above expected theoretical levels.
+Extending this analysis to the spectral dimension we find more significant deviation from theoretical behavior.
+The theoretical signature of the HI signal during reionization is one of a relatively short burst of power at small $k$ values, as the ionized bubbles are briefly at their largest, before the signal from the ever increasing ionized medium drops sharply \citep{Pritchard:2008p8123}.
+The observed residual is relatively constant with redshift, rising at both ends of the band, mimicking the RMS noise curve in Figure \ref{fig:noise}.
 This suggests that the accuracy of the foreground filter decreases towards the edges of the band.
@@
 The tradeoff is that data in the outer parts of the spectrum are heavily down weighted.
-For, in the subtracted foreground model, data at 126MHz are down-weighted $\sim$75\% compared to 150Mhz.
-This amounts to making an assumption about the spectral evolution of the data, namely that 126MHz can be mostly modeled by extrapolation from 150MHz.
+For example, in the subtracted foreground model, data at 126\,MHz are down-weighted $\sim$75\% compared to 150\,Mhz.
+This amounts to making an assumption about the spectral evolution of the data, namely that 126\,MHz can be mostly modeled by extrapolation from 150\,MHz.
 The precision of this subtraction can be judged by examining residuals inside the horizon in the $P(k)$ power spectra plotted in the upper part of Figure \ref{fig:pspecs}\footnote{Note that this notch is not present in Figure \ref{fig:P14compare} because the foregrounds have been left in for comparison with P14.}.
-The relative depth of the redshift 10.5 P(k) foreground residual is substantially shallower then at 7.55 and is generally correlated with the amount of residual found at higher $k$s.
+The relative depth of the redshift 10.5 $P(k)$ foreground residual is substantially shallower then at 7.55 and is generally correlated with the amount of residual found at higher $k$s.
 
@@
 Cuba Astronomical Foundation.
-J.C.P.
+D.C.J.
+is supported by an NSF Astronomy and Astrophysics Fellowship under award AST \# 1401708 while J.C.P.
 is supported by an NSF Astronomy and Astrophysics Fellowship under award AST \#1302774.
